### 2007-04-24 doc/hed/counter.tex
Counter library requirements and API proposal

\documentclass[a4paper,11pt]{article}

\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[small]{caption}
\usepackage{url}

\title{ARC Counter Library\\
Requirements and Design}
\author{Markus Nord\'en}

\begin{document}

\maketitle

\begin{abstract}
  This is a draft summary of requirements and a proposition of an API
  for the counter library of the ARC HED. Comments are encouraged and
  expected. In particular, feedback is needed on globality
  (section~\ref{globality}), persistency (section~\ref{persistency})
  and the suggested API (section~\ref{api}).

  Please, send any feedback to: \url{nordugrid-discuss@nordugrid.org}
\end{abstract}

\section{Introduction}

During the ``Uppsala HED Implementation F2F Meeting'', it was decided
that a counter library shall be provided for services and other
components to use for e.g. allocation and housekeeping of resources.

This document contains requirements on that counter library based on
e-mail discussions on the \url{nordugrid-discuss} e-mail list.

\section{Requirements}

\subsection{Purpose}

The purpose of a counter is to provide housekeeping of a resource that
can be allocated and deallocated such as e.g. disk space or network
bandwidth. The counter itself will not be aware of what kind of
resource it limits the use of. Neither will it be aware of what unit
is being used to measure that resource. Counters are thus very similar
to semaphores.

The users of the resource must thus first call the counter in order to
allocate (make a reservation of) an appropriate amount of the
resource, then use the resource and finally call the counter again to
deallocate what was previously allocated.

Furthermore, the users of the resource must agree on what unit they
use to measure the resource, e.g. if disk space is measured in GB, MB,
kB or single bytes. The counter itself will thus only contain a number.

\subsection{Counter types}

Integers provide sufficient resolution for internal representation of
available resources. Floating point numbers could be an alternative,
but no such requirement has been expressed.

If there should arise a need to represent quantities of a resource
that are smaller than one unit, the choice of unit for the
corresponding counter can be reconsidered.

\subsection{Atomicity and Synchronization}

Counters must be able to handle concurrent operations in a consistent
manner. This implies that all operations must be atomic. Furthermore,
no operation may be performed until it is guaranteed that it will
succeed, i.e. a call to the allocation method must block until there
is a sufficiently large amount of the resource free.

No requirements for simultaneous operations on sets of counters have
been requested.

\subsection{Globality}
\label{globality}

Counters for different purposes (limiting different kinds of
resources) may need to have different scopes:

\begin{description}
\item[Single thread] The counter can only be accessed from within a
single thread. There is no concurrency or need for synchronization and
a special counter class is superfluos.
\item[Single process] (several threads) The counter can be accessed
from all threads within a single process, but is not accessible from
outside that process.
\item[Single computer] (several processes) The counter can be accessed
from all processes running on CPUs with access to a common shared
memory. This requires inter-process communication.
\item[Single cluster] (several computers) The counter can be accessed
from any process running on any computer within a cluster.
\item[Entire grid] The counter can be accessed from any process
running on any computer connected to a certain grid -- probably by
means of a counter service.
\end{description}

It deserves to be mentioned here that the scope of the counter itself
is not necessarily the same as the scope of the resource it limits the
use of. For example, there may be counter available only on the front
end machine of a cluster that keeps track of the number of free CPUs
in the entire cluster. This counter can be used by a process that runs
on the front end machine but starts jobs on other computers in the
cluster.

Requests have been made for single process counters, single computer
counters and single cluster counters.

\subsection{Persistency}
\label{persistency}

Different kinds of counters have different lifetimes. The lifetime of
single computer counters is an open question.

\begin{description}
\item[Process lifetime] The counter is initialized at process start up
  and vanishes when the process finishes. Applicable only for single
  process counters.
\item[System lifetime] The counter ts initialized at system start up
  and is reset upon system restart. Possibly applicable for single
  computer counters.
\item[Eternal lifetime] The counter is initialized once and is then
  available until it is explicitly removed. Applicable for single
  cluster, entire grid and possibly also for single computer counters.
\end{description}

\subsection{Identification of counters}

Counters are identified by unique names. The form of a name will
depend of the globality of a counter.

Single process counters may, for example, be implemented as global
objects accessible from all threads of the process. The ``name'' of a
counter will in this case be the address of the corresponding counter
object. This address can be referred to by an arbitrary number of
pointers and references as well as one or zero ``ordinary variables'',
i.e. just as any other object.

A single computer counter or a single cluster counter may be accessed
through an arbitrary number of counter objects in different processes.
There may thus be many counter objects that correspond to the same
counter and manipulates it in a synchronized way. Upon construction of
such a counter object, the name of the counter shall be provided. Such
names can follow the same rules as e.g. identifiers in the C
programming language. Furthermore, the name of some global object,
e.g. a file, will probably also be required as an ``environment'' for
the counters.

The name of an entire grid counter will probably consist of the URI of
a counter service (the ``environment'' part) and a counter name as
described above.

\subsection{Priority}

In some cases it may be desirable to let important tasks exceed the
limit of a counter to a certain extent. This may be implemented by
means of a credit. That an allocation is prioritized may be expressed
by means of an extra, optional, parameter to the allocation method.

\subsection{Time Awareness}

The basic use of a counter is that a certain amount of it is acquired
by a call to an allocation method and subsequently returned by a call
to a deallocation method.

There has also been requests for self expiring allocations, i.e. that
it is specified at allocation for how long that acquisition shall
last. After that time has elapsed, deallocation will occur
automatically.

It has also been mentioned that reservation of resources in advance
would be a nice feature, but this is considered to be less important.

\section{Design and API}
\label{api}

There will be an abstract base class that defines the interface of a
counter, i.e. what operations can be performed on it. The different
kinds of counters (different globality) will be implemented as
subclasses of that base class.

The signature of the constructors of the different classes will be
determined later based on the needs of the implementation.

The following operations will be available for counters:

\subsection{void wait(int amount, bool prioritized, int duration)}

Blocks until the requested amount is free and allocates it.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be allocated.
\item[prioritized] True if this is a prioritized request, false
  otherwise. Optional, default value is false.
\item[duration] Zero for allocations that subsequently will be
  explicitly deallocated. If positive, duration of a self expiring
  allocation. Optional, default value zero.
\end{description}

\subsection{bool tryWait(int amount, bool prioritized, int duration)}

Tries to allocate the requested amount and returns immediately.
Returns true if the allocation succeded and false if it did not
succeed.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be allocated.
\item[prioritized] True if this is a prioritized request, false
  otherwise. Optional, default value is false.
\item[duration] Zero for allocations that subsequently will be
  explicitly deallocated. If positive, duration of a self expiring
  allocation. Optional, default value zero.
\end{description}

\subsection{void post(bool amount)}

Deallocates a certain amount.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to be deallocated.
\end{description}

\subsection{int getLimit()}

Returns the current limit of the counter.

\subsection{int setLimit(int limit)}

Sets the limit of the counter. Returns the new limit.

\paragraph{Parameters:}
\begin{description}
\item[limit] The new limit.
\end{description}

\subsection{int changeLimit(void amount)}

Changes the limit of the counter by a certain amount. Returns the new
limit.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to add to the the limit. May be positive or
  negative.
\end{description}

\subsection{int getCredit()}

Returns the current credit of the counter.

\subsection{int setCredit(void credit)}

Sets the limit of the counter. Returns the new limit.

\paragraph{Parameters:}
\begin{description}
\item[credit] The new credit.
\end{description}

\subsection{int changeCredit(void amount)}

Changes the credit of the counter by a certain amount. Returns the new
credit.

\paragraph{Parameters:}
\begin{description}
\item[amount] The amount to add to the the credit. May be positive or
  negative.
\end{description}

\subsection{int getValue()}

Returns the current value of the counter, i.e. the number of units
currently available for allocation.

\section{Implementation}

The implementation will be based on ``lower level'' libraries. Posix
and System V semaphores have been considered so far but have been
rejected due to limited functionality and portability issues. The
current candidate is glibmm, possibly combined with file storage for
counters that cannot reside in memory.

\end{document}