### 2008-12-05 doc/sec/Sec_paper.tex
More sections of the sec doc

% This file was converted to LaTeX by Writer2LaTeX ver. 0.5
% see http://www.hj-gym.dk/~hj/writer2latex for more info
\documentclass{article}
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{array}
\usepackage{hhline}
\usepackage{hyperref}
\hypersetup{pdftex, colorlinks=true, linkcolor=blue, citecolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue, pdftitle=, pdfauthor=qiangwz, pdfsubject=, pdfkeywords=}
\usepackage[pdftex]{graphicx}
% List styles
\newcommand\liststyleLi{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleLii{%
\renewcommand\labelitemi{${\bullet}$}
\renewcommand\labelitemii{${\circ}$}
\renewcommand\labelitemiii{${\blacksquare}$}
\renewcommand\labelitemiv{${\bullet}$}
}
\newcommand\liststyleLiii{%
\renewcommand\labelitemi{${\bullet}$}
\renewcommand\labelitemii{${\circ}$}
\renewcommand\labelitemiii{${\blacksquare}$}
\renewcommand\labelitemiv{${\bullet}$}
}
\newcommand\liststyleLiv{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
\newcommand\liststyleLv{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\arabic{enumii}}
\renewcommand\theenumiii{\arabic{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
% Page layout (geometry)
\setlength\paperwidth{8.5in}
\setlength\paperheight{11in}
\setlength\voffset{-1in}
\setlength\hoffset{-1in}
\setlength\topmargin{0.7874in}
\setlength\oddsidemargin{0.7874in}
\setlength\textheight{9.4251995in}
\setlength\textwidth{6.9251995in}
\setlength\footskip{0.0cm}
\setlength\headheight{0cm}
\setlength\headsep{0cm}
% Footnote rule
\setlength{\skip\footins}{0.0469in}
\renewcommand\footnoterule{\vspace*{-0.0071in}\setlength\leftskip{0pt}\setlength\rightskip{0pt plus 1fil}\noindent\textcolor{black}{\rule{0.25\columnwidth}{0.0071in}}\vspace*{0.0398in}}
% Pages styles
\makeatletter
\newcommand\ps@Standard{
  \renewcommand\@oddhead{}
  \renewcommand\@evenhead{}
  \renewcommand\@oddfoot{}
  \renewcommand\@evenfoot{}
  \renewcommand\thepage{\arabic{page}}
}
\makeatother
\pagestyle{Standard}
\title{}
\begin{document}
{\centering
Towards cross-middleware authentication and single sign-on for ARC grid
middleware
\par}

\bigskip

{\centering
Abstract
\par}

\ In order to make the access to Grids as simple as possible, security
is one of the most important challenges in production Grids, especially
when the Grid applications are across multiple administration domains
as well as across heterogeneous Grid middlewares, such as some widely
scaled eScience applications which needs to coordinate resources
sharing among collection of autonomous institutions with different Grid
middlewares running on these resources. In this paper, we describe the
security implementation and consideration in new version of Advanced
Resource Connector (ARC) middleware, where the above issues have been
exploited. The main goal of ARC implementation in terms of security is
to let the middleware be capable to \ interoperate with other Grid
middlewares by leveraging standard specifications. The key aspect of
the work we have done is to enhance the current proxy certificate based
delegation and authentication by utilizing the Security Assertion
Markup Languages (\textit{SAML}) single sign-on (SSO) profile,
WS-Security in order to achieve cross-domain authentication. \

\bigskip

\liststyleLi
\begin{enumerate}
\item[] {\centering
1. Introduction
\par}
\end{enumerate}
Grids provides the technology for large scaled, cross domains
collaboration where the users work on resources and sharing data which
are distributed across domains. This kind of collaboration is called
virtual organization (VO)[ref1]. The resources that contribute to the
VO could come from independent administrative domains, and even require
different security approaches for control the access to themselves. The
same heterogeneous situation even critically applies to users who need
to access the VO, because users holding one kind of credential could
need to travel through multiple resources with different security
approaches. To solve the above issues, the Grid middleware should
provide the following ability: The user which participates the VO
should be able to access data and resources without being required to
constantly do authentication at each resource or data; The VO should be
able to easily manage the privileges or memberships to the users, and
VO together with its resources should be able to easily manage the
authorization, given the condition that the resources are actually
owned by different administrative domains.

In grid community, Grid Security Infrastructure (GSI)[ref2] is the
\textit{de facto} standard about authentication and \ transport level
communication, which is builds on Public Key Infrastructure (PKI), an
architecture based on X.509 public key certificate. GSI implements some
enhancement (such as certificate delegation) based on standard SSL/TLS
protocol. Mutual authentication is required by GSI, and is the default
configuration for GSI based grid deployment; and X.509 certificates is
required for both of the client and service sides in order to achieve
mutual authentication. X.509 certificates is issued by trusted third
party called certificate authorities (CA), and then CAs constitute
trust federation and guarantee two different X.509 certificates from
different CAs can accomplish authentication to each other. So if a user
would access grid system, he/she should own a X.509 certificate which
is issued by a CA that is trusted by other{\textquotesingle}s entity in
the GSI based grid system.

In terms of the virtual organization issues mentioned earlier, GSI and
some other related solutions like virtual organization management
service (VOMS) [ref3] has partly solved the them in some sense. The
identity delegation and X.509 proxy certificate initiated by GSI can
achieve single sign-on, so that the need to re-enter the
user{\textquotesingle}s pass phrase for authentication can be avoided
by creating the proxy certificate, even if some local or remote
entities access access resources on behalf of this user. \ As the most
successful grid authorization model, VOMS adopts X.509 Attribute
Certificate[ref4]. Attribute certificate is signed by VOMS server and
then binded to proxy certificates by client, \ and attributes
certificate includes the VO memberships which is associated with the
proxy certificate{\textquotesingle}s identity. Essentially the
resources can make access control decision based on parsing the VO
membership attributes from the homogeneous attribute certificate.

\ \ \ \ The VO is essentially supposed to be able to provide resource
sharing capability with resources running under heterogeneous
systems/middlewares. This goal can be achieved using Web Service
Architecture (WSA). Therefore, the Web Service technology has been
converged into Grid computing middlewares. The Web Service technology
has been adopted by Globus toolkit, gLite, as well as the ARC.
Actually, utilizing Web Service technology and providing seamless
interoperability with other middlewares is one of the main goals of the
new version of ARC middleware.

Unfortunately, the existing grid security solutions such as GSI and VOMS
do not apply to Web Service \ based application very well due to the
following reasons:

\liststyleLii
\begin{itemize}
\item Firstly, the common implementation for Web Service does not
necessarily recognize and then properly verify and parse the
certificate, since comparing with usual X.509 certificate, proxy
certificate needs different treatment such as proxy path verification. 
\item Secondly, the X.509 Attribute Certificate has not been adopted by
any Web Service implementation at all. 
\item Thirdly, what is even worse is that the GSS API based confidential
communication has also not been adopted by Web Service implementation.
\item Finally, identity delegation is a good solution for user or
user{\textquotesingle}s representative (on behalf of the user) to
move/travel from one resource to another resource. But since the
identity delegation has been completely coupled with the GSI
confidential communication process, it is not possible for the existing
identity delegation solution to be used by Web Service applications.
\end{itemize}
There is some effort to add GSI support to adapt to the strict security
requirement including GSI from grid middlewares, such as GSI plug-in
for gSOAP[ref6]. However, this kind of solution
doesn{\textquotesingle}t help the existing grid middlewares
interoperate with other common Web Service implementation. Instead, Web
Service implementations commonly use the WS-Security, Security
Assertion Markup Languages (SAML), and standard SSL/TLS as well as
standard X.509 certificate when considering security. \

\ \ \ \ Moreover, the current grid security solutions require each user
to possess a X.509 certificate. But the process to get a certificate
may take quite long time, since the requester needs to be checked and
approved by the Registration Authority (RA) and then the Certificate
Authority (CA) can issue the certificate according to the approval from
RA. Meanwhile the process to use the certificate is also not so easy
for the wider less-IT focused research community to deal with, since
the user need to generate the proxy certificate by using the grid
specific command lines, such as \textit{grid-proxy-init}, and
\textit{voms-proxy-init}. On the other hand, users could most normally
have some local community/institutional credentials such as
username/passwords to which users are more familiar. To enable users
use their own community credential instead of the X.509 certificate to
access grid is a promising solution to make the grid be more easily
accessible.

\ \ \ \ The above analysis shows that in order to achieve
interoperability and accessibility for user to access grid system, we
need to address the mismatch between the different security solutions
from different middlewares while keeping the changes as minimum as
possible and the benefits as much as possible; also need to address the
mismatch between user{\textquotesingle}s normal experience and
middleware requirement in order to ensure the access as simple as
possible.

\ \ The approach in the new version of ARC is to utilize the existing
Web Service standards, such as SAML and WS-Security, to achieve the
interoperability and accessibility without breaking existing grid
specific security protocols.

\ The rest of this paper is organized as follows: Section 2 shows the
ARC grid middleware, especially the architecture if the new version of
ARC. Section 3 describes the solution about cross-domain authentication
and single sign-on is presented, including a few solutions implemented
for bridging the mismatch in terms of authentication. Section 4
discusses the related research, and Section 5 concludes.

\bigskip

{\centering
2. ARC grid middleware 
\par}

ARC(Advanced Resource Connector) is an open source grid middleware
solution released under GPL licence. ARC middleware initially aimed at
developing a grid middleware which provides characteristics such as
self-organized, fault-tolerant, non-intrusive, easy-management [fgcs
ARC paper]. \ Now \ the classic version of ARC provides grid services
such as grid job submission and management, resource characterization,
resource aggregation and discovery, basic data management, integration
of grid security solution, etc. Classic ARC has been deployed and used
in production environment, and been one of the mainly deployed grid
middlewares in Europe.

The new generation of ARC is funded by KnowARC project, and based on the
functionality and capabilities of the classic ARC middleware, it aims
at implementing a service-oriented middleware which will provide higher
levels of resource and user abstracting through well-defined web
service interface [Design document] in order to provide
interoperability with other service-based grid middlewares, as well as
other Web Service compatible applications.

As the key part of the implementation of new ARC middleware, there is a
lightweight web service container called HED (Hosting Environment
Daemon) which provides a host place for various services in application
level, as well as a host place for a bunch of components to support
flexible, interoperatible, and efficient communication mechanism. The
whole design of the HED is built around the idea of flexibility and
modularity, which means for the developer, he can easily concentrate on
the application level Web Service implementation by only using the core
minimum amount of components if he only would do application level
development, or he can easily concentrate on the middleware level
implementation such as supporting another communication protocol or
implementing authentication mechanism by using the core minimum amount
of components and external dependencies; also for the deployer, he can
easily configure and deploy the middleware and application for
different kinds of requirements without being bothered to know much
about the implementation.

{\centering
Figure 1. The architecture of Host Environment Daemon
\par}

\begin{center}
\includegraphics[width=4.0673in,height=3.1047in]{Secpaper-img1.png}
\end{center}
The architecture of the Host Environment Daemon is illustrated as Figure
1. In general, there are a few \ components called MCC (Message Chain
Components) which are in charge of processing different messages for
different protocol levels. For instance, as shown in the example
message flow, HTTP MCC will process secure socket stream from TLS MCC
to generate output HTTP message to SOAP MCC, and also process SOAP
documentation from SOAP MCC to generate HTTP message to TLS MCC.

Two examples of MCC components configuration in Figure 1 shows MCC are
configurable: real line shows SOAP over HTTPS (HTTP over TLS), while
dotted line shows SOAP directly over TCP. Service administrator can
configure the MCCs according to his requirement, such as the
interoperability requirement with the other part. For instance, the
later configuration is compatible to WSE (Web Services Enhancement for
.NET) {\textquotesingle}s SOAP message mechanism (see
WSE{\textquotesingle}s \textit{SoapSender} and \textit{SoapReceiver}).
Another configuration could be SOAP over HTTPG (HTTP over GSI) which
can be used to interoperate with some service based on HTTPG such as
SRM (Storage Resource Manager) [ref] service. The easily configurable
characteristics shows the flexibility of HED in terms of protocol
supporting.

Because in the current state HED mostly provides modules for building
SOAP based Web Services, it is easy to think that HED is just another
Web Services development framework like Axis, gSOAP, XFire or any other
out of the numerous implementations. Instead, the idea of HED is to
provide framework for gluing functionalities and not a
re-implementation of various standards. Effectively that means if
Apache httpd web server is considered by developers as necessary for
serving as front-end to services there could be plugin code written
which places Apache2 into a chain of other plugins of the HED, the same
for the Apache Axis SOAP implementation, as shown in Figure 1.

Even though the external solutions like Axis and Apache httpd can be
used for protocol supporting, \ two protocols are completely developed
in HED to support HTTP and SOAP in order to make the solution
lightweight and needed an implementation of those supported protocols
that is both simple and lightweight. However external solutions are
used whenever is appropriate - as in the case of TLS (openssl is used),
GridFTP (GridFTP API is used), LDAP and some other cases, this
principle applies to Axis and Apache httpd if we used the API from
those solution to implement the plugins for supporting SOAP and HTTP
protocol.

HED contains a framework for implementing and enforcing authentication
and authorization. Each Message Chain Component (MCC) or service has a
common interface for implementing various authentication and
authorization functionality. This functionality is implemented by using
pluggable and configurable components (plug-ins) called
\textit{SecHandler}. The \textit{SecHandler} components are C++ classes
and provide method for processing messages traveling through Message
Chains of the HED. Each MCC or Service \ usually implement two queues
of SecHandlers -- one for incoming messages and one for outgoing called
{\textquotedblleft}incoming{\textquotedblright} and
{\textquotedblleft}outgoing{\textquotedblright} respectively. All
\textit{SecHandler} components attached to the queue are executed
sequentially. If any of them fails, message processing fails as well.
In figure 1, the {\textquotedblleft}AuthZ{\textquotedblright} and
{\textquotedblleft}AuthN{\textquotedblright} sub-module inside MCCs and
Services are examples of \textit{SecHandler}.

As explained above, HED is supposed to act as a service holder, but on
the client side, from implementation perspective, there is similar
architecture implemented for processing messages from different
protocols and handling security functionality, also there is a specific
application programming interface (API) for developers to easily write
Web Service client code to service.

\bigskip

{\centering
3. Cross-domain authentication and Single Sign-On
\par}

For cross domain authentication, our work comprises the following five
major aspects:

\liststyleLiii
\begin{itemize}
\item \textit{Integration of community authentication} -{}-{}- to
utilize the standardized community authentication from Shibboleth
Identity Provider software when users authenticate against Grid
services.
\item \textit{Short lived credential service} \ {}-{}-{}- to issue the
short lived X.509 certificate for user by using the community
authentication, in order to interoperate with services that depends on
PKI based mutual authentication.
\item \textit{Credential delegation service} -{}-{}- to process the
X.509 credential delegation as a normal Web Service, in order to
provide single sign-on characteristic.
\item \textit{Message level authentication} -{}-{}- to provide message
level authentication by implementing WS-Security.
\item 
\bigskip
\end{itemize}

\bigskip

\ \ 3.1 Community based authentication for Grid by utilizing Shibboleth
Identity Provider

AAI (Authentication and authorization Infrastructure) is a solution for
the authentication and authorization for inter-organization resource
sharing, such as electronic resource sharing between libraries, etc.
AAI implicitly applies to community/institutional/organization based
authentication where users are from different home communities but need
to get resources from other communities by using some federation
mechanism. Unlike the X.509 based authentication solution in current
grid systems, AAI does not require user to provide X.509 certificate,
instead, it can support different types of authentication, such as
username/password based authentication, IP address based
authentication, etc. There are a few implementations about AAI, among
which Shibboleth [ref] is one implementation which has been widely
deployed.

Shibboleth provides cross-domain single sign-on and attribute-based
authorization while preserving user privacy. Shibboleth is based on the
OASIS Security Assertion Markup Language (SAML), especially the \ new
version of Shibboleth supports SAML 2.0 specification. For
authentication, the main SAML profile that Shibboleth implements is the
SAML2.0 web browser SSO profile, which \ define two functional
components, an Identity Provider and a Service Provider. The Identity
Provider (IdP) is responsible for creating, maintaining, and managing
user identity, while the Service Provider (SP) is responsible for
controlling access to services and resources by using the SAML
assertion produced and issued by IdP upon request. In order to discover
which home community does a user from, Shibboleth specifies an optional
third component called {\textquotedblleft}Where Are You
From?{\textquotedblright} (WAYF) service to aid in the process of IdP
discovery, and this IdP discovery process is also standardized and
\ defined in SAML 2.0 specification and called as
{\textquotedblleft}Identity Profile Discovery
Profile{\textquotedblright}.

We utilized the SAML2.0 web browser SSO profile for the authentication
in ARC middleware. But since the SSO profile is initially supposed to
protect web applications and provide authentication for web users, we
implemented some external code on the client and service side to
integrate SSO profile. On the client side, apart from the WS-Client API
for writing Web Service client, we implemented the user agent
functionality of web browser in order to mimic the behavior of web
browser{\textquotesingle}s user agent, such as http redirection, cookie
processing, etc. In fact, the implementation of user agent is also
based on the client interface of ARC, specifically, the https client
interface, since the client interface of ARC can support different
protocols which are incarnated by different message chain component
(MCC). Hence client developers who would use SAML2.0 SSO profile should
call the user agent interface and then the WS-Client interface. On the
service side, we implemented the Service Provider functionality (based
on the HTTP MCC configured together with TLS MCC) which is called SP
Service. For Identity Provider, the Shibboleth IdP implementation is
used.

Figure 2 shows the process of SAML2.0 SSO integrated in ARC client and
service.

\begin{center}
\includegraphics[width=3.4063in,height=2.1417in]{Secpaper-img2.png}
\end{center}

\bigskip

\bigskip

{\centering
Figure 2. SAML2.0 SSO profile in ARC
\par}

The steps shows in Figure 2 are described as follows:

\liststyleLiv
\begin{enumerate}
\item The client use the user agent interface to launches a http request
including the IdP name (to which the user belongs) to the service side.
The host-name and port of destination URL is the same as that of the
real service, the last part of the destination URL is
{\textquotedblleft}saml2sp{\textquotedblright} which is specific for
pointing to the Service Provider. Note that we use Identity Provider
(IdP) name here to simplify the IdP discovery process, because we
suppose that user who would access the service should better know where
is he from initially.
\item The SP Service (Service Provider) searches the metadata (we use
the same metadata with the same format defined in Shibboleth) and gets
the location of the single sign-on service (hosted in IdP) and also the
location of assertion consuming service (hosted in this SP itself) in
order to compose the SAML {\textless}AuthnRequest{\textgreater}
message. Then SP Service issues this
{\textless}AuthnRequest{\textgreater} message by using its own X.509
certificate (Note in the SAML SSO profile, certificate is still needed
by IdP and SP) and sends back to user agent.
\item User agent sends the {\textless}AuthnRequest{\textgreater} message
to Identity Provider.
\item Identity Provider requires an act of authentication. The
authentication mechanism is outside of the SAML2.0 SSO profile. The
shibboleth IdP implementation will choose some login handler for
authentication. The current user agent implementation is compatible to
the Username/Password login handler of Shibboleth IdP. Through the http
protocol, user agent will feed IdP with the username/password which has
been given by the caller of user agent interface.
\item Once the authentication has been succeeded, the IdP issues a SAML
response including an encrypted (encrypted by destination
SP{\textquotesingle}s public key) SAML assertion, and then this SAML
response will be delivered by the user agent to the Service Provider. 
\item The SP Service verifies and checks the SAML response, and decrypts
and stores the SAML assertion into session/connection context. The SAML
assertion actually includes the {\textless}AuthnStatement{\textgreater}
and {\textless}AttributeStatement{\textgreater}.
\item The WS-Client launches the real Grid/Web Service request via the
same connection as the one which is used by user agent to contacts SP
Service.
\item The Grid/Web Service checks the
{\textless}AuthnStatement{\textgreater} from the session context to see
if the session is still valid through the \textit{SecHandler} called
{\textquotedblleft}SAML SecHandler{\textquotedblright}. And if valid,
service handles the real service processing and returns the response to
WS client. Note that service requires WS client is from the same
connection as the one on which user agent contact SP service in order
to guarantee that the validity of SSO profile result effects the real
client/service interaction.
\end{enumerate}
The SP service and real functional service(s) are hosted by the same
container, and they use the same X.509 credential. The client side
authentication is switched off, so that client doesn{\textquotesingle}t
need to use any X.509 credential. Only the trusted certificates (CA
certificate for SP and IdP) need to be configured for client side so
that SP and IdP can authenticate themselves to the client. As required
by SAML2.0 profile, the SP and IdP should have trust relationship to
each other.

One benefit of SAML2.0 SSO profile that is worth mentioning is that: the
Identity Provider could cache the authentication result through session
management once the user agent has succeeded to authenticate, then in a
short period this authentication result is valid so that user agent
don{\textquotesingle}t need to feed IdP with user{\textquotesingle}s
username and password the next time (if the time is not out of the
scope of valid period) it authenticates against IdP. So user (or the
client on behalf of this user) can travel across multiple security
domains with only inputing his name and password once, which is the
characteristic of single sign-on.

Since the Shibboleth implementation of SAML is standard-compliant and
widely deployed, the solution implemented in ARC can easily
interoperate with other SAML implementations with minimum changes, and
more importantly, this solution can succeed to utilize the widely
deployed SAML implementation for authentication in grid systems by
avoiding the usage of X.509 certificate.

Moreover, even though the implementation is based on ARC middlerware,
the idea can be adopted by other grid middlewares if they can only
require server SSL authentication instead of mutual SSL authentication.

\bigskip

\ \ 3.2 Short lived credential service

However, most of the widely used grid middlewares are based on GSI while
GSI requires mutual authentication. Also for Web Service based grid
solution, we can not prevent service side from requiring client X.509
certificate. \ Based on the solution described in Section 3.1, we
implemented a short lived credential service (SLCS) by which user can
get a short-lived X.509 certificate without being bothered to contact
any registration authority (RA) or certificate authority (CA).

The SLCS service is actually a specific Web Service (standard web
service implemented by using ARC service interface) in Figure 2, and
the SLCS client is a specific ARC client which includes the user agent
and WS client. The whole process from step 1 to step 8 is the same as
in Figure 2.

\bigskip

\begin{center}
\includegraphics[width=3.7689in,height=3.2055in]{Secpaper-img3.png}
\end{center}
{\centering
Figure 3. Short lived credential service
\par}

\ \ \ SLCS client generates a X.509 certificate request, launches a Web
Service request which includes the certificate request; SLCS service
then gets the certificate request, composes a distinguished name (DN),
issues a certificate (short lived, 12 hours by default) with the SAML
attribute (from the SAML2.0 SSO profile) as the X.509 certificate
extension, and puts the certificate in to the Web Service response;
SLCS client get the response and stores the X.509 certificate into
local repository.

The command line interface for the SLCS client is like this:

\$ ./arcslcs -S https://127.0.0.1:60000/slcs -I
https://idp.testshib.org/idp/shibboleth -U myself -P myself

\ \ \ Since the lifetime of the short lived credential is normally
short, it is not a must to protect the private key by pass-phrase. As
illustrated in step a and step b in Figure 3, if the private key is not
protected, through Web Service client, the user can use the X.509
certificate to access Grid Service or Web Service from any kind of
middleware. If the private key is protected, the can use the X.509
certificate to generate a proxy certificate (by using command line
interface utility such as \textit{grid-proxy-init},
\textit{voms-proxy-init}, or \textit{arcproxy}), and the use the proxy
certificate to access Grid Service or Web Service.

\ \ \ It is worth mentioning that since ARC middleware can support GSI
communication by configuring the GSI MCC, \ together with the X.509
certificate, the Web Service client developed by ARC client interface
can interoperate with Grid Service that requires GSI communication.

\ \ \ It might be noticed that how to compose the distinguished name
(DN) for the certificate is critical issue for SLCS service. Since the
Shibboleth Identity Provider uses the eduPerson schema [ref] for the
definition of {\textless}saml:Attribute{\textgreater} in
{\textless}saml:AttributeStatement{\textgreater}, we pick the
relatively distinguishable attribute
{\textquotedblleft}eduPersonPrincipalName{\textquotedblright} for the
DN. A typical \ eduPersonPrincipalName \ value could be
\href{mailto:alice@example.org}{alice@example.org}, then the DN is
{\textquotedblleft}/O=knowarc/OU=example.org/CN=alice{\textquotedblright}.

The obvious benefit of SLCS service is that: If a user passes the
authentication to his home Identity Provider, he can get the X.509
credential at anywhere only by running the SLCS client command together
with inputing his username and password to this home IdP, and then
access the grid system conveniently; meanwhile, thanks to the single
sign-on characteristic discussed in Section 3.1, this user
\ doesn{\textquotesingle}t not need to input his username and password
in a valid period after the first time he succeeds to authenticate
against his home IdP by running the SLCS client command on the same
node, even if this SLCS client command is supposed to run against a few
SLCS services to get a few SLCS credentials.

\bigskip

\ \ 3.3 X.509 Credential delegation service

The delegation of credentials proposed in Grid Security Infrastructure
(GSI) is a good solution for the supporting of single sign-on for
users, except the coupling of the GSI communication which is not so
widely accepted in Web Service applications. In the design of the new
version ARC middleware, we keep the credential delegation idea, while
utilizing the SOAP communication (and normal TCP or TLS for transport
level communication) instead of the GSI communication for the
credential delegation process.

As shown in Figure 4, there is a specific WS client and Web Service as
well for processing delegation: delegation client, and delegation
service. The delegation service has two main wsdl operations, one for
processing the delegation initiation; the other for processing/storing
the signed proxy certificate. The delegation process is detailed in
Figure 5.

For the service invocation process, the WS client functionality (inside
each service or the initial client) which is included to access the
target service (step a1 or a2 in Figure 4), should use the
user{\textquotesingle}s X.509 certificate or proxy certificate (on
behalf of the user) for authentication and secure communication. For
the delegation process, the principle is the same.

Since each WS client accesses the target service (delegation service or
other functional service) on behalf of the user, the trust relationship
between the user{\textquotesingle}s certificates (X.509 certificate and
proxy certificate) and services is required, while the trust
relationship between service and service is not required.

\begin{center}
\includegraphics[width=3.802in,height=3.0689in]{Secpaper-img4.png}
\end{center}

\bigskip

\bigskip

{\centering
Figure 4. Credential delegation service
\par}

\bigskip

\begin{center}
\includegraphics[width=2.5319in,height=1.8571in]{Secpaper-img5.png}
\end{center}
{\centering
Figure 5. Credential delegation process
\par}

\bigskip

\ \ 3.4 Message level authentication

Transport level security (TLS or SSL) is not sufficient for securing Web
Services, because the secure channel created by SSL/TLS make it
impossible to differentially protect the SOAP message, e.g. \ to
encrypt and/or sign only particular components of the SOAP message,
which is relevant when non-sensitive portions of the message need to be
accessed or changed by intermediate actors. Additionally, SSL/TLS is
incapable of providing end-to-end protection for SOAP message which
might flow through multiple intermediate actors, because it only allows
each hop to be protected with the resultant security gaps at
intermediate actors. Web Services Security (WS-Security) specification
(ref) provides means for applying security to Web Service.

The WS-Security related specifications are implemented in ARC as
\textit{SecHandler}. There are three kinds of WS-Security version 1.1
specifications are completely or partly implemented: Username Token
profile, X.509 Token profile, and SAML Token profile.

For those WS clients on behalf of a user, at any step of service
invocation, the WS client can augment the SOAP message with WS-Security
token by simply configuring the WS-Security \textit{SecHandler} into
container{\textquotesingle}s configuration, if the relying party
(target service) requires.

If the WS client owns an X.509 certificate or proxy certificate, the
process of generating an X.509 Token for a SOAP message is: the client
puts the content of certificate into X.509 token as
{\textless}wsse:BinarySecurityToken{\textgreater}, and uses the private
key (corresponding to the certificate) to create XML signature for the
SOAP message body.

However the process of generating a SAML Token is more complicated. In
SAML Token profile, \textit{ConfirmationMethod} is the verifying method
for establishing proof-of-possession for the subject. Two methods are
proposed in SAML Token profile: \textit{holder-of-key},
\textit{sender-vouches}. \ We rely on the \textit{holder-of-key
}method. For the \textit{hold-of-key} subject confirmation method,
there are three interaction parts: the attesting entity, the relying
party and the issuing authority. Firstly the attesting entity
authenticates to issuing authority by using some authentication scheme
such as X.509 Token profile. So then issuing authority is able to make
a definitive statement (sign a SAML assertion) about an act of
authentication that has already taken place. The attesting entity gets
the SAML assertion and then signs the SOAP message together with the
assertion by using its private key (the relevant certificate has been
authenticated by issuing authority, and its relevant public key has
been put into \textit{SubjectConfirmation}

element under SAML assertion by issuing authority. Only the actual owner
of the SAML assertion can do this, as only the subject possesses the
private key paired with the public key in the assertion. This
establishes an irrefutable connection between the author of the SOAP
message and the assertion describing an authentication event.)

\ The relying party is supposed to trust the issuing authority. When it
receives a message from the asserting entity, it will check the SAML
assertion based on its predetermined trust relationship with the SAML
issuing authority, and check the signature of the soap message based on
the public key in the SAML assertion without directly trust
relationship with attesting entity (subject owner).

\bigskip

\ \ 3.5 SAML Token delegation service

\ There is also delegation requirement on the message level, some
entities (client or user) often need to temporarily delegate itself to
another entity to perform action on their behalf. We propose a
delegation mechanism for SAML Token.

The delegation process is similar to the process for delegation proxy
certificate, with the following differences:

\liststyleLv
\begin{enumerate}
\item The communication between delegation client and service is based
on mutual authentication between proxy certificate and service
certificate, instead, it is based on mutual authentication between
service certificate and service certificate.
\item The delegation service will not generate any certificate request,
instead, it picks the public key from its own service certificate, and
sends back to delegation client.
\end{enumerate}
Figure 6 shows the token change due to delegation. On the left side, B
is the identity of the attesting entity{\textquotesingle}s X509 token
mentioned above, while A is the identity of issuing authority, and the
SOAP message is signed by B{\textquotesingle}s private key. On the
right side, C is the identity of the delegation service, ans now the
SOAP message is signed by C{\textquotesingle}s private key.

\bigskip

\begin{center}
\includegraphics[width=4.2571in,height=2.152in]{Secpaper-img6.png}
\end{center}
{\centering
\ \ \ \ \ \ \ \ \ Figure 6. Token before and after delegation
\par}

\bigskip

\bigskip

\bigskip

\bigskip

{\centering
4. Related work
\par}

Shibboleth

GridShib

gLite Delegation Service?

SWITCH SLCS Service

the CCGrid 06 paper: Shibboleth-Protected privilege management
infrastructure

Delegation service : gridsite

\bigskip

\bigskip

{\centering
5. Conclusion and Future work
\par}

\bigskip

\bigskip

Reference:

Foster I, Kesselman C, Tuecke S. The anatomy of the Grid: Enabling
scalable virtual organizations. International Journal of Supercomputer
Applications 2001; 15(3):200--222.

\bigskip

The Shibboleth Project.
\href{http://shibboleth.internet2.edu/}{Http://shibboleth.internet2.edu/}

\bigskip

eduPerson and eduOrg Object shema.
\url{http://middleware.internet2.edu/eduperson/}

\bigskip
\end{document}