### 2009-02-04 doc/client/resubmit_and_migrate_plans.tex
implementations plans for resubmission and migration

\documentclass[12pt,oneside,a4paper,english]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{graphicx,epic,eepic}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Thoughts on resubmission within ARC1}
\author{Peter Rosendahl}

\begin{document}
%\maketitle
\section*{Implementation plan for the resubmission and migration functionality in ARC1}
Before discussing the implementation of the resubmission and migration functionality in ARC1 we first make a distinction between migrating, resubmitting and resuming a job.

{\bf Resubmission} of a job means that the exact same job is redone from scratch, i.e. no changes in the job description or input files. If the job description or input files has changed the job no longer remains the same job and the user should submit a new job instead of resubmitting the previous job.

{\bf Resuming} a job means restarting the job from its last resumable state. The last resumable state of a job is reported by the cluster running the job, thus this action will be dependent on the cluster's middleware. We will only talk about resuming a job if the job has already started running otherwise there will be nothing to resume.

{\bf Migrating} a job is defined as moving a job from a cluster(queue) to a new cluster(queue) where it will then be executed.

\subsection*{Solutions}
In order to solve the scenarios mentioned above, we suggest to implement functionality on both server and client side to provide the commands \texttt{arcresume}, \texttt{arcmigrate} and \texttt{arcresub}.

\subsection*{\texttt{arcresume}}
The {\texttt arcresume} command will restart the job from its last resumable state (as reported by the cluster) on the same cluster (computing element) as the job initially was running on. A typical use of this functionality could be if a job has run successfully, but fails in the uploading of output files, e.g. due to the expiration of the user proxy. The user can resolve this situation by uploading a new proxy to the cluster and invoking the {\texttt arcresume} command which resumes the job and the file upload. Since the job will resume at the same cluster (computing element) the {\texttt arcresume} command should be possible to implement for all supported middlewares if the middleware itself supports the resume action. {\bf This command is neither a migration nor a resubmission of the job, and thus not foreseen to be a part of the D2.2-1 deliverable, but will be implemented as a part of Task 1.2.}

\subsection*{\texttt{arcmigrate}}
The \texttt{arcmigrate} command should enable users to move jobs from one cluster to another. In this first implementation the command will only support the migration of jobs which are queued in the batch system, i.e. there will be no support for check-pointing etc allowing for migration of running jobs. Nevertheless migration of queueing jobs is the foreseen most frequent use case, in which users would like to migrate queueing jobs from one cluster (computing element) to another cluster (computing element) which meanwhile have been freed or made available in other ways.

Upon migration of the job, both the job description and the local input files will be copied directly from the cluster where the job is queueing to the new cluster where the job will be executed\footnote{Input files from storage elements will be copied from the storage element onto the new cluster}. No files or job description will pass via the user's host, thus enabling the user to migrate a job even if the local input files has been changed or deleted. Additionally this will enable the user to perform job migration from a host different from the submission host, e.g. migrating jobs from an internet cafe.

{\bf Keep in mind:} Upon job migration the initial job is killed and hence the job receives a new job-id. The old job-id and accompanying job information (endpoint etc) thus have to be updated in the user's local XML file which holds the list of all active jobs.

\subsubsection*{Technical imlplementation:}

\subsection*{\texttt{arcresub}}
The resubmission command will redo an entire job from scratch by first trying to kill the existing job (if possible) and then resubmit the job. The command could e.g. be used if the contact to the cluster running a job is lost, and the user would like to resubmit the job to a different cluster. In this particular use case no information about the job can be retrieved from the cluster, thus everything constituting the job (job description and input files) will have to be retrieved from elsewhere. This is solved by implementing a local sandbox which upon job submission automatically stores the job description and checksums of the local input files. The local sandbox ensures that the user is able to resubmit the job to a new cluster even if the cluster running the job goes down, and that the resubmitted job contains the exact same input files as the original job (reliable resubmission)\footnote{The reliable resubmission only governs the local input files. Input files taken from storage elements are not under this control}.

The feature of storing job descriptions and checksums in a local sandbox could in the future be expanded to storing these in a ``grid sandbox'', e.g. on a grid storage element. The ``grid sandbox'' could act as a remote joblist which would give an alternative way for the \texttt{arcsync} command to syncronise the local joblist. If the grid sandbox even stores local input files upon submission, it could additionally be used to allow the user to resubmit the job from a different host then the original submission host.

{\bf Comparison with ngresub:} The existing \texttt{ngresub} command relies on retrieving the job description from the cluster before resubmitting the job. This prevents the command from being used in the most likely usage scenario in which the original cluster is down. Additionally the \texttt{ngresub} command does not check that local input files are the same, i.e. it can not guarantee that the new job is exactly the same as the old one.

Both the issues above of the \texttt{ngresub} command are resolved with the proposed new implementation (\texttt{arcresub}). The proposed new solution also resolves the situation with multiple middlewares supported by the client library, and enables resubmission to middlewares different from the middleware originally running the job. All this is achieved through the usage of a local sandbox, which on the other side prevents the user from doing resubmission from a different host then the submission host since everything is stored locally. This could in the future be solved by adopting the above mentioned model of a ``grid sandbox'' but local input files are still likely to be a cause of concern.

\subsubsection*{Technical imlplementation:}

% Comments from Balazs:
%relate new job to the old one
%clarify what it means to be started
\end{document}