### 2009-02-19 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Tutorial on programming web services with ARC-1. This some interim state, uploaded to help our Luebeck-internal review as a start. An announcement to the list will follow next week.

\chapter{SEC echoservice}

\section{SecureHandler}
And what are PDP's?

Currently, the ARC1 comes in severalfour policy decision point implementations:
\begin{itemize}
 \item arc.pdp – Arc PDP\\
The Arc PDP will organize the security attributes into the ARC specific authorization request, call the
policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository,
and get back the evaluation result. See paragraph 3 for detail information about request schema and policy
schema.
\item delegtion.pdp – Delegation PDP\\
The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from
remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See
section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
\item simplelist.pdp – Simplelist PDP\\
The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity
extracted from the remote credential (or proxy credential) with local list of permitted identities.
\item pdpservice.invoker – PDP Service Invoker\\
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the
same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP
message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation
engine can be accessed remotely and maintained centrally.
 \item allow.pdp\\
 Does something unknown!
\end{itemize}

Following Flavors:
\subsection{arc.authz – Authorization SecHandler}
The arc.authz is responsible for calling the interface of policy decision point and getting back the
authorization result, and then making decision according to this authorization result. There is one simple
interface (see Figure 4) defined in PDP, which will be called by arc.authz if configured inside once there
is message (incoming or outgoing) need to be processed.

\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:secHandler_arc.authz,
language=XML,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
...
<Service name="sececho" id="sececho">
	<next id="sececho"/>
	<echo:prefix>[ </echo:prefix>
	<echo:suffix> ]</echo:suffix>
	<SecHandler name="arc.authz" id="pdps" event="incoming">
		<PDP name="simplelist.pdp" location="/etc/grid-security/grid-mapfile"/>
	</SecHandler>
</SecHandler>
...
\end{lstlisting}
\end{minipage}

\begin{lstlisting}[
label=lst:invokation,
language=ksh,
frame=single,
numbers=none,
backgroundcolor=\color{lightgray},
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ tail -n5 /etc/grid-security/grid-mapfile
"/C=Au/O=Science/OU=ARC/CN=Baldur Meier" griduser
"/C=UK/O=MoreScience/OU=ARC/CN=Thor Schulz" griduser
"/C=Au/O=Organelle/OU=ARC/CN=Loki Svensson" griduser
"/C=SK/O=Grid.Org/OU=ARC/CN=Iduna Apple" griduser
"/C=RU/O=MegaGrid/O=ARC/CN=Fenrir Wodaneater" griduser
\end{lstlisting}
Add your client certificate into that list!

Latest: Policy.xsd
http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd

\subsection{identity.map – Identity Mapping SecHandler}

The identity.map is a specific authorization oriented security handler. It will map the global identity in
the message into local identity like system username based on the result returned by Policy Decision Point
components.

\subsection{delegation.collector – Delegation SecHandler}

The delegation.collector is responsible for collecting the delegation policy information from the remote
proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy
into message’s security attribute for the usage of other components, such as delegation.pdp.

\subsection{ usernametoken.handler – UsernameToken SecHandler}

The task of the usernametoken.handler is to generate the WS-Security Username-Token and add it into
header of SOAP message which is the payload of outgoing message. It can also extract the WS-Security
Username-Token from the header of SOAP message which is the payload of incoming message.

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/secechoclient/secechoclient.cpp}

\lstsetCPP

\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/secechoservice.cpp}

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/arched_sec_echoservice.xml}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
	<soap-env:Body>
		<echoRequest:echo>
			<echo:say>Der_Berg_ruft</echo:say>
		</echoRequest:echo>
	</soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <echo:echoResponse>
      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
    </echo:echoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}