### 2009-02-23 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex

 
\chapter{Persistent Webservice}

export ARC\_BARTENDER\_URL