### 2009-02-24 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Small modifications...

 
\chapter{Persistent Webservice}

export ARC\_BARTENDER\_URL

    key\_file = os.environ.get('ARC\_KEY\_FILE', None)
    cert\_file = os.environ.get('ARC\_CERT\_FILE', None)
    proxy\_file = os.environ.get('ARC\_PROXY\_FILE', None)
    ca\_file = os.environ.get('ARC\_CA\_FILE', None)
    ca\_dir = os.environ.get('ARC\_CA\_DIR', None)

export ARC\_KEY\_FILE=/local/glodek/clientKey.pem
export ARC\_CA\_FILE=/local/glodek/clientCA.pem
export ARC\_CERT\_FILE=/local/glodek/clientCert.pem

does the daemon runs the service single threaded

race conditions
\begin{itemize}
 \item Transaktionen
 \item nur lesen
 \item jeder prozess schreibt in seine eigene Datei
 \item Gibt es einen     // andere Zugriffe abblocken
    flock (\$fp, LOCK\_EX);  (atomare befehle)
 	wie bei php??
 \item Peterson Algorithmus
 \item Was passiert beim Ausfall (Time out - Transaktion zurück nehmen)
\end{itemize}

\section{Schlusswort}

Other tutorials are:\\
\\
%Oxana reported the status of OGF tutorials:
\begin{itemize}
 \item first tutorial will be given by Oxana and will deal with how to add the 
resources to Nordugrid, .... (will be related mostly to production ARC)
 \item second tutorial will be given by Gabor and he will show the 
interoperability of ARC (video demo will be shown). A-REX - Unicore 
interoperability.
 \item there will be also a tutorial related to ETICS system where the 
possibility of ETICS system to submit a job to ARC and Unicore will be 
presented. The ETICS team has not contacted so far.
\end{itemize}