### 2009-02-27 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex

\chapter{SEC echoservice}

The TLS echo service, presented in the previous chapter enables the adminstrator the control the access to the grid. Due to a certificate owner may be a member of a large VO it is sometimes recommended to have a more fine grained security within the daemon. 
The goal to be reached in this chapter is to enhance the possibilities to limitate the access to particular users and to narrow their rights on the workstation.\\

The security framework of ARC is implemented based on the same concept of modularity as HED. 
To extend the security of a MCC one only has to modify the service configuration file. This is to be done by including a \textit{SecHandler} element into a component or service which again has to contain a \textit{pdp} element. A small example of a chain component is shown in Listing~\ref{lst:sec_service_component_xml}. The \textit{SecHandler} is able to intervene the message flow
by: manipulating the message content, the attributes or just disrupt the message. 
% Attributes done by: identity.map, delegation.collector
% manipulating content: usernametoken.handler
% disrupt by: arc.authz
\lstsetJUSTXML
\begin{lstlisting}[
label=lst:sec_service_component_xml,float=htb,
caption={[Example of a SecHandler within the component \textit{http.service}.]
         \textbf{Example of a SecHandler within the component \textit{http.service}}}]
<Component name="http.service" id="http">
	<next id="soap">POST</next>
	<SecHandler name="arc.authz" id="auth" event="incoming">
		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
	</SecHandler>
</Component>
\end{lstlisting}
Each MCC or Service usually implement two queues of SecHandlers --- one for incoming messages and one for outgoing called
``incoming'' and ``outgoing'' respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well~\cite{QIANG_2008}.
In order to decide which message shall be modified by the \textit{SecHandler} so called PDP (Policy Decision Point) are used. They come in a bunch of flavours such that the variabilty of security which can be realised is big.\\

In this tutorial only a small set of \textit{SecHandler}s and \textit{PDP} will be introduced. For further reading it is recommended to consult the document ``Security framework of ARC1'' written by Weizhong Qiang and Alexandr Konstantinov. The present chapter is closly related to it.\\

  \begin{table}[htb]
  \centering
  \caption{List of SecHandlers}
  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
	\hline
 	\textbf{identity.map}          & Maps the identity of the user to an identity of the workstation.\\
 	\textbf{arc.authz}             & Restricts the access to the next layer.\\
 	\textbf{delegation.collector}  & Collects the information for the delegation policy (Related to \textit{delegate.pdp}). \\
 	\textbf{usernametoken.handler} & Generates the WS-Security username-token and adds it into the SOAP header of outgoing
					 messages. In case of incoming messages the SecHandler extracts the WS-Security username-token out of it. \task{Figure out what the username-token does...}\\
	\hline
  \end{tabular*}
  \end{table}

  \begin{table}[htb]
  \centering
  \caption{List of PDPs}
  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
	\hline
 	\textbf{allow.pdp}           & bla \\
 	\textbf{simplelist.pdp}      & blub.\\
 	\textbf{arc.pdp}             &\\
 	\textbf{delegate.pdp}        & \task{Explain that well}\\
 	\textbf{pdpservice.invoker}  &\\
	\hline
  \end{tabular*}
  \end{table}

In order to realise security within an MCC 
\begin{verbatim}
SecHandler 
==========
	PDP
	===	
		elements/attributes required
		============================
identity.map						--- maps the identity of users to identity within the local workstation
	allow.pdp					--- means concering all identities
							  (I suppose it can be used in every MCC)
		<LocalName>nobody</LocalName>		  (Due to the SecHandler within the PDP the local name has to be spechified)
		???LocalList???
		???LocalSimplePool???
	simplelist.pdp					--- maps only persons listed within the file
								(TLS identity is used here, thus it can be used only after the TLS)
		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
	arc.pdp						--- Complexer policy to permit or deny someone. 
		<PolicyStore>					(Is it possible to add parameters for the SOAP service here?)
		<Location type="file">policy.xml</Location>
		</PolicyStore>
	[...]

arc.authz					--- restricts access to next layer
	arc.pdp					--- using policy.xml
	simplelist.pdp				--- using a file with listed identities
	delegate.pdp				--- like arc.pdp but poicy is parsed from a remote proxy
	pdpservice.invoker			--- Is a client which can be used to invoke the PDP Service which 
							implements the same functionality as Arc PDP, except that the
							evaluation request and response are carried by SOAP
							message. The benefit of implementing PDP Service and PDP Service 
							Invoker is that the policy evaluation engine can be accessed remotely
							and maintained centrally.
							- wasn't that also possible with the delegate.pdp??

delegation.collector 				--- Delegation SecHandler
						--- is responsible for collecting the delegation policy information and 
						putting this policy into message’s security attribute for the 
						usage of other components, such as delegation.pdp
usernametoken.handler				--- The task is to generate the WS-Security Username-Token and add it into
						header of SOAP message which is the payload of outgoing message. 
						It can also extract the WS-Security Username-Token from the header of 
						SOAP message which is the payload of incoming message.
						why should one want to do that???

\end{verbatim}

- have fine grained but complex authorization system (policy evaluator) which uses information stored on the server (arc.pdp) or an an proxy (delegation.pdp) or which is evaluated using a SOAP based service such that the request can be evaluated remotly (pdpservce.invoker). 
- have a simple list which matches identities (simplelist.pdp)
- map users to an identity within the local workstation

have a smaller set of users,
have a policy which may map certain users to certain settings

For HED is build up to  % entsrprechend einer Mod Structur
be modular, the security can easily be extend to our given example.
Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler. The SecHandler components provide a method for processing messages traveling through Message Chains of the HED. 
Each MCC or Service usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called “incoming” and “outgoing” respectively. All SecHandler components attached to the queue are executed sequentially. 
If any of them fails, message processing fails as well. 
Each SecHandler is configured inside the arched configuration file used for configuring whole chain of MCCs 
Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP). PDP can process ARC specific Request and Policy documents which are as well written in XML format~\cite{QIANG_2008}..

\section{SecureHandler}
And what are PDP's?

Following Flavors:
\subsection{arc.authz – Authorization SecHandler}
The arc.authz is responsible for calling the interface of policy decision point and getting back the
authorization result, and then making decision according to this authorization result. There is one simple
interface (see Figure 4) defined in PDP, which will be called by arc.authz if configured inside once there
is message (incoming or outgoing) need to be processed.

\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:secHandler_arc.authz,
language=XML,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
...
<Service name="sececho" id="sececho">
	<next id="sececho"/>
	<echo:prefix>[ </echo:prefix>
	<echo:suffix> ]</echo:suffix>
	<SecHandler name="arc.authz" id="pdps" event="incoming">
		<PDP name="simplelist.pdp" location="/etc/grid-security/grid-mapfile"/>
	</SecHandler>
</SecHandler>
...
\end{lstlisting}
\end{minipage}

\begin{lstlisting}[
label=lst:invokation,
language=ksh,
frame=single,
numbers=none,
backgroundcolor=\color{lightgray},
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ tail -n5 /etc/grid-security/grid-mapfile
"/C=Au/O=Science/OU=ARC/CN=Baldur Meier" griduser
"/C=UK/O=MoreScience/OU=ARC/CN=Thor Schulz" griduser
"/C=Au/O=Organelle/OU=ARC/CN=Loki Svensson" griduser
"/C=SK/O=Grid.Org/OU=ARC/CN=Iduna Apple" griduser
"/C=RU/O=MegaGrid/O=ARC/CN=Fenrir Wodaneater" griduser
\end{lstlisting}
Add your client certificate into that list!

Latest: Policy.xsd
http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd

\subsection{identity.map – Identity Mapping SecHandler}

The identity.map is a specific authorization oriented security handler. It will map the global identity in
the message into local identity like system username based on the result returned by Policy Decision Point
components.

\subsection{delegation.collector – Delegation SecHandler}

The delegation.collector is responsible for collecting the delegation policy information from the remote
proxy credential (proxy certificate is compatible to RFC3820) inside the message, and putting this policy
into message’s security attribute for the usage of other components, such as delegation.pdp.

\subsection{ usernametoken.handler – UsernameToken SecHandler}

The task of the usernametoken.handler is to generate the WS-Security Username-Token and add it into
header of SOAP message which is the payload of outgoing message. It can also extract the WS-Security
Username-Token from the header of SOAP message which is the payload of incoming message.

POLICY STUFF:
copy the request out of the LOGGEr and use that to write your policy.xml
%/C=DE/ST=Germany/L=L\xFCbeck City/O=Instite of Neuro- and Bioinformatics/OU=Mitarbeiter

Deny-Overrides - this is default if no algorithm specified.
...
Administrator of service can configure Authorization SecHandler - arc.authz - for each MCC and Service
and define reasonable and meaningful policy. While defining policy the administrator must take into
account that the attributes defined in the policy should be already collected by previous components in a
chain. For instance, policy with AttributeId “http://www.nordugrid.org/schemas/policy-
arc/types/http/path” should not be configured inside SecHandler attached to MCCTLS.

\lstsetCPP
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/clients/secechoclient/secechoclient.cpp}

\lstsetCPP

\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/secechoservice.cpp}

\lstsetARCHEDXML
\begin{minipage}[t]{\textwidth}
\lstinputlisting
	[
	label=lst:arcecho_arched_xml,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/secechoservice/arched_sec_echoservice.xml}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}
\end{minipage}

\lstsetKSH
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:invokation_arched_timeservice,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
	<soap-env:Body>
		<echoRequest:echo>
			<echo:say>Der_Berg_ruft</echo:say>
		</echoRequest:echo>
	</soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\lstsetJUSTXML
\begin{minipage}[t]{\textwidth}
\begin{lstlisting}[
label=lst:timeservice_cpp_source,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <echo:echoResponse>
      <echo:hear>[ Der_Berg_ruft ]</echo:hear>
    </echo:echoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}
\end{minipage}

\section{Other PDPs}

In order the adminisitrator wants to maintain the grid access using one workstation the delegation.pdp can be used.
The client requests access to a service. In case the request is valid a certificate will be deployed on the desired server such that it can be accessed now \task{Am I right?} \cite{QIANG_2008}\\