### 2009-04-20 doc/usagereporter/aur-tech-draft.tex
Created document for Usage Reporter

%\documentclass{book}
\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-TECH-NN

{\raggedleft NORDUGRID-TECH-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC accounting component}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Technical document}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large P\'eter D\'ob\'e \footnote{dobe@iit.bme.hu} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

%\tableofcontents                          %Comment if use article style
\newpage

\section{Introduction}

\section{Purpose}
  metering resource usage
  record it in a database
    SGAS service used
    standard OGF Usage Records
  possibly billing
  job-level accounting
  user-based accounting
  project name

\section{Architecture}
  (diagram)

  A-REX
  logger
  SGAS service

\section{Operation}

  \subsection{Invocation}
    stand-alone exe
    run by A-REX hourly
    command line arguments
    no config, everything from A-REX

  \subsection{Parsing job log files}
    ``id.randomletters''
    UR generated from them

  \subsection{Accessing LUTS}
    insert method
    certificates from:...

\section{Implementation}
  c++
  sw dependencies
    full arc
  install location

\appendix

\section{Configuration}
  jobreport
  jobreport\_options
  jobreport\_credentials

\section{Usage Record properties}
  input file format
  filled properties
  missing properties

\bibliography{grid}
\end{document}