### 2009-04-21 doc/client/UnicoreSupport.tex
Adding documents on Unicore

\documentclass[11pt,oneside,a4paper,english]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{graphicx,epic,eepic}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{a4wide}
\usepackage{xspace}

\newcommand{\arcmigrate}{\texttt{arcmigrate}\xspace}

\begin{document}
%\maketitle
\section*{Support for Unicore resource discovery and job submission/management}
The next generation ARC client library will offer support for discovering grid resource running Unicore middleware and to submit and manage grid jobs (basic functionality) at these resources. This writeup documents the present status of the Unicore support, what is missing and what is the work plan to implement the missing functionality.
\subsection*{Versions of Unicore supported}
Text here about the supported versions
\subsection*{Status of ongoing work}
\subsubsection*{Resource discovery}
Resource discovery. Can we find all resources? Are there defaults in the config file? Can we get detailed information from the clusters?
\subsubsection*{Job submission and management}
Can we submit jobs? Can we get jobs? Can we kill jobs?

\subsection*{Work to be done}
I guess this section is is to write once the present status is documented. Any depencies? i.e. work that has to be done be others?

\subsection*{Work Plan}
Assuming that the Unicore work can start fulltime again first week of May. How much time would be needed in order to finalize the work listed above in the work to be done section. What should be done first etc.

\end{document}