### 2009-04-28 doc/manuals/user_interface/ui.tex

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {},
  pdftitle = {ARC Clients},
  pdfsubject = {User's manual},
  pdfkeywords = {Grid,KnowARC,ARC,client},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}[1]{\colorbox{shadecolor}{\texttt #1}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\uicommand}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

   \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
   \end{tabular}

   \hrulefill

   {\raggedleft NORDUGRID-MANUAL-?\par}

   {\raggedleft \today\par}

   \vspace*{2cm}

   %%%% ---- The title ----

   \begin{center}

      \textsc{\Large ARC Clients}
      \Large \par \textit{User's Manual}

      \vspace*{2cm}

      % {\centering \large Refers to ARC release series 0.4 and up \large \par}

   \end{center}

   \vspace*{2cm}

   % \abstract{
   %   This is the user's  manual to ARC User Interface.
   % }

\end{titlepage}
% \thispagestyle{empty} $ $
% \newpage
% $\ $
\tableofcontents
\newpage

\chapter{Introduction} \label{sec:intro}

The command line user interface of ARC consists of a set of commands
necessary for job submission and manipulation and data management. A
special utility also exists for test purposes.

\chapter{Commands} \label{sec:cli}

\section{Job submission and management} \label{sec:ui} 
\index{job management}
The following commands are used for job submission and management,
such as status check, results retrieval, cancellation, re-submission
and such. The jobs must be described using a job description language. ARC supports the languages JSDL, xRSL and JDL~\cite{xrsl}.

\input{ui-job}

\section{Data manipulation} \label{sec:dm} 
\index{data management} ARC provides basic data management tools,
which are simple commands for file copy and removal, with eventual use
of data indexing services.

\input{ui-data}

\section{Test suite} \label{sec:testsuite}
%\input{arctest}

\section{Third-party commands} \label{sec:ui-other}
%\input{ui-other}

\chapter{URLs}\label{sec:urls}
\index{URLs}
\input{xrsl-urls}

\chapter{Configuration} \label{sec:client.conf} 
\section{ARC Client Configuration}
%\input{ui-conf}

\bibliography{grid}
\printindex

\end{document}