### 2009-04-28 doc/tech_doc/xrsl/xrsl.tex

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {Oxana Smirnova},
  pdftitle = {Extended Resource Specification Language},
  pdfsubject = {Reference manual},
  pdfkeywords = {Grid,NorduGrid,ARC,Globus,RSL},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-MANUAL-4\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----

    \begin{center}

    \textsc{\Large Extended Resource Specification Language}
     \Large \par \textit{Reference Manual}

    \vspace*{2cm}

%    {\centering \large Refers to ARC release series 0.4 and up \large \par}

    \end{center}

    \vspace*{2cm}

%    \abstract{
%      The paper represents an overview and a reference manual for the
%      NorduGrid Extended RSL, developed on the basis of the \globus\ 
%      RSL.
%    }

  \end{titlepage}
  %\thispagestyle{empty} $ $
  %\newpage
  %$\ $
  \tableofcontents
  \newpage

  \chapter{Introduction}
  \label{sec:intro}

  Execution of jobs at batch clusters is the major part of the data
  analysis, production and other computational tasks and applications
  in many research areas. In the Grid infrastructure~\cite{grid}, such
  clusters are not necessarily local to a job submission site, but can
  be widely distributed. This implies extra requirements for the
  proper description of job options. The \globus\ \index{Globus}
  project~\cite{globus} developed a Grid middleware toolkit, which
  makes use of the \textit{Resource Specification Language
  (RSL)}~\cite{rsl}\index{RSL} to parse job options and definitions to
  resource management systems. The NorduGrid project~\cite{nordugrid}
  \index{NorduGrid} developed the Advanced Resource Connector (ARC),
  -- a 
  solution~\cite{para02,acat02} for a Grid facility, suitable for
  complex tasks, like, for example, a High Energy Physics
  production. To match the complexity of such tasks, this solution
  requires certain extensions to the RSL.

  To describe a task to be submitted to the ARC-enabled resources, an
  extended version of the \globus\  RSL is used. The extensions concern
  not only introduction of new attributes, but also differentiation
  between two levels of the job options specifications:

  \begin{itemize}
  \item[] \textbf{User-side RSL}, i.e., the set of attributes
    specified by a user in a job-specific file. This file is
    interpreted by the \textit{User Interface (UI)}~\cite{ui},
    \index{User Interface} and after the necessary modifications is
    parsed to the \textit{Grid Manager (GM)}~\cite{gm}\index{Grid
    Manager}
  \item[] \textbf{GM-side RSL}, i.e., the set of attributes
    pre-processed by the UI, and ready to be interpreted by the GM
  \end{itemize}

  A user only has to know the user-side part, and utilize it to
  describe the Grid tasks. The Grid Manager, however, uses slightly
  different notations, supplied by the User Interface.

  In what follows, the description of the NorduGrid-extended RSL,
  further denoted as \textbf{xRSL}\index{xRSL}, is given, using the
  following notations:

  \begin{tabular}{ll}
    \verb#<xxxx>#&parameter to be substituted with a corresponding string or a number\\
    \verb#[xxxx]# & optional parameter\\
    \verb#xxx|yyy|zzz# &list of possible values of a parameter\\
    \verb#-"-# &"same as above"\\
  \end{tabular}

  \chapter{xRSL syntax and rules}
  \label{sec:remarks}

  For a complete description of \globus\  RSL, see
  reference~\cite{rsl}. xRSL uses the same syntax conventions,
  although changes the meaning and interpretation of some attributes.

  \section{Syntax}
  \label{sec:syntax}

  A Grid task is described by the mean of xRSL attributes, which can
  be either passed via a command-line, or, more conveniently, be
  collected a so-called xRSL-file (suggested extension
  \textit{.xrsl}). Such a file contains a plain list of attribute-value pairs
  and boolean operands "\&" (for AND) and "\verb#|#" (for
  OR). Attribute names are case-insensitive.

  If the attributes are specified from the command line, the entire
  description must be enclosed either in single or in double quotes.
  Single quotes enclosure is completely analogous to xRSL-file usage,
  while double quotes imply standard shell expansion of the enclosed
  string. This has implications when strings contain local shell
  variables: they \textbf{will not be expanded} unless the task description is
  entered from the command line and is enclosed in double quotes.

  An attribute-value pair is a key element of specification. It
  consists of an expression that assigns one or more values to an
  attribute, and is enclosed in round brackets:

  \begin{shaded}
    \verb#(attribute="value")#\\
    \verb#(attribute="value1" "value2")# (for multi-valued attributes)\\
  \end{shaded}

  Certain attributes do not have assigned value; instead, they contain
  a list of values that should be put in proper correspondence with
  each other:

  \begin{shaded}
    \verb#(attribute=("value1" "value2")("value3" "value4"))#
  \end{shaded}

  In the example above, \verb#value1# and \verb#value3# are put in correspondence to
  \verb#value2# and \verb#value4# respectively, according to the
  context of the attribute.

  Values should be enclosed in quotes if they contain blank
  spaces or special characters. The \textbf{special characters} are:
  \begin{shaded}    
    \verb. +   &   |   (   )   =   <   >   !   "   '   ^   #   $.%$
  \end{shaded}

  To quote a string containing special characters, you can use 
  either single or double quotes. If your string, however, contains
  both such characters, you can define any character as an own
  delimiter, by preceding it with the "carat" (\verb#^#) character:
  \verb#attribute=^*My "good" value^*# makes use of a carat-escaped
  asterisk as a delimiter.

  Typically, an xRSL job description starts with an ampersand
  ("\&") , to indicate implicit \textbf{conjunction} of all the
  attributes:
  \begin{shaded}
    \verb#&(attribute1=value1)(attribute2="value 2")...#
  \end{shaded}

  Whenever a \textbf{disjunct}-request of two or more attributes is
  needed, the following construction can be used:
  \begin{shaded}
    \verb#(|(attribute="value1")(attribute="value2")...)#
  \end{shaded}

  In expressions, the following operands are allowed:
  \begin{shaded}
    \verb#=   !=   >   <   >=   <=#
  \end{shaded}

  \textbf{Commented} lines should start with "(*" and be
  closed with "*)":
  \begin{shaded}
    \verb#(*attribute="value1"*)#
  \end{shaded}
  Comments can not be nested.

  \textbf{Multiple job} description in one file is realized via a standard
  \globus\  RSL multi-request operand "+", which should precede
  multiple job description: 
  \begin{shaded}
    \verb#+(&(...))(&(...))(&(...))#
  \end{shaded}

  The xRSL attributes can be written in a single string, or split in
  lines arbitrary; blank spaces between and inside
  \verb#(attribute="value")# relations are ignored.

  \section{URLs}
  \label{sec:url}

  \input{xrsl-urls}

  \chapter{Attributes}
  \label{sec:attributes}

  Most of the attributes \index{attributes} introduced originally by \globus\ RSL~1.0 are
  supported, some with modifications as indicated in this
  document. Many new attributes are introduced, of which some are
  to be specified in the user's script, and others are internal for
  the GM (are added or modified by the UI).

  Attribute names are case-insensitive, although assigned values may
  well be case-sensitive, if they represent file names, environment
  variables etc..

  \section{User-side attributes}
  \label{sec:user}

  The following attributes \index{attributes:user-side} can be
  specified in a user's xRSL script. Some are to be modified by the UI
  before being passed to the GM. If this is the case, corresponding GM input is described.

  \input{xrsl-user}

  \section{GM-side attributes}
  \label{sec:gm}

  The following attributes are constructed by the UI and are parsed to
  the GM, although users may specify them manually (\textbf{not
  advised!}).

  \input{xrsl-gm}

  \section{Unsupported Globus RSL attributes}
  \label{sec:globus}

  The following \globus\ attributes are not supported by the ARC
  middleware. Whenever they are specified, a warning is issued by the
  UI, and the corresponding attribute is ignored.

  \subsection{Unsupported RSL 1.0 attributes}

  \begin{itemize}
    \item \verb#(resourceManagerContact=<string>)#
    \item \verb#(directory=<string>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(project=<string>)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(label=<string>)#
    \item \verb#(subjobCommsType=blocking-join|independent)#
    \item \verb#(subjobStartType=strict-barrier|loose-barrier|no-barrier)#
  \end{itemize}

  \subsection{Unsupported GRAM RSL attributes}

  \begin{itemize}
    \item \verb#(directory=<string>)#
    \item \verb#(fileCleanUp=<array>)#
    \item \verb#(fileStageIn=<array>)#
    \item \verb#(fileStageInShared=<array>)#
    \item \verb#(fileStageOut=<array>)#
    \item \verb#(gassCache=<path>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(jobType=<string>)#
    \item \verb#(libraryPath=<path>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(project=<string>)#
    \item \verb#(remoteIoUrl=<string>)#
    \item \verb#(scratchDir=<string>)#
  \end{itemize}

  \chapter{XRSL and JSDL}
  \label{sec:jsdl}

  \input{xrsl-jsdl}

  \appendix

  \chapter{Examples}
  \label{sec:examples}

  \section{User-side xRSL script}

  \begin{verbatim}
&
(* test run: if "yes", only submits RSL without actual job start *)
     (dryRun="no")
(* some local variables defined for further convenience *)
     (rsl_substitution=("TOPDIR" "/home/johndoe")) 
     (rsl_substitution=("NGTEST" $(TOPDIR)/ngtest))
     (rsl_substitution=("BIGFILE" "/scratch/johndoe/100mb.tmp"))
(* some environment variables, to be used by the job *)
     (environment=("ATLAS" "/opt/atlas") ("CERN" "/cern"))
(* the main executable file to be staged in and submitted to the PBS *)
     (executable="checkall.sh")
(* the arguments for the executable above *)
     (arguments="pal")
(* files to be staged in before the execution *)
     (inputFiles = ("be_kaons" "")
     ("file1" gsiftp://grid.uio.no$(TOPDIR)/remfile.txt) 
      ("bigfile.dat" $(BIGFILE) )    )
(* files to be given executable permissions after staging in *)
     (executables="be_kaons")
(* files to be staged out after the execution *)
     (outputFiles=
        ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp")
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile")
        ("be_kaons.hbook"  gsiftp://ce1.grid.org$(NGTEST)/kaons.hbook)   )
(* user-specified job name *)
     (jobName="NGtest")
(* standard input file *)
     (stdin="myinput.dat")
(* standard output file *)
     (stdout="myoutput.dat")
(* standard error file *)
     (stderr="myerror.dat")
(* GM logs directory name *)
     (gmlog="gmlog")
(* flag whether to merge stdout and stderr *)
     (join="no")
(* request e-mail notification on status change *)
     (notify="bqfe john.doe@gmail.com jane.doe@mail.org")
(* specific queue to submit the job *)
     (queue="atlas")
(* maximal CPU time required for the job, minutes for PBS*)
     (CpuTime="60")
(* maximal time for the session directory to exist on the remote node, days *)
     (lifeTime="7")
 (* memory required for the job, Mbytes *)
     (Memory="200")
(* wall time to start job processing *)
     (startTime="2002-04-28 17:15:00")
(* disk space required for the job, Mbytes *)
     (Disk="500")
(* required architecture of the execution node *)
     (architecture="i686")
(* required run-time environment *)
     (runTimeEnvironment="Atlas-1.1")
(* number of re-runs, in case of a system failure *)
     (rerun="2")
  \end{verbatim}

  \section{GM-side xRSL script}

Note that the client tool expands all the \texttt{rsl\_substitution}
values and adds double quotes to \textbf{all} strings. It also
streamlines blank spaces.

Comment lines are removed from the GM-side xRSL; below they are shown
to explain details.

  \begin{verbatim}
&
(* saves RSL in a temporary file if "yes" *)
     ("savestate" = "yes" )
(* job submission to be performed if action is "request" *)
     ("action" = "request" )
(* submission host name *)
     ("hostname" = "ce1.grid.org" )
(* client software version *)
     ("clientsoftware" = "nordugrid-arc-0.6.0.3" )
(* walltime value added by the client, in seconds *)
     ("walltime" = "3600" )
(* test run: if "yes", only submits RSL without actual job start *)
     ("dryRun" = "no" )
(* some local variables defined for further convenience *)
     ("rsl_substitution" = ("TOPDIR" "/home/johndoe" ) ) 
     ("rsl_substitution" = ("NGTEST" "/home/johndoe/ngtest" ) )
     ("rsl_substitution" = ("BIGFILE" "/scratch/johndoe/100mb.tmp" ) )
(* some environment variables, to be used by the job *)
     ("environment" = ("ATLAS" "/opt/atlas" ) ("CERN" "/cern" ) )
(* executable *)
     ("executable" = "checkall.sh" )
(* arguments *)
     ("arguments" = "pal" )
(* files to be staged in before the execution *)
     ("inputfiles" = ("checkall.sh" "279320" )
        ("myinput.dat" "39806" )
        ("be_kaons" "8807" )
        ("file1" "gsiftp://grid.uio.no/home/johndoe/remfile.txt" ) 
        ("bigfile.dat" "104857600" )
     )
(* files to be given executable permissions after staging in *)
     ("executables" = "checkall.sh" "be_kaons" )
(* files to be staged out after the execution *)
     ("outputfiles" = ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp" )
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile" )
        ("be_kaons.hbook" "gsiftp://ce1.grid.org/home/johndoe/ngtest/kaons.hbook" )
        ("myoutput.dat" "" )
        ("myerror.dat" "" )
    )
(* user-specified job name *)
     ("jobName" = "NGtest" )
(* standard input file *)
     ("stdin" = "myinput.dat" )
(* standard output file *)
     ("stdout"="myoutput.dat" )
(* standard error file *)
     ("stderr" = "myerror.dat" )
(* flag whether to merge stdout and stderr *)
     ("join" = "no" )
(* request e-mail notification on status change *)
     ("notify" = "bqfe john.doe@gmail.com jane.doe@mail.org" )
(* specific queue to submit the job *)
     ("queue" = "atlas" )
(* CPU time required for the job, converted into seconds *)
     ("cputime" = "3600" )
(* maximal time for the session directory to exist on the remote node, seconds *)
     ("lifetime" = "604800" )
(* memory required for the job, Mbytes *)
     ("memory" = "200" )
(* wall time to start job processing *)
     ("startTime" = "20020128171500" )
(* disk space required for the job, Mbytes *)
     ("disk" = "500" )
(* required architecture of the execution node *)
     ("architecture" = "i686" )
(* required run-time environment *)
     ("runtimeenvironment" = "Atlas-1.1" )
(* number of re-runs, in case of a system failure *)
     ("rerun" = "2" )
(* original client xRSL with expanded string substitutions; shortened here *)
     ("clientxrsl" = "&(""dryrun"" = ""no"" )(... )(""rerun"" = ""2"" )" )
  \end{verbatim}

  \section{JSDL script}

  \begin{verbatim}

<?xml version="1.0" encoding="UTF-8"?>
<jsdl:JobDefinition
    xmlns:jsdl="http://schemas.ggf.org/jsdl/2005/11/jsdl"
    xmlns:jsdl-posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
    xmlns:jsdl-arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
 <jsdl:JobDescription>
    <!-- Generic job description (jobname in XRSL) -->
    <jsdl:JobIdentification>
        <jsdl:JobName>JSDL Hello World job</jsdl:JobName>
        <jsdl:Description>
          This is a simple Hello World job test
        </jsdl:Description>
    </jsdl:JobIdentification>
    <!-- Standard application (same as executable, arguments in XRSL) -->
    <jsdl:Application>
        <jsdl:ApplicationName>Echo</jsdl:ApplicationName>
        <jsdl-posix:POSIXApplication>
          <jsdl-posix:Executable>/bin/echo</jsdl-posix:Executable>
          <jsdl-posix:Argument>Hello world</jsdl-posix:Argument>
        </jsdl-posix:POSIXApplication>
    </jsdl:Application>
    <!-- Get the GM diagnostics (same as gmlog in XRSL) -->
    <jsdl-arc:LocalLogging>
	      <jsdl-arc:Directory>gmlog</jsdl-arc:Directory>
    </jsdl-arc:LocalLogging>
 </jsdl:JobDescription>
</jsdl:JobDefinition>

  \end{verbatim}

  \chapter{JSDL-ARC schema}
  \label{sec:jsdl-arc-schema}

  \input{jsdl-arc-schema}

  \bibliography{grid}
  \printindex

\end{document}