### 2009-04-28 doc/sec/papers/GRID09/paper.tex
Redo some performance testing; change the text

\documentclass[conference]{IEEEtran}
\usepackage[pdftex]{graphicx}
%\usepackage{epstopdf}
\graphicspath{{./png/}{./eps/}}
\DeclareGraphicsExtensions{.pdf,.png}
\usepackage[caption=false,font=footnotesize]{subfig}
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\title{The design and implementation of standards-based Grid single sign-on using federated identity}

\author{\IEEEauthorblockN{Weizhong Qiang, Aleksandr Konstantinov}
\IEEEauthorblockA{Department of Physics, University of Oslo}
}

\maketitle

\begin{abstract}
Security infrastructure is one of the most challenging tasks in the developement, integration and
deployment of Grid middlewares. Even though the Grid community addresses the security issue 
through public key infrastructures (PKI) to support mutual authentication using X.509 certificates,
maintaining the X.509 credential is not that easy for non-IT specialists, and is proved to be an 
obstacle for the more widely deployment of Grid technology. Considering the identity federation 
technology is a emergingly widely used technology which can facilitate cross-domain single sign-on,
without requiring the users to maintain the X.509 credentials rather than their own institutional
accounts, we believe utilizing identity federation for Grid middlwares is a promising direction 
for Grid technology to be more widely used.
This paper describes a single sign-on infrastructure developed as part of the NorduGrid/ARC grid 
middleware. It adopts the identity federation standard (SAML), as well as Web Service standand. 
It focus on single sign-on solution on the middleware level for users to access Grids by only 
using their frequently used accounts, without being bothered to maintain X.509 credentials. 
Unlike other research which utilize identity federation on the Grid portal level, solution on
the middleware level can provide more flexibility.
Also, the performance of single sign-on solution is measured. We identify the performance limitation 
of security-related services inside this solution, and analyze the ways to avoid the limitation.
To our knowledge, the work described in this paper is the first evaluated implementation that utilizes 
identity federation for Grid usage on the middlware level.
\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Grid infrastructures facilitate the interoperability between wide-scale, cross-domain heterogeneous 
resources, as well as the accessibility provided to users. In terms of security issue, the Grid security 
infrastruture should provide accessibility as much as possible without loosing the security benifits.

The current Grid community uses Grid Security Infrastructure (GSI)~\cite{IFoster98} as the de facto 
standard for authentication and transport level security, which builds on the Public Key Infrastructure
(PKI) to support authentication using X.509 certificate. Mutual authentication is required 
by GSI based Grid deployment. Mutual authentication means that Grid users have to possess X.509 
credentials. To obtain a X.509 credential, a user needs to contact the Registration Authority (RA)
which will check the user's information and approve his certificate request; afterwards,
the Certificate Authority (CA) can issue the certificate according to the approval from 
the RA. The whole process is arduous and may take some time.
Meanwhile, to maintain a X.509 certificate is also not so easy for the non-IT community to deal 
with, since users have to periodically create proxy certificates by using Grid-specific command line 
utilities, such as grid-proxy-init or voms-proxy-init. Also maintaing the CAs or RAs is not a  
easy task, especially when the community of Grid users is getting much wider.

On the other hand, users normally could already have had some frequently-used campus or institutional 
accounts/credentials such as username/password. We believe that enabling Grid users to use their institutional 
credentials rather than the X.509 certificate to access Grids is a promising way 
to make the Grids more easily accessible, as well as to extend the user community of Grids
to a much larger scale.

To enable users to utilize the institutional credentials to access Grid, some issues needs to be 
addressed. Firstly, the Grid middleware needs to be improved to support the authentication based on
institutional credentials instead of X.509 certificates; Secondly, in order to interoperate with
those Grid infrastructures which require X.509 certificate based mutual authentication, an approach
should be provided for obtaining X.509 certificate based on the institutional credentials; Thirdly,
the single sign-on characteristic should be provided so that users can authenticate once and then transverse 
from Grid resources to Grid resources without being prompted to authenticate again at each of the resources.

The approach in the new version of ARC grid middleware is to utilize the identity federation 
standards, specifically, the Security Assertion Markup Language (SAML) specification, and Web 
Service standards, to facilitate the Grid single sign-on.

The rest of this paper is organized as follows: Section \ref{sec:arcmiddleware} presents the 
ARC Grid middleware, especially the architecture of the new version of ARC. Section 
\ref{sec:intergrationSAML2SSO} describes the solution of using federated identity for 
Grid authentication by using SAML2SSO profile; Setion \ref{sec:fedidtoX509} describes
how to obtain a X.509 credential based on federated identity, and how to use it for 
accessing multiple Grid resources; Section \ref{sec:perfeval} is performance evaluation;
Section \ref{sec:relatedwork} describes the related work; Section \ref{sec:conclusion} 
contains conclusions and future work outlook.

%-------------------------------------------------------------------------
\section{ARC Grid middleware}
\label{sec:arcmiddleware}

ARC (Advanced Resource Connector) is an open source Grid middleware solution released 
under Apache license. ARC middleware aims at developing self-organized, 
fault-tolerant, non-intrusive, easy-manageable Grid middleware~\cite{MEllert07}. 
The current production version of ARC provides Grid services for job submission and 
management, resource characterization, resource aggregation and discovery, basic data 
management, integration of Grid security solution, and so on. This release has been 
deployed and used in production environment, and is one of the widely deployed Grid 
middlewares in Europe.

The development version of ARC components is developed by the KnowARC project~\cite{KnowARClink}, 
based on the functionality of the current production ARC middleware. It aims at 
implementing a Web Service oriented Grid middleware which will provide higher levels of 
resource and user abstraction through well-defined Web Service interface~\cite{KnowARCDesignlink} 
in order to provide interoperability with other Web Service oriented Grid middlewares, as well as 
other Web Service compatible applications.

As the key part of the implementation of new ARC middleware, there is a lightweight Web 
Service container called Hosting Environment Daemon (HED) which provides hosting environment 
for various services at application level, as well as a number of modules to support flexible, 
interoperable, and efficient communication mechanism for building SOAP based Web Services. 
The design of the HED is built around the idea of flexibility and modularity,
such that the application service developers can simply concentrate on the application 
level Web Service implementation by only using the core minimum amount of components. It 
also simplifies work on the middleware level, e.g. making possible to implement another 
communication protocol or authentication mechanism. A system administrator can easily 
configure and deploy the middleware and application for different kinds of requirements 
without having to know much about the implementation.

The architecture of the HED is illustrated in Figure \ref{fig:HED}. In general, there 
are few components called Message Chain Component (MCC) which are in charge of implementing 
different protocol levels. For instance, as shown in the example message flow, the HTTP MCC 
will process stream from the TLS MCC to parse the HTTP message and pass its body to the SOAP 
MCC, and also process the SOAP response from the SOAP MCC to generate the HTTP message 
for the TLS MCC.

Dotted line in Figure \ref{fig:HED} shows an alternative path for the information 
to propagate among MCCs. A service administrator can configure the MCCs according to the 
interoperability requirements with the other part. For instance, the configuration marked 
with the dotted line is compatible with the WSE (Web Services Enhancement for .NET) SOAP 
message mechanism (see WSE’s SoapSender and SoapReceiver)~\cite{WSElink}. Another configuration 
could be SOAP over HTTPG (HTTP over GSI) which is needed to interoperate with services 
like the Storage Resource Manager (SRM)~\cite{A.Shoshan03} service. This shows the 
flexibility of HED in terms of protocols support.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED.pdf}
\caption{The example of Host Environment Daemon deployed with A-REX and File services}
\label{fig:HED}
\end{figure}

HED contains a flexible security framework for implementing and enforcing security 
related functionality, such as authentication and authorization. 
Each security related functionality can be implemented as a pluggable and configurable 
component (plug-ins) called SecHandler. Each MCC or Service is usually configured with 
two queues of SecHandler -- one for the incoming message and one for the outgoing message 
respectively. In Figure \ref{fig:HED}, the ``AuthZ'' and ``AuthN'' sub-modules inside 
MCCs and Services are examples of SecHandler.

%-------------------------------------------------------------------------
\section{Grid authentication using federated identity: Integrating SAML2SSO profile}
\label{sec:intergrationSAML2SSO}
Identity federation is an emerging technology which enables the identity information to be 
trustily transfered across autonomous security domains. By using identity federation, users 
from one domain can access another domain without the need for directly trust relationship 
between users and accessed domain, i.e., users can use accounts from their host domains to 
access external domains, and accessed domain does not need to maintain accounts for those 
external users. Identity federation doesn't enforce specific authentication mechanisms, so 
that various types of authentication can be supported between user and its host domain, and 
user is able to use his frequently used account/credential to accomplish authentication. 
Considering the Grid use cases, we see that utilizing the existing account to access Grid 
without being bothered to applying and manipulating a X.509 credential should be attractive 
for Grid users.

Shibboleth is the one of the several implementations of the identity federation, which is 
based on the OASIS Security Assertion Markup Language (SAML) 2.0 specifications. In terms 
of authentication, the Shibboleth implements SAML2.0 web browser SSO profile, which  defines 
two functional components, an Identity Provider and a Service Provider. The Identity Provider 
(IdP) is responsible for creating, maintaining, and managing user identity, while the Service 
Provider (SP) is responsible for controlling access to services and resources by analyzing 
the SAML assertions produced and issued by the IdP upon request (request from client application 
to the service/resource that is protected by SP).

In the implementation of ARC middleware, SAML2.0 web browser SSO profile is supported by 
implementing the functionality of service provider and web browser's user agent, and utilizing 
Shibboleth IdP implementation for the functionality of identity provider. The user agent 
functionality is for mimicing the behavior of web browser's user agent for authentication, 
such as the HTTP redirection and the HTTP cookie processing. The implementation of the user 
agent is based on the the HTTPS client interface of ARC, since the client interface of ARC 
can support different protocols which are incarnated by different MCCs.

\begin{figure}
\includegraphics[width=1.0\columnwidth]{SAML2SSO_UML.pdf}
\caption{Sequence diagram of SAML2.0 SSO profile integration in SOAP invocation 
between ARC WS Client and Web Service}
\label{fig:SAML2SSOUML}
\end{figure}

Figure \ref{fig:SAML2SSOUML} shows how is SAML2.0 SSO profile integrated into the SOAP 
invocation between ARC WS Client and Web Service. Step 1 to step 5 in Figure \ref{fig:SAML2SSOUML} 
is similar to those steps depicted in Web Browser SSO profile[ref], with the differences that
we specify the way about how does the service provider determine which identity provider to use (identity
provider discovery), by enabling the client to send the IdP name to service provider; as well 
as specify the way about how does the client/user agent authenticate against service provider, 
by implementing the form-based HTTP authentication which uses username and password for authentication between 
user agent and service provider (rather than X.509 client authentication) and is compatible 
with the Username/Password login handler of Shibboleth IdP.

In step 6, the service provider (SP Service) verifies and checks the SAML response, decrypts and stores 
the SAML assertion into session/connection context. That assertion includes the AuthnStatement and 
AttributeStatement elements. From step 8 to step 10, the WS client launches the SOAP invocation via 
the same connection as the one which is used by user agent to contact SP Service, then Grid/Web Service 
checks the AuthnStatement stored in the session context to see if the AuthnStatement is valid or expired. 
If valid, service handles the SOAP request and returns the SOAP response to the WS client. Note
the ``connection'' and ``session'' are interchangeably used in this paper, with the same meaning for 
TCP connection.

Note that in the current implementation the service requires WS client to reuse the same 
TCP connection as the one used by user agent in step 5, in order to guarantee that the validity of 
the SAML2SSO result is applicable to the SOAP invocation process. On the service side, 
the sesion reusing is accomplish by deloying the SP service and other functional Grid/Web Service(s) 
together in the same service container.

There are a few benifits about integrating SAML2.0 SSO profile for Grid/Web service authentication. 
Firstly, the client certificate authentication can be switched off, so that users do not need 
to apply and maintain a X.509 credential. Secondly, some implementation of identity provider, such as 
Shibboleth IdP, can cache the authentication result through session management once the user
has succeeded to authenticate, and for a short period this authentication result is valid, so that 
the next time the user authenticates against IdP, the user agent can just feed IdP with the last 
successfully authenticated session's identity rather than feed IdP with username and password again. So 
the user (or the client on behalf of that user) can travel across multiple security domains with 
only providing his name and password once, which is also the characteristic of single sign-on.
Lastly, even though there are a few implementations of identity providers, most of them are 
standard-compliant, so the solution implemented in ARC middleware can interoperate with other 
identity provider implementations with minimum change.

%-------------------------------------------------------------------------
\section{Bridging federated identity and X.509 credential}
\label{sec:fedidtoX509}
Many widely used Grid middlewares are based on GSI which requires mutual X.509 authentication. 
Meanwhile, Web Service based Grid applications mostly do require client certificate authentication.  
To bridge the gap beween federated identity and X.509 credential, a short lived credential 
service (SLCS) is implemented through which a user can get a short-lived X.509 certificate only by
providing his username/password and authenticating through the SAML2SSO profile.

Also, in order to provide single sign-on capability for services that act on behalf of user to 
invoke other services, a X.509 credential delegation service is implemented, which is based on 
SOAP communication rather than the proprietary GSI communication which is used in the delegation
services such as MyProxy.

\subsection{Short-lived credential service}
\label{sec:slcs}
The sequence diagram of SLCS process is the same as the diagram shown in 
Figure \ref{fig:SAML2SSOUML}, with SOAP request (the message of step 8 of 
Figure \ref{fig:SAML2SSOUML}) including X.509 request 
and SOAP response (the message of step 10 of Figure \ref{fig:SAML2SSOUML}) 
including an signed X.509 certificate and the CA certificate that is used to 
sign this X.509 certificate. 
In detail, the SLCS client firstly accomplishes the SAML2SSO profile, then sends 
SOAP request to the SLCS service with the X.509 request embedded. The SLCS service makes
access control decision according to the SAML assertion stored in session context, and issues 
a certificate (with 12 hours lifetime) with the SAML attributes as the X.509 certificate 
extension. The SLCS client then gets the SOAP response with the X.509 certificate embedded.

Since the lifetime of the short lived credential has a short lifetime, it is permissible to protect 
the private key by using the local file system permissions instead of encrypting it by 
using a pass-phrase. Therefore, no manual interactivity is required to uncrypted the private
key. Hence single sign-on is provided, since the user only needs to input his password when 
he starts SAML2SSO profile.

How to compose a distinguished name (DN) for the certificate is a critical issue for 
the SLCS service. Since the ``eduPersonPrincipalName'' (see eduPerson schema~\cite{eduSchemalink})
is identical to the user that accomplishes SAML2SSO profile. We pick up ``eduPersonPrincipalName'' 
from SAML attributes that returned from Shibboleth IdP, and use it to compose the DN. A 
typical example of the eduPersonPrincipalName 
could be alice@example.org. In such case composed DN is for example\\``/O=knowarc/OU=example.org/CN=alice''.

By using the SLCS service, the user can easily access the Grid services that requires 
X.509 credential on client side, at anywhere, simply by running the SLCS client command 
and providing his username and password.

\subsection{X.509 credential delegation service}
\label{sec:creddeleg}

\begin{figure}
\includegraphics[width=1.0\columnwidth]{Delegation_UML.pdf}
\caption{Sequence diagram of the usage of delegation service}
\label{fig:DelegUML}
\end{figure}

The sequence diagram is shown in Figure \ref{fig:DelegUML}. Firstly, the client needs to delegate
the proxy certificate or short-lived certificate to a delegation service through step 1 to step 7.
Then this client invokes the peer Web Service by appending the delegation information (URL of delegation
service and the delegation ID corresponding to the above delegation) into the SOAP body. The Web 
Service then sends a X.509 request and the delegation ID to the delegation service 
which is specified by client side, and gets back a delegated certificate which can be composed together
with the private key for a proxy certificate. The Web Service could use this proxy certificate 
to represent the user to invoke another Web Service, which either could cause one more process of 
delegation (step 1 to step 13) if another delegation service is specified, or could directly start
from step 8 if the same delegation service is specified as the one in the former delegation process.

The client only needs to delegate the certificate to delegation service one time per session, it does
not need to do delegation for subsequent SOAP invokations after the first one.

The functionality of both client and service sides are implemented as pluggable handlers which can be 
configured into the client and service configuration, so that the service or client implementation itself
does not need to be changed.

We suppose the short-lived certificate is used by client, so that the SLCS service together with
the delegation service can provide single sign-on from user's federated identity, while keeping the 
interoperability to those Grid/Web services that requires mutual authentication using certificate.

%-------------------------------------------------------------------------
\section{Performace Evaluation}
\label{sec:perfeval}
The following sections evaluate the performance of the HED versus Axis2/C, as 
well as the performance about the SAML2SSO profile, short-lived credential
service, and delegation service.

\subsection{Performance comparison between HED and Axis2/C}
\label{sec:perhedandaxis}
Since the implementation of HED includes an alternative SOAP implementation which is the base of 
the security related web servies described in this paper, it is useful to compare the performance 
of HED and other SOAP implementations. Although there are many SOAP implementations such as 
AxisJava, gSOAP, XSOAP4, etc., the goal of this paper is not about complete comparison between
HED and all the other SOAP implementations, so we use the Axis2/C (v1.5.0) to demostrate the performance
HED.

For the service side, we implement two simple ``echoString'' services in HED and Axis2/C. The 
``echoString'' provides a simple SOAP portType which will output the same value as it gets from
input.
For client side, we develope a client by using the client API of ARC, to invoke the ``echoString''
services from Axis2/C and HED. The service invocation is just sending a short string to service, and 
getting back the same string. One test/experiment consists of a set of clients which run simultaneously. 
Each client runs for a fixed period of time, and invokes the service as many times as it can during 
the test period. Multiple clients are launched by creating multiple threads in one test.
Two values are measured: average response time and throughput. Average response time is computed
by counting the average value of the response time for each invocation. Throughput is computed
by counting the number of invocation during each second.

The two ``echoString'' services run in a test machine equipped with 3GHz Intel Pentium dual-core 
CPU and 2GB memory, Red Hat Enterprise Linux 5. The client runs in another machine equipped 
with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. The Axis2/C service is run using 
the httpd module on Apache Http Server version 2.2.11 instance. We keep all the configurations 
default in Apache Http Server.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{HED2Axis_thread10.pdf}
\caption{Performance results for HED and Axis2C: 1 service and 10 clients}
\label{fig:HED2Axis_thread10}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{TCP_thread_all.pdf}
\caption{Performance of HED and Axis2C echoString services without TLS: performance
changes with client numbers; session non-reused }
\label{fig:TCP_thread_all}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{TLS_thread_all.pdf}
\caption{Performance of HED and Axis2C echoString services with TLS: performance
changes with client numbers; session non-reused}
\label{fig:TLS_thread_all}
\end{figure}

Figure \ref{fig:HED2Axis_thread10} is the results of performance comparison with ten 
clients being launched.
Two types of experiments are shown: session reused, session non-reused.
``Session reused'' means all of the SOAP invocations on each client share the same session 
(tcp connection); while ``Session non-reused'' means each invocation launches a new session.

For ``Session reused'' experiments, we see that the average response time in HED is about 81ms
which is twice of the value in Axis2/C, and the throughput in HED is about half of the value in Axis2/C. 
One obersavation is that, the performance results of HED is quite stable, with the average response time 
and throughput almost kept unchanged under different time points.
For ``Session non-reused'' experiments, the average response time in HED is almost the same as
the value in Axis2/C, with the value 15ms. On the other hand, the throughput in HED is close to
 the value in Axis2/C.

Since the echoString service is a simple service which occupies little and stable time for processing
service logic itself, and also the experiments are done under local network, we can speculate 
the response time we get in Figure \ref{fig:HED2Axis_thread10} is mainly about the processing time 
in container itself.
With the stability characteristic of performance, we can speculate that the service container 
itself will stablly impact on the performance of other services.

To more deeply compare the performance of services from HED and Axis2/C, some tests have
been done to investigate how does the number of simultaneous clients impact on the performance.
Figure \ref{fig:TCP_thread_all} is the performance results for echoString services with client number
being changed. In this set of experiments, the session is not reused, and secure communication is not
used. Each experiment is executed for a relatively long duration (two minutes), so that the there are
enough times of service invokations and then the performance results are reliable.
We see that both the response time of services in HED and Axis2/C starts from similar value 
(when client number is five) and changes linearly with number of clients, while service in HED 
provides shorter response time with the value being around two thirds of the response time of 
service in Axis2/C when client number becomes bigger.
In case of throughput, service in HED provides stable throughtput, while service in Axis2/C provides 
throughput which slightly becomes smaller when client number becomes bigger.

Figure \ref{fig:TLS_thread_all} is the performance results when secure communication is used.
For Axis2/C, we configure the http server (that hosts Axis2/C) with SSL module enabled.
Each experiment is executed for a duration of two minutes.
We see that, similar to Figure \ref{fig:TCP_thread_all}, the both the response time of services 
in HED and Axis2/C also starts from similar value (when client number is five) and changes linearly 
with number of clients, while the service in HED also provides much shorter response time 
with the value being around sightly more than one third of the response time of service in Axis2/C 
when client number becomes bigger.
In case of throughput, the throughput of service in HED increases linearly with client numbers intially,
however, it starts to keep stable to be a fixed value (around 13000 requests per minutes) when client 
number is bigger than 30. On the other hand, the throughput of service in Axis2/C starts to keep 
stable when client number is bigger than 10, with the value being around sightly more than one 
third of the throughput of service in HED.

We conclude that the HED can provide better performance than Axis2/C, in both cases of secure 
communication being enabled and disabled, especially when the number of simultaneous clients 
is bigger.

\subsection{Performance of the integration SAML2 SSO profile}
\label{sec:perfsaml2sso}
In our next experiment, we evaluate how does the integration of SAML2SSO profile impact on the 
performance of Web Service invocation. We configure a ``echoString'' service together with a SP (
Service Provider) service, and develope a client by using SAML2SSO related API of ARC, to invoke 
the ``echoString'', as well as interact with the IdP (Identity Provider). Service and clients run on
the same set of machine as section \ref{sec:perhedandaxis}. Meanwhile, we deploy the Shibboleth IdP 
implementation on a machine equipped with 2.8GHz Intel Pentium CPU and 1GB memory, Red Hat Enterprise 
Linux 5.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread10.pdf}
\caption{Performance of HED echoString services with SAMLSSO profile: 1 service and 10 clients}
\label{fig:SAML2SSO_thread10}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SAML2SSO_thread_all.pdf}
\caption{Performance of HED echoString services with SAMLSSO profile:
performance changes with client numbers; session non-reused }
\label{fig:SAML2SSO_thread_all}
\end{figure}

Figure \ref{fig:SAML2SSO_thread10} depicts the average response time and throughputs for 10 clients
running simultaneously against the ``echoString'' service enhanced with SAML2SSO profile. Since 
the success of authentication from IdP can be regarded as valid for the whole session between client 
and service, the client with ``Session reused'' need to accomplish the SAML2SSO profile only once 
for all SOAP invocations on this client, while the client with ``Session non-reused'' will need to 
accomplish this profile once per SOAP invocation.

The average response time of ``Session non-reused'' experiments changes little with the execution 
duration, and the value is about 3000ms; while the throughput is around 200 requests per minutes.
Considering that the average response time in case of ``Session non-reused'' and pure SOAP invocation 
is around 15ms when client number is 10 (see section \ref{sec:perhedandaxis}), therefore we can 
calculate that the consumption of time for SAML2SSO profile is about 99.5\% of the whole SOAP 
invocation with SAML2SSO profile integrated.

Deeper investigation about recording the time of each step of SAML2SSO profile supports the 
above obersavation. The average time cost for getting an authencation request from SP service 
is about 60ms; the average time cost for authenticating against IdP is about 2750ms; the 
average time cost for sending the SAML response (from IdP) to SP service is about 20ms;
the average time cost for verifying (by SP service) the signature of SAML response is 
around 130ms.

In order to improve the performance of SAML2SSO profile, optimizing the process between 
client and IdP is the correct direction, and probably one of the easiest way is to enhance the 
hardware configuration.

One the other hand, the average response time of ``Session reused'' experiments is about 82 ms,
which is almost as short as the value in the pure SOAP invocation experiment in section 
\ref{sec:perhedandaxis}. The result is resonable because the time consumption caused by 
only once of SAML2SSO execution on one client is relatively small comparing with the time 
consumption caused by the following hundreds of pure SOAP invocations on the same client.

Figure \ref{fig:SAML2SSO_thread_all} depicts the performance results for echoString services
that are enhanced with SAML2SSO profile, with client number being changed. In this set of experiments, 
the session is not reused, and secure communication without client certificate authentication is used.
The results show the average response time increases linearly with the client number, while the 
throughput increases initially and then keeps in a fixed value (around 215 requests per minute) 
when the client number is bigger than 4. When the client number is four, the response time is 
around one second, while the throughput of the service is 215 requests per minutes which is almost
the fixed value.
We speculate that in order to get acceptable response time, there should be less than four (including four) 
concurrent clients if they need to continuously participate one SAML2SSO profile.

Considering that the throughput of ``echoString'' service enhanced with mutual TLS is 13000, and 
the concurrent client number when arriving fixed throughput is 30, the performance of
``echoString'' service enhanced with SAML2SSO profile is much worse, with the values being 215
and 4 correspondingly.

However for some typical grid application scenarios, normally the client does not need 
to continuously accomplish SAML2SSO profile. For instance, a user could submit a job 
which needs a period of time (normally this period is much longer than a few seconds)
to execute, which launches an execution of SAML2SSO profile; 
afterwards, this user could query the information as well as the result about this job, which 
launches another execution of SAML2SSO profile. On the other hand, statistically, users will most
probably not access grid system in a concurrent way, rather than a random way. Therefore, if we 
suppose a user could access grid once per minute, in our test environment, we can claim that 
around 215 clients randomly can be supported with acceptable performance.

\subsection{Performance of SLCS service}
\label{sec:perfslcsserv}

\begin{figure}
\includegraphics[width=0.9\columnwidth]{SLCS_thread_all.pdf}
\caption{Performance of SLCS service: performance changes with client numbers}
\label{fig:SLCS_thread_all}
\end{figure}

Figure \ref{fig:SLCS_thread_all} is the performance about the SLCS service. Since the SLCS is 
based on the SAML2SSO profile, the reponse time should be the accumulation of time consumption 
about the SAML2SSO processing and the SLCS processing. In this experiment, SLCS service and 
client run on the same set of machine as section \ref{sec:perhedandaxis}.

The results show that the throughput of SLCS service has the same characteristic as the throughput
of ``echoString'' service enhanced with SAML2SSO profile, with the value being around 215 
requests per minute; and the average response time resembles the value of ``echoString'' service as 
well, with the value being increased for about 80ms. The increased time consumption is caused by 
the creating of X.509 request (1024-bit RSA key pair generation) on the client side.

Even though the average response time of SLCS service is not very short, considering the short-lived 
credential has a relatively long lifetime (12 hours by default) and therefore client only needs to invoke
SLCS service once per 12 hours, the performance is completely acceptable.

\subsection{Performance of delegation service}
\label{sec:perfdelegserv}

In the last experiment, we evaluate how does the usage of delegation service impact on the 
performance of Web Service invocation. We deployed a ``echoString'' service plugged with delegation
security handler on a machine equipped with 3GHz Intel Pentium dual-core CPU and 2GB memory, Red 
Hat Enterprise Linux 5, and also configure a client together with delegation security handler plugged 
on a machine equipped with 2.0GHz AMD Athlon CPU and 1GB memory, Red Hat Enterprise Linux 5. Aside from 
``echoString'' service and client, a delegation service is deployed on a machine equipped with 
3GHz Intel Pentium dual-core CPU and 2GB memory, Red Hat Enterprise Linux 5.

\begin{figure}
\includegraphics[width=0.9\columnwidth]{Delegation_thread_to_perf.pdf}
\caption{Performance of HED echoString service with credential delegation: performance 
changes with client numbers; echoString service contacts delegation service upon each service 
invokation from client}
\label{fig:Delegation_thread_to_perf}
\end{figure}

In the experiment, the client launches a new credential delegation for each service invokation, 
and ``echoString'' service acquires delegated credential from delegation service upon each 
service invokation.
Figure \ref{fig:Delegation_thread_to_perf} depicts the average response time and throughput of 
the ``echoString'' service with delegation security handler embedded. The average response time 
inreases linearly with the client number, while the throughput starts to keep at fixed value
(around 400 request per minutes) when the client number is 8, after initially increasing.
We conclude that if the client number is bigger than eight, in our test environment, there 
is no throughput benefit can be achieved by adding clients.

%-------------------------------------------------------------------------
\section{Related Work}
\label{sec:relatedwork}
There are other researches that utilize Shibboleth for protecting usage of Grid resources. 
The work in ~\cite{J.Watt06} uses Shibboleth to protect a GridSphere portal and then 
indirectly protect Grid services which are deployed in this GridSphere protal. Another 
work in ~\cite{R.O.Sinnott06} also uses Shibboleth to protect Grid portal, moreover, it 
maintains a few common used server certificates which can then be used to access Grid 
resources after users have successfully authenticated themselves at Grid protal through
Shibboleth. Unlike these two solutions that use Shibboleth to protect Grid portal, the solution
in this paper uses Shibboleth or similar standards-based identity federation to directly protect 
Grid resource/services, which provide possibility for those Grid applications that are directly
based on Grid interface (e.g., applications that invokes SOAP operations of remote Grid services),
to benefit from identity federation.

GridShib~\cite{T.Scavo07,VWelch05} provides one solution that users can use X.509 
certificate to authenticate to Shibbolth IdP, get back SAML assertion and embed it in 
the proxy certificate. The Grid service will then use this SAML assertion for access control.
Unlike the work described in this paper, GridShib still requires users to possess a X.509 
credential, rather than username/password. GridShib also provides another 
solution~\cite{TBarton06} in which a credential service (MyProxy server) is deployed together
with Shibboleth IdP for acting as an online-CA, authenticating the user through username/password
rather than X.509 certificate, and issuing short-lived X.509 credential. The second solution
is similar to the short-lived credential service in our solution, while our solution
also provides the option for users directly authenticating to Grid services via SAML2.0 SSO 
profile using their username/password.

Another short-lived credential service is also provided by ~\cite{switchslcslink}. 
The advantage of our work comparing to the work in ~\cite{switchslcslink} is that 
our SLCS service is based on Web Services standard, so that it is easier to achieve 
interoperability.

In the aspect of credential delegation, apart from the credential delegation mechanism of 
GSI~\cite{IFoster98,VWelch04}, several authors~\cite{MAhsant04} propose a delegation protocol 
based on the WS-Trust~\cite{WSTrustlink} to provide a standard and interoperable protocol 
for the delegation in Grids. WS-Trust will also be adopted by the ARC middleware to express the 
specifications required to define delegation protocol in a standard way. Gridsite project also
implements a X.509 credential delegation solution~\cite{GridSitelink} based on the Web 
Service, with which ARC delegation client can be easily made interoperate.

%-------------------------------------------------------------------------
\section{Conclusion And Future Work}
\label{sec:conclusion}
In this paper we propose a Grid single sign-on framework that can utilize users' 
federated identity instead of X.509 mutual authentication for authentication, which
we believe can make further uptake of Grid resources as well as Grid middleware by 
wider scale of users.

We implement the SAML2 SSO profile in the ARC Grid middleware, so that the Grid applications
that developed by using the application interface provided by this middleware can easily 
be switched to benefit from federated identity based authentication. 
The performance evaluation shows even though the introducing of SAML2 SSO profile causes 
much performance downgrade, this downgrade can be avoided by reusing the authentication 
result for multiple SOAP invocations.
Meanwhile, in order to be compatible with those Grid applications that require X.509 
certificate on the client side, we also implement one Web Service which can be used for 
obtaining an X.509 credential based on federated identity, and the other Web Service 
which can be used for delegating X.509 credential. We believe that the evaluated 
performance is enough for creating X.509 certificate and proxy certificate, considering that 
the certificates should have a relaively much longer lifetime comparing to the response time of 
these two services.

Although only authentication issue has been discussed in this paper, we know single
sign-in should also include the authorization issue. The SAML assertion which includes
users' attributes can be used on the serrvice side to achiece attribute-based authorization.
Therefore providing plugin on the service side to enforce access control using SAML attribute
is considered one part of the future work.

SimpleSAMLphp~\cite{simplesamllink} provides another implementation of SAML2 SSO profile, and it 
has been widely deployed around nordic countries for identity federation. Hence integrating
ARC based Grid applications with simpleSAMLphp is considered as another of future work.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

\section*{Acknowledgment}

The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

%\bibliographystyle{IEEEtran}

\begin{thebibliography}{1}

\bibitem{IFoster98}
I. Foster, C. Kesselman, G. Tsudik, and S. Tuecke, A Security Architecture for 
Computational Grids, ACM Conference on Computers and Security, 1998, 83-91.

\bibitem{AlfieriR05}
R. Alfieri, R. Cecchini, V. Ciaschini, L. dell’Agnello, A. Frohner, K. Lorentey, 
and F. Spataro. From gridmap-file to voms: managing authorization in a Grid environment. 
Future Generation Comp. Syst., 21(4), 549-558 (2005)

\bibitem{RFC3821link}
RFC 3821- An Internet Attribute Certificate Profile for Authorization. http://www.faqs.org/rfcs/rfc3281.html

\bibitem{GTlink}
Globus Toolkit. http://www.globus.org/toolkit/

\bibitem{gLitelink}
gLite: lightweight middleware for Grid computing. http://glite.web.cern.ch

\bibitem{ARClink}
Advanced Resource Connector. http://www.nordugrid.org/middleware/

\bibitem{WSSeclink}
OASIS Web Services Security. www.oasis-open.org/committees/wss/

\bibitem{SAMLlink}
OASIS Security Assertion Markup Languages (SAML). www.oasis-open.org/committees/security/

\bibitem{MEllert07}
M. Ellert et al. Advanced resource connector middleware for lightweight computational 
grids, Future Generation computer systems, 23(2), 219-240 (2007)

\bibitem{KnowARClink}
KnowARC project.  https://www.knowarc.eu/

\bibitem{KnowARCDesignlink}
Design document of new version ARC. https://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf

\bibitem{WSElink}
Web Services Enhancements 2.0 Service Pack 2. http://msdn.microsoft.com/en-us/library/

\bibitem{A.Shoshan03}
A. Shoshani, A. Sim, and J. Gu, Storage Resource Managers: Essential Components for the Grid, 
Grid Resource Management: State of the Art and Future Trends, Kluwer Publishing, 2003.

\bibitem{Shiblink}
The Shibboleth Project. http://shibboleth.internet2.edu/

\bibitem{eduSchemalink}
eduPerson and eduOrg Object shema. http://middleware.internet2.edu/eduperson/

\bibitem{R.O.Sinnott06}
R.O. Sinnott, J. Jiang, J. Watt, and O. Ajayi. Shibboleth-based access to and 
usage of grid resources. Proceeding of 7th IEEE/ACM International Conference on 
Grid Computing, September 2006, Barcelona.

\bibitem{J.Watt06}
J. Watt, O. Ajayi, J. Jiang, J. Koetsier, R.O. Sinnott. A Shibboleth-Protected 
Privilege Management Infrastructure for e-Science Education. in proceeding of Sixth 
IEEE International Symposium on Cluster Computing and the Grid (CCGRID'06), May 2006,
Singapore.

\bibitem{T.Scavo07}
T. Scavo and V. Welch. A Grid Authorization Model for Science Gateways. International 
Workshop on Grid Computing Environments, 2007.

\bibitem{VWelch05}
V. Welch, T. Barton, K. Keahey and F. Siebenlist. Attributes, Anonymity, and 
Access: Shibboleth and Globus Integration to Facilitate Grid Collaboration. in 
proceeding of 4th Annual PKI R\&D Workshop, 2005.

\bibitem{TBarton06}
T. Barton et al. Identity Federation and Attribute-based Authorization 
through the Globus Toolkit, Shibboleth, Gridshib, and MyProxy. in proceeding 
of 5th Annual PKI R\&D Workshop, 2006.

\bibitem{switchslcslink}
SWITCH Short Lived Credential Service. http://www.switch.ch/Grid/slcs/

\bibitem{VWelch04}
V. Welch et al. X.509 proxy certificate for dynamic delegation. in proceeding 
of the 3rd Annual KI R\&D Workshop, 2004.

\bibitem{MAhsant04}
M. Ahsant, J. Basney, and O. Mulmo. Grid Delegation Protocol. UK Workshop on 
Grid Security Experiences, Oxford. 2004.

\bibitem{WSTrustlink}
OASIS WS-Trust specification. http://docs.oasis-open.org/ws-sx/ws-trust/200512

\bibitem{GridSitelink}
Gridsite delegation service. http://www.gridsite.org/wiki/Delegation\_protocol

\bibitem{simplesamllink}
http://rnd.feide.no/simplesamlphp

\end{thebibliography}

\end{document}