### 2009-05-12 doc/tech_doc/infosys/infosys_technical.tex
Moving more from handbook. Started adding security parts.

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

{\raggedleft NORDUGRID-TECH-21\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large ARC Information System}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Documentation and developer's guide}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large \footnote{@} \large \par}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\renewcommand{\thefootnote}{\arabic{footnote}}

\chapter{Design Overview}
\label{cha:design_overview}

\chapter{ISIS}
\label{cha:isis}

\section{Registration handling}
\label{sec:isis_registration_handling}

\subsection{Functionality}
\label{sub:isis_functionality}

 The functionality of ISIS is made of two parts. On the one hand they are working as ordinary Web Services, and on the other hand maintain a peer-to-peer network - \texttt{ISIS cloud}.

 Main functionality of ISIS service visible from outside of ISIS cloud is to accept registration and provide collected information to clients. For that ISIS implements operations described in following section.
The single ISIS service accepts Registration Records pushed to it by other services (including ISIS services too) and stores them in local XML database. Stored records can be queried by clients using mandatory and service-specific attributes for selection criteria.

 In case of multiple ISIS services they form a peer-to-peer network. Inside that cloud Registration Records are propagated between services in such a way that all ISIS instances continiously try to synchronise their databases.

% TODO: put figure here

% subsection functionality (end)

\subsection{Interface} 
\label{sub:isis_interface}

\subsubsection{Operation Register}

\begin{description}

  \item[Input]~\begin{description}
    \item[Header]~\begin{description}
      \item[RequesterID] Identifier of the client.
      \item[MessageGenerationTime] Time when following set of RegEntry was generated. There may be multiple RegEntry elements.
    \end{description}
    \item[RegEntry]~\begin{description}
      \item[SrcAdv]~\begin{description}
        \item[Type] Type of service being registered. This element is opaque string for now. There shall be service types defined later.
        \item[EPR] Endpoint Reference of service being registered in terms of WS-Addressing.
        \item[SSPair] Set of key/value pairs representing service specific information.
      \end{description}
      \item[MetaSrcAdv]~\begin{description}
        \item[ServiceID] Globally unique and persistent identifier of the service.
        \item[GenTime] (Generation Time) The actual timestamp of information (called wake\_up\_time in the pseudo code below)
        \item[Expiration] Validity period of Service Advertisement record.
      \end{description}
    \end{description}
  \end{description}

  \item[Output]~\begin{description}
    \item[Fault] Optional element describing fault which occured while performing registration. If missing registration succeeded.
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation is usually called by service which wants to register it's presence in ISIS. This message consists of one or more \textbf{Registration Entry} and at most one \textbf{Registration Header}. The \textbf{Registration Entry}(RegEntry) contains a \textbf{Service Advertisement}(SrcAdv) and a corresponding \textbf{Service Advertisement Metadata}(MetaSrcAdv). This structure is shown on Figure \ref{fig:registration_message}.
\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{RegistrationMessage.png}}}
\caption{\label{fig:registration_message}Embedded structure of Registration Message} }
\end{figure}

Service must supply mandatory information. Among those the \texttt{Endpoint Reference} is used to contact service - only required element is contact URL of service. \texttt{Type} specifies kind of service and is used to find out functionality nad interface of service. \texttt{ServiceID} is used to distinguish between registered services and to deal with case of service changing it's contact URL. For more information about mandatory and optional information see Section \ref{service_advertisement}

As result of this operation new Registration Record is stored inside ISIS internal database and eventually propagated to other ISISes. If registration record of same ID already existed it will be renewed.

This operation is also used by ISIS services to propagate Registration Records inside ISIS cloud.

\subsubsection{Operation RemoveRegistrations}

\begin{description}

  \item[Input]~\begin{description}
    \item[MessageGenerationtime]
    \item[ServiceID] Multiple identifiers of services, whose records has to be removed.
  \end{description}

  \item[Output]~\begin{description}
    \item[RemoveRegistrationResponseElement]~\begin{description}
      \item[ServiceID] Identifier of service whose record was not removed
      \item[Fault] Description of failure reason
    \end{description}
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation is used to explicitely requesting the removal of zero or more Registration Entries associated with specified ServiceID values stored in the Information System. If corresponding record does not exist it's identifier will be present in response message together with corresponding Fault element.

\subsubsection{Operation GetISISList}

\begin{description}

  \item[Input]~\begin{description}
    \item[none]
  \end{description}

  \item[Output]~\begin{description}
    \item[EPR] Multiple Endpoint References of known ISIS services
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

In response to this operation the EndpointReferences to all known ISIS services are returned. The operation is used for obtaining a list of known ISIS instances from any particular ISIS. Clients can then use the obtained list to run direct queries against the ISIS instances. The operation is provided for fault tolerance and for providing optional performance boost. This operation returns the known peer-to-peer neighbors. If client is interested in all ISIS instances they can be obtained using the Query operation because they are ordinary services registered into the Information System.

\subsubsection{Operation Query}

\begin{description}

  \item[Input]~\begin{description}
    \item[QueryString] XPath query expression
  \end{description}

  \item[Output]~\begin{description}
    \item[any] Result of query
  \end{description}

  \item[Faults]~\begin{description}
    \item[none]No specific faults are defined
  \end{description}

\end{description}

This operation allows any XPath queries to be performed on stored Registration Records. The records are treated as merged in one XML document with each record being equivalent to \texttt{RegEntry} element of \texttt{Register} operation. In response all elements produced by XPath query are returned. The purpose of this operation is to make it possible to obtain any kind of information related to the Indexing Database.

% subsection interface (end)

% section registration_handling (end)

\section{Peer-to-Peer}
\label{sec:isis_peer_to_peer}

\subsection{Functionality}
\label{sub:peer_to_peer_functionality}

%\subsubsection{Peer-to-Peer functions}
%\label{ISIS Peer-to-Peer functions}
%\subsection{Client}
%\label{Client}

%The client part of the information service is for querying the data stored in the database of the ISIS 
%services. This is implemented as a part of the ARCLib and using the ISIS service's interface for the 
%various data queries. For further details see Section \ref{Query data}.

%\section{Operations}
%\label{Operations}
%The different operations are performed by different Information system components. These cases are described below in the dedicated subsections.

% subsection functionality (end)

\subsection{Interface}
\label{sub:isis_peer_to_peer_interface}

% subsection interface (end)

% section peer_to_peer (end)

\section{Authorization}
\label{sec:isis_authorization}

\subsection{Client Authorzation}
\label{sub:isis_authorization_client_authorization}

% TODO figure with mutiple service types

To ensure information stored in ISIS cloud can't be tampered and only available to proper clients following authorization framework is implemented. All actions perfored by ISIS clients are divided into three following groups:

\begin{itemize}
 \item Operations initiated by other ISISes in the cloud. Those include:
 \begin{itemize}
  \item Register with Registration Message containing information not about contacting client
  \item RemoveRegistrations with request to remove Registration Message representing not contacting client
  \item Connect % TODO: describe this operation
 \end{itemize}

 Those operations may cause uncontrollable changes in colleccted information and must be granted only for higly trusted entities like ISISes themselves.

 \item Operations initiated by the Services registering to Information System. Those are:
 \begin{itemize}
  \item Register with Registration Message containing information about contacting client
  \item RemoveRegistrations with request to remove Registration Message representing contacting client
 \end{itemize}

 \item Operations which are allowed for any liable client of particular Grid infrastructure.
 \begin{itemize}
  \item Query
  \item GetISISList
 \end{itemize}
\end{itemize}

% subsection client_authorization (end)

\subsection{Information Authorzation}
\label{sub:isis_authorization_information_authorization}

%* Clasification of information
%
% 1. From acccessibility point of view information can be devided
%into 3 groups:
% a. public
% b. available to trusted clients (like broker or scheduler service)
% c. private - available to owner of information object and entities
%trusted by owner.
%Distinction between 2 and 3 is not clear. One may accept that 2 is
%imposed by architecture/deployment of system and 3 is controlled by
%entities which are users of the system.
%
 2. From storage/access location point of view information is divided
%into:
% a. local - stored at endpoint representing resource which is
%described by information. This information is obtainable through
%ALIS interface.
% b. propagated to dedicated services (ISIS) which collect information
%needed for pre-selection and initial lookup of services
%representing resources.
%
%NOTE: Definition of propagation of information does not exactly belong
%to security architectire of information system. But it affects it significantly
%and hence at least minimal set of definitions is needed.
%
%OUT OF BAND NOTE: Those services could provide pre-selection capability.
%
% Sets 1 and 2 may intersect in any way. In first approach to simplify
%implementation 2b may be considered to always belong to 1a.
%
%
%* Implementaion guidelines
%
% To implement set 1 information objects need to have Access Control
%Rules (ACR) assigned (usually used term Access Control List seems to
%be inadequate due to meaning of word List).
%
% To implement set 2 in similar way information objects have Propagation
%Control Rules (PCR) assigned. Because curently information system is made
%of one layer of ISIS PCRs may consist of one boolean property.
%
% Information is structured in tree-like fashion - XML document.
%Access control is defined at node and follows all branches - similar
%to file systems. Propagation control is similar to access control (if
%needed at all). Children nodes can't widen access defined at parent
%nodes. This should decrease amount of ACR to be evaluated.
%
%EXAMPLE: if parent node A allows access only to VO_1 then children
%node B can narrow access to Adminsitrator of VO_1 and can't open
%access for VO_2.
%
% By default all nodes are public and are propagatable. Properties
%assigned to clildren nodes limit access and propagation. Access
%permission to particular information node is evaluated by traversing
%all nodes prom parent to child. At first node which gives negative
%result evaluation is stopped.
%
% Because evaluation of permission may be costly (CPU intensive,
%causing big latency if access to external service is needed, etc.)
%other tricks to reduce cost are needed.
%
% To reduce access control evaluation cost ACRs may/should/must(?)
%be assigned to information nodes in one-to-many way - one ACR per
%multiple information nodes. This grouping may be initiated either
%interactively (by clients explicitely specifying that same ACR is
%applicable to multiple nodes) or automatically (by comapring new
%ACR to already existing).
%
% If information nodes containing ACRs are propagated to ISIS (second
%implementation step) propagation algorithm must take ACR into account.
%For informtion propagation between ISISes this rule also applies.
%EXAMPLE: Information with access open to VO_1 must not be propagated
%to ISIS belonging to VO_2.
%
%
%** Ideas for in-memory XML documents
%
% Because informational documents are normally produced dynamically and
%have short lifetime there seems to be reasonable to keep them in memory.
% Each node with assigned ACR contains attribute (even if that breaks schema)
%containing id of ACR. ACR is defined as XML element (even if that breaks
%schema) located at higher, same or child level of considered node.
% Because informational documents are produced dynamically there are few
%ways for processing ACRs:
% 1. Right after document is generated it is traversed and ACRs applied.
%Parts of document which do not pass ACR evaluation are cut off. ACRs are
%cut off too.
% 2. While calling document generation procedure identity of client is
%supplied and ACRs are applied immediately.
% 3. While calling document generation procedure hooks/callbacks are supplied
%which are used and implement ACR evaluation.
% Combination of 1 and 3 looks most promising because it is flexible and
%keeps information system and policy evaluation code separated.
%
%
%* References
%"Information access control use cases" http://wiki.nordugrid.org/index.php/ARC1/Infosys/Access_Use_cases
%

% subsection information_authorization (end)

% section authorization (end)

% chapter isis (end)

\chapter{Service}
\label{cha:service}

\section{Information generation}
\label{sec:service_information_generation}

The service developers have to ensure that the services are providing the necessary information about themselves. This is done by implementing the subclass of the Arc::Service class - the RegistrationCollector function has to provide up-to-date status information about the service and anything else it wants to be advertised. This information package is called \textbf{Service Advertisement}.
\label{service_advertisement}
The \textbf{Service Advertisement} can contain any information service wants to advertise but the mandatory elements have to be always present:
\begin{itemize}
  \item Service ID: A globally unique identifier of the service.
  \item Service Type: The Glue2 type of service.
  \item Endpoint URL: The URL where the service can be contacted provided as part of EPR element.
\end{itemize}

Because there may be multiple registration processes running in parallel it is important to ensure that implementation RegistrationCollector is thread safe or there are internal locks implemented.

% section information (end)

\section{Registration}
\label{sec:service_registration}

The registration of service is carried out by internal module called Registrant. The Registrant is active module of the HED (Hosting Environment Daemon) which is bound to a set of ISISes. In practice, the configuration part of the Registrant contains exactly one ISIS to bind, and the Registrant will collect the necessary information about the other ISISes belonging to the same network.

To register services to more than one ISIS network multiple Registrant instances has to be configured. In this case, the default Registrant will be used for registering every services unless configured expilicitely.
The registration of service can be done once or periodically based either on the configuration of the Registrant or overwritten for every service separately. The Registrant is also performing message aggregation of all services linked to it if possible. The simplified algorithm of the Registrant is presented below.

\begin{framed}
  Registrant - pseudo algorithm\\
  \\
  \verb#// Initialize phase#\\
  \verb#Read the configuration and store the information about the services in a list#\\
  \verb#do { // Cyclic phase in a different Thread#\\
  \verb# wake_up_time = now();#\\
  \verb# messages = null;#\\
  \verb#  if ( 0 < count(service where service.next_run <= wake_up_time)) {#\\
  \verb#    foreach( service where service.next_run <= wake_up_time) {#\\
  \verb#      messages.add(service.RegistrationCollector);#\\
  \verb#      service.next_run = wake_up_time + service.period;#\\
  \verb#    }#\\
  \verb#    if (0 < count(messages)) {#\\
  \verb#      sent_message = assemble message with headers(messages);#\\
  \verb#      send(sent_message);#\\
  \verb#    }#\\
  \verb#  } else {#\\
  \verb#    sleep(min(service.next_run) - now()); #\\
  \verb#  }#\\
  \verb#} while(true)#\\
\end{framed}
Current implementation does not allow value of the period to be less than 2 minutes.

Service provides \textbf{Service Advertisement} part of information sent to ISIS (see Section \ref{sub:isis_interface}).
Before sending this information the Registrant extends it with additional data (\textbf{Service Advertisement Metadata}).

An example layout of services is shown on Figure \ref{fig:registration_overview}. In this configuration example 
there are two Registrants configured in one HED container for three services. The \textit{Registrant A} is the 
default Registrant and the services configured in the following way:

\begin{itemize}
  \item Service 1: There is no Registrant configured so the default one will register it.
  \item Service 2: The \textit{Registrant A} is configured explicitly.
  \item Service 3: The \textit{Registrant B} is configured explicitly.
\end{itemize}

So the Service 1 and 2 are handled both by \textit{Registrant A} and Service 3 by \textit{Registrant B}.
In the first step each Registrant performs information collection from all assigned services sequentially. Registrants themselves are executing in parallel. For Figure \ref{fig:registration_overview} the approximate sequence of information collection is (\textit{A1}-\textit{A2}, \textit{B1}-\textit{B2}) followed by (\textit{A3}, \textit{A4}). After collection  the aggregated information is sent to the corresponding ISISes independently by each other Registrant (\textit{A5}, \textit{B5}).

\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{Registration.png}}}
\caption{\label{fig:registration_overview}Overview of the registration process} }
\end{figure}

This registration operation is done once during the start-up phase and periodically according to (per service) configured periods.

% section registration (end)

\section{Configuration}
\label{sec:service_configuration}

% The registration operation is done by the Registrants and described in Section \ref{service_registrantion}.

% section configuration (end)

% chapter service (end)

% \bibliography{grid}
\end{document}