### 2009-05-19 doc/manuals/janitor/images/dependencies.tex
Temporary solution to save the nightlies - use old pgf style

\begin{tikzpicture}[>=latex,join=bevel,]
  \pgfsetlinewidth{1bp}
%%
\pgfsetcolor{black}
  % Edge: Installer -> Util
  \draw [->] (649bp,144bp) .. controls (648bp,136bp) and (648bp,127bp)  .. (647bp,108bp);
  % Edge: Janitor -> Filefetcher
  \draw [->] (670bp,221bp) .. controls (687bp,210bp) and (714bp,193bp)  .. (742bp,175bp);
  % Edge: Catalog -> Tag
  \draw [->] (352bp,152bp) .. controls (376bp,141bp) and (416bp,124bp)  .. (449bp,108bp) .. controls (451bp,107bp) and (452bp,107bp)  .. (463bp,101bp);
  % Edge: BaseSystem -> Note
  \draw [->] (410bp,72bp) .. controls (410bp,64bp) and (410bp,55bp)  .. (410bp,36bp);
  % Edge: Installer -> Execute
  \draw [->] (665bp,146bp) .. controls (674bp,136bp) and (686bp,123bp)  .. (703bp,105bp);
  % Edge: Janitor -> Catalog
  \draw [->] (625bp,228bp) .. controls (568bp,216bp) and (430bp,184bp)  .. (355bp,168bp);
  % Edge: Catalog -> MetaPackage
  \draw [->] (304bp,155bp) .. controls (279bp,146bp) and (241bp,131bp)  .. (215bp,108bp) .. controls (195bp,90bp) and (180bp,65bp)  .. (166bp,36bp);
  % Edge: Janitor -> TarPackage
  \draw [->] (624bp,231bp) .. controls (564bp,225bp) and (415bp,208bp)  .. (293bp,180bp) .. controls (191bp,156bp) and (166bp,146bp)  .. (67bp,108bp) .. controls (65bp,107bp) and (63bp,107bp)  .. (52bp,101bp);
  % Edge: Filefetcher -> Execute
  \draw [->] (751bp,145bp) .. controls (746bp,137bp) and (739bp,126bp)  .. (728bp,107bp);
  % Edge: Janitor -> Config
  \draw [->] (678bp,231bp) .. controls (731bp,226bp) and (845bp,210bp)  .. (869bp,180bp) .. controls (884bp,160bp) and (869bp,132bp)  .. (848bp,105bp);
  % Edge: Janitor -> Execute
  \draw [->] (663bp,218bp) .. controls (671bp,207bp) and (680bp,193bp)  .. (687bp,180bp) .. controls (696bp,160bp) and (704bp,136bp)  .. (712bp,108bp);
  % Edge: Catalog -> TarPackage
  \draw [->] (302bp,159bp) .. controls (255bp,152bp) and (151bp,135bp)  .. (67bp,108bp) .. controls (65bp,107bp) and (63bp,106bp)  .. (51bp,102bp);
  % Edge: Catalog -> DebianPackage
  \draw [->] (355bp,156bp) .. controls (392bp,146bp) and (463bp,128bp)  .. (521bp,108bp) .. controls (523bp,107bp) and (526bp,106bp)  .. (538bp,102bp);
  % Edge: Installer -> Config
  \draw [->] (674bp,153bp) .. controls (707bp,140bp) and (764bp,117bp)  .. (810bp,99bp);
  % Edge: Tag -> Note
  \draw [->] (469bp,75bp) .. controls (458bp,65bp) and (444bp,51bp)  .. (425bp,33bp);
  % Edge: MetaPackageCollection -> MetaPackage
  \draw [->] (131bp,72bp) .. controls (135bp,64bp) and (140bp,54bp)  .. (150bp,36bp);
  % Edge: Common -> Config
  \draw [->] (833bp,144bp) .. controls (833bp,136bp) and (833bp,127bp)  .. (833bp,108bp);
  % Edge: Janitor -> DebianPackage
  \draw [->] (636bp,219bp) .. controls (627bp,208bp) and (615bp,194bp)  .. (606bp,180bp) .. controls (593bp,160bp) and (582bp,135bp)  .. (571bp,108bp);
  % Edge: Catalog -> Package
  \draw [->] (313bp,147bp) .. controls (302bp,137bp) and (287bp,123bp)  .. (267bp,105bp);
  % Edge: Catalog -> NoteCollection
  \draw [->] (329bp,144bp) .. controls (329bp,136bp) and (329bp,127bp)  .. (329bp,108bp);
  % Edge: Janitor -> Installer
  \draw [->] (651bp,216bp) .. controls (651bp,208bp) and (651bp,199bp)  .. (651bp,180bp);
  % Edge: Janitor -> Util
  \draw [->] (637bp,219bp) .. controls (629bp,208bp) and (619bp,194bp)  .. (615bp,180bp) .. controls (610bp,164bp) and (610bp,159bp)  .. (615bp,144bp) .. controls (617bp,134bp) and (622bp,124bp)  .. (633bp,106bp);
  % Edge: Execute -> InstallLogger
  \draw [->] (717bp,72bp) .. controls (717bp,64bp) and (717bp,55bp)  .. (717bp,36bp);
  % Edge: Janitor -> Common
  \draw [->] (675bp,226bp) .. controls (704bp,216bp) and (755bp,198bp)  .. (797bp,180bp) .. controls (799bp,179bp) and (800bp,179bp)  .. (811bp,173bp);
  % Edge: NoteCollection -> Note
  \draw [->] (347bp,74bp) .. controls (358bp,64bp) and (373bp,51bp)  .. (393bp,33bp);
  % Edge: Catalog -> BaseSystem
  \draw [->] (345bp,147bp) .. controls (357bp,137bp) and (372bp,123bp)  .. (393bp,105bp);
  % Edge: Catalog -> MetaPackageCollection
  \draw [->] (305bp,154bp) .. controls (271bp,142bp) and (208bp,120bp)  .. (156bp,102bp);
  % Node: Note
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (410bp,18bp) ellipse (27bp and 18bp);
  \draw (410bp,18bp) node {Note};
\end{scope}
  % Node: Execute
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (717bp,90bp) ellipse (27bp and 18bp);
  \draw (717bp,90bp) node {Execute};
\end{scope}
  % Node: Package
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (251bp,90bp) ellipse (27bp and 18bp);
  \draw (251bp,90bp) node {Package};
\end{scope}
  % Node: DebianPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (565bp,90bp) ellipse (35bp and 18bp);
  \draw (565bp,90bp) node {DebianPackage};
\end{scope}
  % Node: Config
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (833bp,90bp) ellipse (27bp and 18bp);
  \draw (833bp,90bp) node {Config};
\end{scope}
  % Node: MetaPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (159bp,18bp) ellipse (32bp and 18bp);
  \draw (159bp,18bp) node {MetaPackage};
\end{scope}
  % Node: NoteCollection
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (329bp,90bp) ellipse (33bp and 18bp);
  \draw (329bp,90bp) node {NoteCollection};
\end{scope}
  % Node: Util
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (645bp,90bp) ellipse (27bp and 18bp);
  \draw (645bp,90bp) node {Util};
\end{scope}
  % Node: Catalog
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (329bp,162bp) ellipse (27bp and 18bp);
  \draw (329bp,162bp) node {Catalog};
\end{scope}
  % Node: Tag
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (485bp,90bp) ellipse (27bp and 18bp);
  \draw (485bp,90bp) node {Tag};
\end{scope}
  % Node: Filefetcher
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (761bp,162bp) ellipse (27bp and 18bp);
  \draw (761bp,162bp) node {Filefetcher};
\end{scope}
  % Node: Common
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (833bp,162bp) ellipse (27bp and 18bp);
  \draw (833bp,162bp) node {Common};
\end{scope}
  % Node: MetaPackageCollection
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (122bp,90bp) ellipse (46bp and 18bp);
  \draw (122bp,90bp) node {MetaPackageCollection};
\end{scope}
  % Node: Installer
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (651bp,162bp) ellipse (27bp and 18bp);
  \draw (651bp,162bp) node {Installer};
\end{scope}
  % Node: Janitor
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (651bp,234bp) ellipse (27bp and 18bp);
  \draw (651bp,234bp) node {Janitor};
\end{scope}
  % Node: TarPackage
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (29bp,90bp) ellipse (29bp and 18bp);
  \draw (29bp,90bp) node {TarPackage};
\end{scope}
  % Node: BaseSystem
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (410bp,90bp) ellipse (30bp and 18bp);
  \draw (410bp,90bp) node {BaseSystem};
\end{scope}
  % Node: InstallLogger
\begin{scope}
  \pgfsetstrokecolor{black}
  \draw (717bp,18bp) ellipse (33bp and 18bp);
  \draw (717bp,18bp) node {InstallLogger};
\end{scope}
%
\end{tikzpicture}