### 2009-06-08 doc/manuals/janitor/tex_maintenance/maintenance.tex

\chapter{Maintenance}

This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
It gives detailed instructions how to create new packages for the Janitor.
To administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e,
as explained in the next section.
In the last section, a typical use case in maintaining the Janitor will be shown.

% * Maintainance
% ** Protege and RDF files
% ** Creating own tar balls

\section{Catalog}\label{sec:catalog}

The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catalog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Metapackage.}{}
         \label{tab:main_class_meta}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	homepage       & single                & String       \\
	instance       & multiple              & Instance of Package       \\
	lastupdated    & single                & String       \\
	name           & required single       & String       \\
	tag            & multiple              & Instance of Tag       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Basesystem.}{}
         \label{tab:main_class_base}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	distribution   & required single       & String       \\
	name           & required single       & String       \\
	short\_description & required single   & String       \\
	url            & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Tag.}{}
         \label{tab:main_class_tag}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	name           & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class DebianPackage.}{}
         \label{tab:main_class_debian}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	debconf        & multiple              & String       \\
	depends        & multiple              & Instance of MetaPackage  or Package      \\
	package        & required multiple              & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class TarPackage.}{}
         \label{tab:main_class_tar}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	depends        & multiple              & Instance of MetaPackage or Package      \\
	environ        & multiple              & String       \\
	url            & required multiple     & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\section{HTML interface of the catalog}

The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and its automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
This script is meant to be run by a mod\_perl-enabled Apache. The script itself does not contribute to the 
core functionality of the Janitor.
In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

\section{Introducing new packages}\label{sec:catalog}

This section describes how to add new packages to the Catalog. In the current implementation, only tar based packages are processed by the Janitor.
Within the example they are assigned to be used together with Debian Etch. This limitation is only literal. 
There is no limitiation to newer Debian distributions.

\subsection{Debian Etch (tar based)}

At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
a well accepted file format throughout the UNIX community. The reflects the traditional manual approach towards RTE in ARC,
for which one directory is made available to all compute nodes. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

\begin{figure}
  \begin{center}
    \includegraphics[width=4cm]{images/tar_folder.pdf}
    \mycaption{Directory structure in the tar files for automated installation.}{}
    \label{fig:tar_folder}
  \end{center}
\end{figure}

The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
Once the tar file was prepared, it must its entry to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.

From such Catalogues, the Janitor finds all information to install packages that possibly have never been installed
on the site before. The offers of RTEs in a catalogue are cross-checked against the local infrastructure and a subset
of the available packages will be accepted as "installable". This list of installable RTEs is forwarded to the
grid information system.

\subsection{Protoypes}

In order to have an impression how the tar files are created, several
prototypes are provided at \url{http://dre.knowarc.eu}.

The WEKA package for machine learning~\cite{FRANK_20074} and the Java Runtime Environment
are available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
installation via the prior mentioned traditional page representing Runtime Environments for ARC. 
The corresponding tar file is named \task{somewhat??}. The \texttt{data} directory simply contains a ZIP file
which needs to be unzipped in the installation directory. For that reason, the \texttt{control/install} script is 
written as follows:
% dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
\begin{verbatim} 
#!/bin/sh
set -e  # Makes the script to terminate at the first line it fails.

WEKA_ZIP="weka-3-4-8a.zip"
unzip $WEKA_ZIP
rm -f $WEKA_ZIP
\end{verbatim}
The runtime script sets the environment variable of the Java Classpath:
\begin{verbatim}
#!/bin/sh

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}
The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
is nothing more to be done.\\

%\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
%provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
%dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
%thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
%projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

To address the concerns of the physicists using ARC, a dynamic runtime environment for the
ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}

What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
5 GB. This was a test illustrating the feasibility of using dynamic REs in High Energy Physics. The application
of the dynamic RE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
service, e.g. a software manager of a big experiment will be able to deploy software packages on production
sites simply by creating a tarball and adding an entry to the Catalog.

%\subsection{Debian Etch}
%
%These kind of packages are not yet supported.
%
%\subsection{Debian Sid}
%
%These kind of packages are not yet supported.
%
%\subsection{Adminstrating the Catalog}

\section{Typical use cases}\label{sec:catalog}

% ** setstate REMOVAL_PENDING
% no warranty interferences in the file syste,