### 2009-09-18 doc/manuals/python/app/ex-2-1-2.tex

\section{Examples of \ref{icommon}}
\subsection{String operator - simple node}
(Example \ref{strex1})
\label{cstrex1}
\begin{verbatim}
import arc
#create an XMLNode
mynode = arc.XMLNode(arc.NS({'me':'http://example.com/myExample'}),'me:myNode')
#set text content
mynode.Set('Hello, World!')
#representation of the node
mynode.GetXML()
#String operator in action
str(mynode)
\end{verbatim}

\subsection{String operator - complex node}
(Example \ref{strex2})
\label{cstrex2}
\begin{verbatim}
# String operator of XMLNode works with the text content.
# It does not care about child nodes or text contents of those nodes.

# build a tree and set text content in nodes
#       r
#     / | \
#    /  |  \
#   /   |   \
# c01  c02  c03
#      /\    |
#     /  \  c06  
#   c04  c05

# create the root node
r = arc.XMLNode(arc.NS(),'r')
# set text content
r.Set('R')
# create rest of the tree
c01 = r.NewChild('c01')
c01.Set('C-01')
c02 = r.NewChild('c02')
c02.Set('C-02')
c03 = r.NewChild('c03')
c03.Set('C-03')
c04 = c02.NewChild('c04')
c04.Set('C-04')
c05 = c02.NewChild('c05')
c05.Set('C-05')
c06 = c03.NewChild('c06')
c06.Set('C-06')
# String operator does not care about child nodes
str(c03)
str(c02)
str(r)
# XML representation of a complex node
c02.GetXML()
\end{verbatim}

\subsection{Occurence of out\_xml\_str as an output value}
(Example \ref{outxml})
\label{coutxml}
\begin{verbatim}
import arc
#create node
n = arc.XMLNode(arc.NS({'me':'myNS'}),'me:myNode')
#set content
n.Set('Hello, World!')
#get document
#GetDoc according to API:
#void GetDoc(std::string &out_xml_str, bool user_friendly=false) const
#out_xml_str is turned into an output value and user_friendly is an optional 
#parameter so there are no necessary parameters this time
#result will hold the output value
result = n.GetDoc()
#show result
result
\end{verbatim}

\subsection{Using LogStream to add new destination to the root logger}
(Example \ref{logstream})
\label{clogstream}
\begin{verbatim}
import arc
import sys
#get root logger
root_logger = arc.Logger_getRootLogger()
#create a LogStream; sys.stdout would be OK
stream = arc.LogStream(sys.stdout)
#add destination to root logger
root_logger.addDestination(stream)
#log a message
#result immediately appears on sys.stdout
root_logger.msg(arc.INFO, 'Hello, World!')
\end{verbatim}