### 2009-09-18 doc/manuals/python/app/ex-2-1-4.tex

\section{Examples of \ref{iclient}}
\subsection{ClientSOAP - process}
(Example \ref{procex})
\label{cprocex}
\begin{verbatim}
import arc
#create default config
cfg = arc.MCCConfig()
#create URL
url = arc.URL('http://localhost:50000/Echo')
#create payload
payload = arc.PayloadSOAP(arc.NS({'echo':'urn:echo'}))
#create payload content
#and set echo message
payload.NewChild('echo:echo').NewChild('echo:say').Set('Hello, World!')
#create client
client = arc.ClientSOAP(cfg,url)
#let the client call the service
response, status = client.process(payload)
#see response
print response.GetXML()
#check status
status.getKind() == arc.STATUS_OK
\end{verbatim}