### 2009-09-18 doc/manuals/python/app/ex-2-1-5.tex

\section{Examples of \ref{idata}}
\subsection{DataMover - Transfer}
(Example \ref{dmex})
\label{cdmex}
\begin{verbatim}
import arc
# create DataMover
mover = arc.DataMover()
# import datapoint_from_url from arcom package
# DataPoints are required to use DataMover
from arcom import datapoint_from_url
# create source DataPoint
src = datapoint_from_url('file://./file1')
# create destination DataPoint
dst = datapoint_from_url('file://./file2')
# do the transfer
status, failure = mover.Transfer(src, dst, arc.FileCache(), arc.URLMap())
# check status
str(status)
'Operation completed successfully'
# check for failure message
failure
\end{verbatim}