### 2009-09-18 doc/manuals/python/app/ex-3-7.tex

\section{Examples of \ref{axtree}}
\subsection{Creating an XMLTree}
(Example \ref{crxt})
\label{ccrxt}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# create XMLTree sources
# xmlstr - will be used to create an XMLTree from string
# xmlnodestr - will be used to create 'node' (an XMLNode)
# node - will be used to create an XMLTree from XMLNode
# treestruct - for creating XMLTree from tree structure
xmlstr = '<echo><say>Hello!</say></echo>'
xmlnodestr = '<node><subnode>Subnode</subnode></node>'
node = XMLNode(xmlnodestr)
treestruct = ('root', [('leaf','Leaf')])

# from_tree has the highest priority
t1 = XMLTree(from_node = node, from_string = xmlstr, from_tree = treestruct) 
t1.get()

# no from_tree supplied; from_node will be used
t2 = XMLTree(from_node = node, from_string = xmlstr)
t2.get()

# from_string only
t3 = XMLTree(from_string = xmlstr)
t3.get()
\end{verbatim}

\subsection{XMLTree - forget namespace}
(Example \ref{xtfn})
\label{cxtfn}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# Create XMLTree objects to demonstrate
# the use of forget namespace feature
# node - is an XMLNode
# tree - is an XMLTree that is created from 'node'
node = XMLNode('<soap-env:Envelope xmlns:hash="urn:hash" \
xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" \
xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" \
xmlns:xsd="http://www.w3.org/2001/XMLSchema" \
xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">\
<soap-env:Body>\
<hash:get>\
<hash:IDs>\
<hash:ID>0</hash:ID>\
<hash:ID>1</hash:ID>\
<hash:ID>2</hash:ID>\
</hash:IDs>\
</hash:get>\
</soap-env:Body>\
</soap-env:Envelope>')

tree = XMLTree(node)
# show tree structure
print tree

# forget_namespace is ignored when from_tree is used
tree2 = XMLTree(from_tree = tree, forget_namespace = True)
print tree2

# forget_namespace with from_node
tree3 = XMLTree(from_node = node, forget_namespace = True)
# it works as expected
print tree3

# forget_namespace with from_string
strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
</myns:foo>'
tree4 = XMLTree(from_string = strnode, forget_namespace = True)
# it also works
print tree4
\end{verbatim}

\subsection{XMLTree - rewrite}
(Example \ref{xtrew})
\label{cxtrew}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

# Create XMLTree objects to demonstrate
# the use of rewrite feature
strnode = '<myns:foo xmlns:myns="urn:foo"><myns:dummy>Hello, World!</myns:dummy>\
</myns:foo>'
tree = XMLTree(from_string = strnode,\
               rewrite = {'myns:foo':'myns:boo'},\
               forget_namespace = False)

# 'myns:foo' is turned into 'myns:boo'
print tree

# combine with forget_namespace
tree2 = XMLTree(from_string = strnode,\
                rewrite = {'foo':'boo'},\
                forget_namespace = True)
# 'forget_namespace' is True so namespace prefix 'myns' is being stripped off
# ('myns:foo' -> 'foo'; 'myns:dummy' -> 'dummy')
# then 'foo' is turned into 'boo' 
print tree2
\end{verbatim}

\subsection{XMLTree - add\_to\_node}
(Example \ref{xtadd2n})
\label{cxtadd2n}
\begin{verbatim}
from arc import XMLNode
from arcom.xmltree import XMLTree

node = XMLNode('<node/>')

# create XMLTree
tree = XMLTree(from_string = '<root><dummy/><hello>World</hello></root>')

# from 'tree' add 'dummy' to 'node'
tree.add_to_node(node, '/root/dummy')

# show result
print node.GetXML(True)

# from 'tree' add 'hello' to 'dummy' in 'node'
dummy = node.Get('dummy')
tree.add_to_node(dummy, '/root/hello')

# show result
print node.GetXML(True)
\end{verbatim}

\subsection{XMLTree - pretty\_xml}
(Example \ref{xtpretty})
\label{cxtpretty}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create XMLTree
tree = XMLTree(from_string = '<root><Hello>World</Hello></root>')

# select entire tree
# indent children with space character
# apply prefix for every new line
print tree.pretty_xml(indent=' ', path='', prefix='# ')

# note that it is possible to select multiple parts
# create an example XMLTree to demonstrate that
tree2 = XMLTree(from_string = '<root><sub><Hello>World</Hello></sub>\
<sub><dummy>foo</dummy></sub></root>')

# show '/root/sub'
print tree2.pretty_xml(indent=' ', path='/root/sub', prefix='# ')
\end{verbatim}

\subsection{XMLTree - \_\_str\_\_}
(Example \ref{xtstr})
\label{cxtstr}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create XMLTree
tree = XMLTree(from_string = '<root><dummy>foo</dummy></root>')

# show tree
# print will result in __str__ being called
# when trying to convert XMLTree to string
print tree
\end{verbatim}

\subsection{XMLTree - get}
(Example \ref{xtget})
\label{cxtget}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
tree = \
('root',\
  [('trunk',\
    [('leaf','1'),\
     ('leaf','2')]\
  )]\
)

# create XMLTree from tree structure
xt = XMLTree(from_tree = tree)

# if called with path being None
# get() selects the root node
xt.get()

# Path is just a plain path.
# Empty tag name matches everything, so - in this particular case -
# all expressions below will produce the same result.
xt.get('/root/trunk/leaf')
xt.get('//trunk/leaf')
xt.get('/root//leaf')
xt.get('/root/trunk/')
xt.get('///leaf')
xt.get('///')
\end{verbatim}

\subsection{XMLTree - get\_trees}
(Example \ref{xtgettrees})
\label{cxtgettrees}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
tree = \
('trunk',\
  [( 'branch',\
     [('leaf','1'),\
      ('leaf','2')]\
   ),\
   ( 'branch',\
     [('leaf','3')]\
   )]\
)

# create XMLTree from tree structure
xt = XMLTree(from_tree = tree)

# get branches ('/trunk/branch') in a list
blist = xt.get_trees('/trunk/branch')

# show branches in blist
for b in blist:
    print b

# get leaves ('/trunk/branch/leaf') in a list
llist = xt.get_trees('/trunk/branch/leaf')

# show leaves in llist
for l in llist:
    print l

\end{verbatim}

\subsection{XMLTree - get\_value}
(Example \ref{xtgetval})
\label{cxtgetval}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = ('branch',[('leaf','1')])

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get_value for '/branch/leaf'
# will return the value from 'leaf'
tree.get_value('/branch/leaf')

# get_value for '/branch'
# will return the value from 'branch'
# ie the list of its children
tree.get_value('/branch')

# get_value for a path that does not match
# provide a default value of 'N/A'
tree.get_value('/dummy','N/A')
\end{verbatim}

\subsection{XMLTree - add\_tree}
(Example \ref{xtaddtree})
\label{cxtaddtree}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = \
('root',\
  [ ('branch',\
      [ ('leaf','1'),\
        ('leaf','2')\
      ]\
    ),\
    ('branch',\
      [ ('leaf','3'),\
        ('leaf','4')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# add a new leaf ('leaf','5') to '/root/branch'
# the new leaf will be added to the first node 
# matched by path ('/root/branch') 
tree.add_tree(('leaf','5'), '/root/branch')

# show tree
print tree
\end{verbatim}

\subsection{XMLTree - get\_values}
(Example \ref{xtgetvals})
\label{cxtgetvals}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts = \
('root',\
  [ ('branch',\
      [ ('leaf','1'),\
        ('leaf','2')\
      ]\
    ),\
    ('branch',\
      [ ('leaf','3'),\
        ('leaf','4')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get values of branches ('/root/branch') in a list
# this will return the lists of children of the branches
tree.get_values('/root/branch')

# get values of leaves ('/root/branch/leaf') in a list
# this will return the values of leaves
tree.get_values('/root/branch/leaf')

# try a path that does not match
# this will return an empty list
tree.get_values('/dummy')
\end{verbatim}

\subsection{XMLTree - get\_dict}
(Example \ref{xtgetdict})
\label{cxtgetdict}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts =\
('person',\
  [ ('name', 'Alice'),\
    ('id', '11'),\
    ('job', 'Librarian')\
  ]
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get dictionary from '/person'
# only name and job is to be returned
tree.get_dict('/person',{'name':'name','job':'job'})
\end{verbatim}

\subsection{XMLTree - get\_dicts}
(Example \ref{xtgetdicts})
\label{cxtgetdicts}
\begin{verbatim}
from arcom.xmltree import XMLTree

# create tree structure
ts =\
('persons',\
  [ ('person',\
      [ ('name', 'Alice'),\
        ('id', '11'),\
        ('job', 'Librarian')\
      ]\
    ),\
    ('person',\
      [ ('name', 'Bob'),\
        ('id', '12'),\
        ('job', 'Bartender')\
      ]\
    )\
  ]\
)

# create XMLTree from tree structure
tree = XMLTree(from_tree = ts)

# get dictionaries for '/persons/person'
tree.get_dicts('/persons/person')
\end{verbatim}