### 2009-10-15 doc/tech_doc/client/job_description_mapping.tex
Converting document to LaTeX. Please port to final.

\documentclass{article}
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{array}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage[english]{babel}
\usepackage{multirow}

\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}

\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}

\newcommand{\todo}{TODO}

\newcommand{\same}{{\textquotesingle}{\textquotesingle}}
\newcommand{\subsubsubsection}[1]{\paragraph{#1}}
\newcommand{\subsubsubsubsection}[1]{\subparagraph{#1}}

\newenvironment{inouttabular}%
{\begin{center}\begin{tabular}{l>{\ttfamily\footnotesize}c>{\ttfamily\footnotesize}c}%
\toprule
& \textnormal{\normalsize{In}} & \textnormal{\normalsize{Out}}\\ \cmidrule{2-3}}
{\bottomrule\end{tabular}\end{center}}

\makeatletter
\newcommand\arraybslash{\let\\\@arraycr}
\makeatother

%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

\hrulefill

%-------- Change this to NORDUGRID-XXXXXXX-NN

\raggedleft {NORDUGRID-XXXXXXX-NN\par}

{\raggedleft \today\par}

\vspace*{2cm}

%%%%---- The title ----
{\centering \textsc{\Large The ARC Job Description}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Internal Representation Mapping}\large \par}

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large M. S. Andersen\footnote{skou@nbi.dk} \large \par}

%%%%---- An abstract - if style is article ----
%\begin{abstract}
%The abstract
%\end{abstract}
\end{titlepage}

\tableofcontents                          %Comment if use article style
\newpage

\section{Introduction}
This document contains information about how XRSL, JDL and the various
supported JSDL flavours are mapped to and from the job description
internal representation in ARC{}-lib.

\section{Data Structure}
The internal representation of the job description is contained in the
JobDescription class, which consist of the following objects:
JobIdentification, Application, Resources, DataStaging and JobMetaData.
JobIdentification contains data used to identify the job description
through various types, tags and names. The Application object is used
to describe explicitly the executable which should be executed at the
Computing Endpoint (CE), environment variables, logging, standard in
and out, credential service, etc. The Resources class contains
information about requirements and the type of execution node preferred
by the user. The DataStaging class contains information about input
files and output files/directories created by the job. Finally the job
description may contain meta data, describing the context in which the
description was generated, and it is stored in the JobMetaData class.

\subsection{JobIdentification}
\subsubsection{JobName}
\begin{table}[h]
\begin{inouttabular}
XRSL & jobname & jobname\\
JDL & No matching attribute & Not mapped\\
JSDL & JobIdentification/JobName & JobIdentification/JobName\\
ARC-JSDL &  & \\
\end{inouttabular}
\end{table}

\subsubsection{Description}
\begin{table}[h]
\begin{inouttabular}
ARC-JSDL & JobIdentification/Description & JobIdentification/Description\\
\end{inouttabular}
\end{table}

\subsubsection{JobType}
\begin{inouttabular}
ARC-JSDL & JobIdentification/JobType & JobIdentification/JobType\\
\end{inouttabular}
JDL has a jobtype and a type attribute, however a mapping has not been
made.

\subsubsection{UserTag}
\begin{inouttabular}
JDL & usertags & usertags\\
ARC-JSDL & JobIdentification/UserTag & JobIdentification/UserTag\\
\end{inouttabular}

\subsection{JobVOName}
\begin{inouttabular}
XRSL & jobname & jobname\\
ARC-JSDL & JobIdentification/JobVOName & JobIdentification/JobVOName\\
\end{inouttabular}

\subsection{Application}
\subsubsection{Executable}
\begin{inouttabular}
XRSL & executable, arguments & executable, arguments\\
JDL & Executable, Arguments & Executable, Arguments\\
\multirow{2}{*}{Posix-JSDL} & Application/POSIXApplication/Executable, & Application/POSIXApplication/Executable,\\
&  Application/POSIXApplication/Arguments & Application/POSIXApplication/Arguments\\
\multirow{2}{*}{HPCP-JSDL} & Application/HPCProfileApplication/Executable, & Application/HPCProfileApplication/Executable,\\
&  Application/HPCProfileApplication/Arguments & Application/HPCProfileApplication/Arguments\\
\multirow{2}{*}{ARC-JSDL} & Application/Executable/Path, & Application/Executable/Path,\\
 & Application/Executable/Argument & Application/Executable/Argument\\
\end{inouttabular}
The Executable data member is a composite structure of a string and a
list of strings, where the string holds the executable and the list
holding the arguments.

\subsubsection{Input}
\begin{inouttabular}
XRSL & stdin & stdin, inputFiles\\
JDL & StdInput & StdInput, InputSandbox\\
Posix-JSDL & Application/POSIXApplication/Input & Application/POSIXApplication/Input\\
HPCP-JSDL & Application/HPCProfileApplication/Input & Application/HPCProfileApplication/Input\\
ARC-JSDL & Application/Input & \\
\end{inouttabular}
If the Input data member does not exist in the file list, it will be
added to it.

\subsubsection{Output}
\begin{inouttabular}
XRSL & stdout & stdout, outputFiles\\
JDL & StdOutput & StdOutput, OutputSandbox, OutputSandboxDestURI\\
Posix-JSDL & Application/POSIXApplication/Output & \\
HPCP-JSDL & Application/HPCProfileApplication/Output & \\
ARC-JSDL & Application/Output & \\
\end{inouttabular}
If the Output data member does not exist in the File
list, it will be added to it.

\subsubsection{Error}
\begin{inouttabular}
XRSL & stderr & stderr, outputFiles\\
JDL & StdError & StdError, OutputSandbox, OutputSandboxDestURI\\
Posix-JSDL & Application/POSIXApplication/Error &\\
HPCP-JSDL & Application/HPCProfileApplication/Error &\\
ARC-JSDL & Application/Error & \\
\end{inouttabular}
If the Error data member does not exist in the File list, it will be
added to it.

\subsubsection{Join}
\begin{inouttabular}
XRSL & join & Not mapped\todo{}\\
ARC-JSDL & Application/Join & Application/Join\\
\end{inouttabular}

\subsubsection{Environment}
\subsubsubsection{Name}
\begin{inouttabular}
XRSL & environment* & environment*\\
JDL & Environment** & Environment**\\
Posix-JSDL & Application/POSIXApplication/Environment.name & Application/POSIXApplication/Environment.name\\
HPCP-JSDL & Application/HPCProfileApplication/Environment.name & Application/HPCProfileApplication/Environment.name\\
ARC-JSDL & Application/Environment/Name & ~\\
\end{inouttabular}
The XRSL attribute environment is a list of
pairs. The first element of each pair is mapped to the Name data member.
The JDL attribute Environment is a list of equalities.

\subsubsubsection{Value}
\begin{inouttabular}
XRSL & environment* & environment*\\
JDL & Environment** & Environment**\\
JSDL & No matching attribute & Not mapped\\
Posix-JSDL & Application/POSIXApplication/Environment & Application/POSIXApplication/Environment\\
HPCP-JSDL & Application/HPCProfileApplication/Environment & Applicaion/HPCProfileApplication/Environment\\
ARC-JSDL & Application/Environment/Value & ~\\
\end{inouttabular}

\subsubsection{Prologue}
\begin{inouttabular}
JDL & Prologue, PrologueArgument & Prologue, PrologueArgument\\
\multirow{2}{*}{ARC-JSDL} & Application/Prologue/Path, & Application/Prologue/Path,\\
& Application/Prologue/Argument & Application/Prologue/Argument\\
\end{inouttabular}

\subsubsection{Epilogue}
\begin{inouttabular}
JDL & Epilogue, EpilogueArgument & Epilogue, EpilogueArgument\\
\multirow{2}{*}{ARC-JSDL} & Application/Epilogue/Path, & Application/Epilogue/Path,\\
& Application/Epilogue/Argument & Application/Epilogue/Argument\\
\end{inouttabular}

\subsubsection{LogDir}
\begin{inouttabular}
XRSL & gmlog & gmlog outputFiles\\
ARC-JSDL & Application/LogDir & Application/LogDir\\
\end{inouttabular}

\subsection{RemoteLogging}
\begin{inouttabular}
XRSL & jobreport & jobreport\\
ARC-JSDL & Application/RemoteLogging & Application/RemoteLogging\\
\end{inouttabular}

\subsubsection{Rerun}
\begin{inouttabular}
XRSL & rerun & rerun\\
JDL & RetryCount ShallowRetryCount* & RetryCount ShallowRetryCount**\\
ARC-JSDL & Application/Rerun & Application/Rerun\\
\end{inouttabular}
* Only the maximum value between RetryCount
and ShallowRetryCount is stored.

** Both RetryCount and ShallowRetryCount will be set if
Rerun is set, and they will both have the value of Rerun.

\subsubsection{ExpiryTime}
\begin{inouttabular}
JDL & ExpiryTime & ExpiryTime\\
ARC-JSDL & Application/ExpiryTime & Application/ExpiryTime\\
\end{inouttabular}

\subsubsection{StartOnHold}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{ProcessingStartTime}
\begin{inouttabular}
ARC-JSDL & Application/ProcessingStartTime & Application/ProcessingStartTime\\
\end{inouttabular}

\subsubsection{Notification}
\begin{inouttabular}
XRSL & notify & notify\\
ARC-JSDL & Application/Notification & Application/Notification\\
\end{inouttabular}

\subsubsection{CredentialService}
\begin{inouttabular}
XRSL & credentialserver & credentialserver\\
JDL & MyProxyServer & MyProxyServer\\
ARC-JSDL & Application/CredentialService & Application/CredentialService\\
\end{inouttabular}

\subsubsection{AccessControl}
\begin{inouttabular}
XRSL & acl & acl\\
ARC-JSDL & Application/AccessControl & Application/AccessControl\\
\end{inouttabular}

\subsection{Resources}
\subsubsection{OperatingSystem}
\begin{inouttabular}
XRSL & opsys & opsys\\
ARC-JSDL & Resources/OperatingSystem & Resources/OperatingSystem\\
\end{inouttabular}

\subsubsection{Platform}
\begin{inouttabular}
XRSL & architecture & architecture\\
JSDL & Resources/CPUAchitecture/CPUArchitectureName & ~\\
ARC-JSDL & Resources/Platform &~\\
\end{inouttabular}

\subsubsection{NetworkInfo}
\begin{inouttabular}
JSDL & Resources/IndividualNetworkBandwidth &\\
ARC-JSDL & Resources/NetworkInfo &\\
\end{inouttabular}

\subsubsection{NodeAccess}
\begin{inouttabular}
XRSL & nodeAccess & nodeAccess\\
ARC-JSDL & Resources/NodeAccess & ~\\
\end{inouttabular}

\subsubsection{IndividualPhysicalMemory}
\begin{inouttabular}
XRSL & memory & memory\\
Posix-JSDL & Application/POSIXApplication/MemoryLimit &\\
ARC-JSDL & Resources/IndividualPhysicalMemory &\\
\end{inouttabular}

\subsubsection{IndividualVirtualMemory}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/VirtualMemoryLimit &\\
ARC-JSDL & Resources/IndividualVirtualMemory &\\
\end{inouttabular}

\subsubsection{DiskSpaceRequirement}
\subsubsubsection{DiskSpace}
\begin{inouttabular}
XRSL & disk & disk\\
JSDL & Application/FileSystem/DiskSpace &\\
ARC-JSDL & Resources/DiskSpaceRequirement/DiskSpace &\\
\end{inouttabular}

\subsubsubsection{CacheDiskSpace}
\begin{inouttabular}
ARC-JSDL & Resources/CacheDiskSpace & Resources/CacheDiskSpace\\
\end{inouttabular}

\subsubsubsection{SessionDiskSpace}
\begin{inouttabular}
ARC-JSDL & Resources/SessionDiskSpace & Resources/SessionDiskSpace\\
\end{inouttabular}

\subsubsection{SessionLifetime}
\begin{inouttabular}
XRSL & lifeTime & lifeTime\\
ARC-JSDL & Resources/SessionLifeTime & Resources/SessionLifeTime\\
\end{inouttabular}

\subsubsection{SessionAccessMode}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{IndividualCPUTime}
\begin{inouttabular}
JSDL & Resources/IndividualCPUTime* &\\
ARC-JSDL & Resources/IndividualCPUTime &\\
\end{inouttabular}
* The JSDL element is mapped to the Range part of
IndividualCPUTime.

\subsubsection{TotalCPUTime}
\begin{inouttabular}
XRSL & cpuTime & cpuTime\\
JSDL & Resources/TotalCPUTime* &\\
Posix-JSDL & Application/POSIXApplication/CPUTimeLimit* &\\
ARC-JSDL & Resources/TotalCPUTime &\\
\end{inouttabular}
* The JSDL and POSIX-JSDL elements is mapped to the
Range part of TotalCPUTime.

\subsubsection{IndividualWallTime}
\begin{inouttabular}
ARC-JSDL & Resources/IndividualWallTime &\\
\end{inouttabular}

\subsubsection{TotalWallTime}
\begin{inouttabular}
XRSL & wallTime & wallTime\\
Posix-JSDL & Application/POSIXApplication/WallTimeLimit &\\
ARC-JSDL & Resources/TotalWallTime &\\
\end{inouttabular}

\subsubsection{Homogeneous}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{Benchmark}
\subsubsubsection{Type}
\begin{inouttabular}
\end{inouttabular}

\subsubsubsection{Value}
\begin{inouttabular}
\end{inouttabular}

\subsubsection{CEType}
\begin{inouttabular}
XRSL & middleware & middleware\\
ARC-JSDL & Application/CEType & Application/CEType\\
\end{inouttabular}

\subsubsection{SlotRequirement}
\subsubsubsection{NumberOfSlots}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/ProcessCountLimit &\\
ARC-JSDL & Resources/SlotRequirement/NumberOfSlots &\\
\end{inouttabular}

\subsubsubsection{ProcessPerHost}
\begin{inouttabular}
XRSL & count & count\\
JSDL & Resources/TotalCPUCount & \\
ARC-JSDL & Resources/SlotRequirement/ProcessPerHost &\\
\end{inouttabular}

\subsubsubsection{ThreadsPerProcesses}
\begin{inouttabular}
Posix-JSDL & Application/POSIXApplication/ThreadCountLimit &\\
ARC-JSDL & Resources/SlotRequirement/ThreadsPerProcesses &\\
\end{inouttabular}

\subsubsubsection{SPMDVariation}
\begin{inouttabular}
ARC-JSDL & Resources/SlotRequirement/SPMDVariation &\\
\end{inouttabular}

\subsubsection{CandidateTarget}
\subsubsubsection{EndPointURL}
\begin{inouttabular}
XRSL & cluster & Not mapped\\
JSDL & Resources/CandidateHosts/HostName &\\
ARC-JSDL & Resources/CandidateTarget/EndPointURL & Resources/CandidateTarget/EndPointURL\\
\end{inouttabular}

\subsubsubsection{QueueName}
\begin{inouttabular}
XRSL & queue & queue\\
JDL & QueueName & QueueName\\
ARC-JSDL & Resources/CandidateTarget/QueueName & Resources/CandidateTarget/QueueName\\
\end{inouttabular}

\subsubsection{RunTimeEnvironment}
\begin{inouttabular}
XRSL & runTimeEnvironment & runTimeEnvironment\\
ARC-JSDL & Application/RunTimeEnvironment & Application/RunTimeEnvironment\\
\end{inouttabular}

\subsection{DataStaging}
\subsubsection{File}
\subsubsubsection{Name}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{Source}
\subsubsubsubsection{URL}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Threads}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{DataIndexingService}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{Target}
\subsubsubsubsection{URL}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Threads}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{Mandatory}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{DataIndexingService}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsubsection{NeededReplica}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{KeepData}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{IsExecutable}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsubsection{DownloadToCache}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{Directory}

\subsection{JobMetaData}
\subsubsection{Author}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{DocumentExpiration}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{Rank}
\begin{inouttabular}
XRSL & &\\
JDL & &\\
JSDL & &\\
Posix-JSDL & &\\
HPCP-JSDL & &\\
ARC-JSDL & &\\
\end{inouttabular}

\subsubsection{FuzzyRank}
\begin{inouttabular}
XRSL & ~ & ~\\
JDL & ~ & ~\\
JSDL & ~ & ~\\
Posix-JSDL & ~ & ~\\
HPCP-JSDL & ~ & ~\\
ARC-JSDL & ~ & ~\\
\end{inouttabular}

\section{XRSL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
XRSL input  & Internal representaion & XRSL output\\\hline
executable & Application.Executable.Name & executable\\
arguments & Application.Executable.Arguments & arguments\\
inputFiles & DataStaging.File.Name DataStaging.File.Source & inputFiles\\
executables & DataStaging.File.IsExecutable & executables\\
cache & TODO & TODO\\
outputFiles & DataStaging.File.Name DataStaging.File.Target & outputFiles\\
cpuTime & Resources.TotalCPUTime & cpuTime\\
wallTime & Resources.TotalWallTime & wallTime\\
gridTime & Not mapped & Not mapped\\
benchmarks & Not mapped & Not mapped\\
memory & Resources.IndividualPhysicalMemory & memory\\
disk & Resources.DiskSpaceRequirement.DiskSpace & disk\\
runTimeEnvironment & Resources.RunTimeEnvironment & runTimeEnvironment\\
middleware & Resources.CEType & middleware\\
opsys & Resources.OperatingSystem & opsys\\
stdin & Application.Input & stdin\\
stdout & Application.Output & stdout outputFiles\\
stderr & Application.Error & stderr outputFiles\\
join & Application.Join & Not mapped\\
gmlog & Application.LogDir & gmlog\\
jobName & Identification.JobName & jobName\\
ftpThreads & DataStaging.File.Source.Threads DataStaging.File.Target.Threads & Not mapped\\
acl & Application.AccessControl & acl\\
cluster & Resources.CandidateTarget.EndPointURL & Not mapped\\
queue & Resources.CandidateTarget.QueueName & queue\\
startTime & Application.ProcessingStartTime & Not mapped\\
lifeTime & Resources.SessionLifeTime & lifeTime\\
notify & Application.Notification & notify\\
replicaCollection & DataStaging.File.DataIndexingService & Not mapped\\
rerun & Application.Rerun & rerun\\
architecture & Resources.Platform & architecture\\
nodeAccess & Resources.NodeAccess & Not mapped\\
dryRun & Not mapped & Not mapped\\
rsl\_substitution & Processed internally by the parser & Not mapped\\
environment & Application.Environment & environment\\
count & Resources.SlotRequirement.ProcessPerHost & count\\
jobreport & Application.RemoteLogging & jobreport\\
credentialserver & Application.CredentialService & credentialserver\\
\end{tabular}
\end{center}
\end{table}

\section{JDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
JDL input  & Internal representaion & JDL output\\\hline
JobType & Not mapped & Not mapped\\
Executable & Application.Executable.Name & Executable InputSandbox\\
Arguments & Application.Executable.Arguments & Arguments\\
StdInput & Application.Input & StdInput InputSandbox\\
StdOutput & Application.Output & StdOutput OutputSandbox OutputSandboxDestURI\\
StdError & Application.Error & StdError OutputSandbox OutputSandboxDestURI\\
InputSandbox & DataStaging.File.Name DataStaging.File.Source & InputSandbox\\
InputSandboxBaseURI & DataStaging.File.Source.URI & InputSandbox\\
OutputSandbox & DataStaging.File.Name DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
OutputSandboxDestURI & DataStaging.File.Target DataStaging.File.KeepData & OutputSandbox OutputSandboxDestURI\\
OutputSandboxBaseDestURI & DataStaging.File.Target.URI & OutputSandbox OutputSandboxDestURI\\
Prologue & Application.Prologue.Name & Prologue\\
PrologueArguments & Application.Prologue.Arguments & PrologueArguments\\
Epilogue & Application.Epilogue.Name & Epilogue\\
EpilogueArguments & Application.Epilogue.Arguments & EpilogueArguments\\
AllowZippedISB & Stored internally, not processed & AllowZippedISB\\
ZippedISB & Stored internally, not processed & ZippedISB\\
ExpiryTime & Application.ExpiryTime & ExpiryTime\\
Environment & Application.Environment & Environment\\
PerusalFileEnable & Stored internally, not processed & PerusalFileEnable\\
PerusalTimeInterval & Stored internally, not processed & PerusalTimeInterval\\
PerusalFilesDestURI & Stored internally, not processed & PerusalFilesDestURI\\
InputData (Deprecated) & Not mapped & Not mapped\\
OutputData (Deprecated) & Not mapped & Not mapped\\
StorageIndex (Deprecated) & Not mapped & Not mapped\\
DataCatalog (Deprecated) & Not mapped & Not mapped\\
DataRequirements & Stored internally, not processed & DataRequirements\\
DataAccessProtocol & Stored internally, not processed & DataAccessProtocol\\
OutputSE & Stored internally, not processed & OutputSE\\
VirtualOrganisation & Identification.JobVOName & VirtualOrganisation\\
RetryCount & Application.Rerun & RetryCount ShallowRetryCount\\
ShallowRetryCount & Application.Rerun & RetryCount ShallowRetryCount\\
LBAddress & Stored internally, not processed & LBAddress\\
MyProxyServer & Application.CredentialService & MyProxyServer\\
HLRLocation & Stored internally, not processed & HLRLocation\\
JobProvenance & Stored internally, not processed & JobProvenance\\
NodeNumber & Stored internally, not processed & NodeNumber\\
JobSteps (Deprecated) & Not mapped & Not mapped\\
CurrentStep (Deprecated) & Not mapped & Not mapped\\
JobState (Deprecated) & Not mapped & Not mapped\\
ListenerPort & Stored internally, not processed & ListenerPort\\
ListenerHost & Stored internally, not processed & ListenerHost\\
ListenerPipeName & Stored internally, not processed & ListenerPipeName\\
Requirements & Stored internally, not processed & Requirements\\
Rank & JobMeta.Rank & Rank\\
FuzzyRank & JobMeta.FuzzyRank & FuzzyRank\\
UserTags & Identification.UserTag & UserTags\\
BatchSystem & Stored internally, not processed & BatchSystem\\
QueueName & Resources.CandidateTarget.QueueName & QueueName\\
\end{tabular}
\end{center}
\end{table}

\section{JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
JSDL input  & Internal representaion & JSDL output\\\hline
\bfseries JobIdentification & &\\
JobName & &\\
JobAnnotation & &\\
JobProject & &\\
\bfseries Application & &\\
ApplicationName & &\\
ApplicationVersion & &\\
\bfseries Resources & &\\
CandidateHosts/HostName & &\\
FileSystem/FileSystemType & &\\
FileSystem/Description & &\\
FileSystem/MountPoint & &\\
FileSystem/DiskSpace & &\\
ExclusiveExecution & &\\
OperatingSystem/OperatingSystemType/OperatingSystemName & &\\
OperatingSystem/OperatingSystemVersion & &\\
CPUArchitecture/CPUArchitectureName & &\\
IndividualCPUSpeed & &\\
IndividualCPUTime & &\\
IndividualCPUCount & &\\
IndividualNetworkBandwidth & &\\
IndividualPhysicalMemory & &\\
IndividualVirtualMemory & &\\
IndividualDiskSpace & &\\
TotalCPUTime & &\\
TotalCPUCount & &\\
TotalPhysicalMemory & &\\
TotalVirtualMemory & &\\
TotalDiskSpace & &\\
TotalResourceCount & &\\
\bfseries DataStaging & {}-{}- & {}-{}-\\
FileName & &\\
FileSystemName & &\\
CreationFlag & &\\
DeleteOnTermination & &\\
Source/URI & &\\
Target/URI & &\\
\end{tabular}
\end{center}
\end{table}

\section{POSIX-JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
POSIX-JSDL input  & Internal representation & POSIX-JSDL output\\\hline
Executable & &\\
Argument & &\\
Input & &\\
Output & &\\
Error & &\\
WorkingDirectory & &\\
Environment & &\\
WallTimeLimit & &\\
FileSizeLimit & &\\
CoreDumpLimit & &\\
DataSegmentLimit & &\\
LockedMemoryLimit & &\\
OpenDescriptorsLimit & &\\
PipeSizeLimit & &\\
StackSizeLimit & &\\
CPUTimeLimit & &\\
ProcessCountLimit & &\\
VirtualMemoryLimit & &\\
ThreadCountLimit & &\\
UserName & &\\
GroupName & &\\
\end{tabular}
\end{center}
\end{table}

\section{HPCProfile-JSDL mapping}
\begin{table}
\begin{center}
\begin{tabular}{l|c|c}
HPCProfile-JSDL input  & Internal representaion & HPCProfile-JSDL output\\\hline
Executable & & \\
Argument & & \\
Input & & \\
Output & & \\
Error & & \\
WorkingDirectory & &\\
Environment & &\\
UserName & &\\
\end{tabular}
\end{center}
\end{table}
\end{document}