### 2009-10-17 doc/tech_doc/janitor/tex_concept/concept.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

\chapter{Programming concept}

The main language for the implementation of the functionality of the
dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
integration with the Grid Manager and the Web service) for the Janitor. In
the pre-web-service implementation the Catalog remains a static web page.
The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
The modules can be separated into two
functional groups. One addresses the retrieval of information from
the Catalog RDF file in the left major branch of the figure. The other
addresses the process of fetching and installing the packages.

\begin{landscape}
 \begin{figure}[!h]
 \vspace{4cm}
 \begin{center} 
 %\resizebox{24cm}{!}{\input{images/dependencies}}
 \includegraphics[width=24cm]{images/dependencies.pdf}
 \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
 \vfill
 \label{fig:janitorDependencies} 
 \end{figure}
\end{landscape}

In order to get a more detailed view on the full functionality of the envisioned system it is suggested to consult the 
Design Document\footnote{\href{http://www.knowarc.eu/documents/Knowarc_D1.1-1_07.pdf}
{http://www.knowarc.eu/documents/Knowarc\_D1.1-1\_07.pdf}}.

\section{Runtime environments states}

A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.

With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the execution service \AREX. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time.

\begin{table}[!h]
 \begin{center}
 \begin{tabular}{lp{12cm}}
 State&Description\\
 \hline
 UNAVAILABLE&
	The RE is not available for the BaseSystem (see \ref{sec:catalog}) the site uses. \\
 INSTALLABLE&
	The RE is available for the BaseSystem the site uses and it
	will be automatically installed once a job requests it. \\
 INSTALLING/a&
	A job requested the RE and it is currently being installed \\
 INSTALLING/m&
	The RE-adminstrator requested the installation of the RE. Its
	currently being installed. \\
 FAILED&
	The installation process failed. \\
 INSTALLED/a&
	The RE is installed dynamically. \\
 INSTALLED/m&
	The RE ist installed manually by the RE-administrator \\
 BROKEN/m&
	The RE is installed but failed tests of the RE-administrator \\
 VALIDATED/m&
	The RE is installed and successfully passed the tests of the
	RE-administrator \\
 REMOVAL PENDING&
	The RE is still installed but will be removed as soon as
	possible. It is not available to new jobs. \\
 REMOVING&
	The RE is currently being removed. \\
 INSTALLED/s&
	The RE was installed in the traditional way by the site administrator. \\
 BROKEN/s&
	The RE was installed in the traditional way and failed
	validation by the RE-administrator, \\
 VALIDATED/s&
	The RE was installed in the traditional way and was
	successfully verified.
 \\
 \end{tabular}
 \end{center}
 \mycaption{States a Runtime Environment can possibly be in.}{}
 \label{tab:states}
\end{table}

The concept of ARC prevails to have a single directory into which to
install software to be execute on all compute elements of the site.
Otherwise, the installation of a runtime environment would be
required to be performed just when the jobs initiates its computation.

In principle this would be doable, particularly for those runtime 
environments that are part of a trusted Linux distribution already.
But this would also mean that the compute nodes become inhomogeneous,
and at the time of writing this was considered undesireable.

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=10cm]{images/RE_states.png}
    \mycaption{Relationships between the possible states
 of Runtime Environments.}{Red arcs represent human interaction.
 The distinction between {\bf /a}, {\bf /m} and {\bf /s}
 states does not need to be visible for all clients.}
    \label{fig:RE_states}
  \end{center}
\end{figure}

The manually induced transitions are marked in red, he automated transitions in black.
A transition between states can be induced automatically (\ie by the
advent of a job requesting a particular dynamic RE) or manually by the
site's administrator or an individual with respective rights to use the
Janitor's command line.

Upon presentation of a package to a Catalog, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

Automatically installed packages can be removed by the automatism. A
manually installed package or one that has failed to be installed,
can only be removed upon manual induction.  The \ldots/{\bf s} states
represent those Runtime Environments that are installed in the original
manual way of RE installation in ARC 0.6.

\subsection{Dependencies on Job IDs}

The Janitor was designed with the job execution in mind. As such, all
actions it performs are driven by the need of a job. And those jobs
should then be specified, to learn when the demand for a particular
runtime environment has ended.

This concept partially conflicts with the idea to use the Janitor
as an aid for the manual invocation of an installation -- there is no
job ID that could be assigned for that process. Today, the administrator
is suggested to use a special number, \eg 0, to indicate such manual
installs. This will change in near future.

\section{Job states}

The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}.
After a job has been succesfully registered with the Janitor, its state
will be set to \texttt{PREPARED}. Invalid jobs are not cached by the Janitor. 
After the Janitor is requested to deploy the runtime environment, the
state of the job will change to \texttt{INITIALIZED}. If an unforeseen
exception occures during that process, Janitor will drop the job from
its database and set the affected runtime environments to the state
\texttt{FAILED}.

\section{Integration into AREX}

\newenvironment{note}
{\rule{1ex}{1ex}\hspace{\stretch{1}}}
{\hspace{\stretch{1}}\rule{1ex}{1ex}\\}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=5cm]{images/arex-stages.pdf}
    \mycaption{}{}
    \label{fig:arex_states}
  \end{center}
\end{figure}

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=12cm]{images/janitor_integration_2nd_edition.pdf}
    \mycaption{}{}
    \label{fig:janitor_integration_2}
  \end{center}
\end{figure}

\section{WebService Interface}

Default port number: 55555\\
Client command equal, except assignment of HED.xml\\
 (from /arc1/trunk:12561)

\begin{figure}[!h]
  \begin{center}
    \includegraphics[width=12cm]{images/WS_structure.pdf}
    \mycaption{}{\task{To be translated and beautificated. SVG file is missing!}}
    \label{fig:WS_structure}
  \end{center}
\end{figure}

Proposal for SOAP messages:\\
 namespace: dynamicruntime or janitor\\
\task{Create WSDL files for that}
\task{Permission concepts: Depending on certificates. Certain certificates may \texttt{sweep}. Defined in service\_HED.xml. Evaluated in:?? }
\lstsetJUSTXML
\begin{lstlisting}[
        label=lst:arc_conf2,
        caption={ [Example \textit{arc.conf} settings for janitor.]
                  \textbf{Example \textit{arc.conf} settings for janitor.}}
        ]
<Request action="SEARCH|SWEEP|LIST|DEPLOY|REMOVE|CHECK|REGISTER">
	<Initiator jobid="1234"/>  <!-- Needed for: CHECK|REGISTER|DEPLOY|REMOVE -->
                                  <!-- May contain no jobID in this case a new one will 
                                       be created and returned via the response message-->

	<Runtimeenvironment type="dynamic">      <!-- Needed for: SEARCH|REGISTER-->
		<Package name="APPS/BIO/WEKA-3.4.10"/>
		<Package name="APPS/BIO/WEKA-3.4.11"/>
	</Runtimeenvironment>
                                  <!-- SWEEP and LIST only works, if the TLS-adminstrator identity 
                                       (which is assigned in the arched configuration file) is
                                       to be found by the SecHandler. Both need neither initiator 
                                       nor runtimeenvironment elements -->
</Request>
\end{lstlisting}

\lstsetJUSTXML
\begin{lstlisting}[
        label=lst:arc_conf3,
        caption={ [Example \textit{arc.conf} settings for janitor.]
                  \textbf{Example \textit{arc.conf} settings for janitor.}}
        ]
<response action="SEARCH|SWEEP|LIST|DEPLOY|REMOVE|CHECK|REGISTER">
	<initiator jobid="1234"/>  <!-- Needed for REMOVE|REGISTER|DEPLOY|CHECK-->
	<result code="0" message="Sucessfully initailized job."> <!-- -->
	<jobs> <!--LIST|CHECK-->
		<job jobid="1234">
			<created>1234567890</created>    <!-- in unix time-->
			<age>0</age>           <!-- in seconds-->
			<runtimeenvironment>
				<package>APPS/BIO/WEKA-3.4.10</package>
			</runtimeenvironment>
			<state>INITIALIZED</state>
		</job>
		<job jobid="4321">
			<created>1234567891</created>
			<age>0</age>
			<package>APPS/BIO/WEKA-3.4.10</package>
			<state>INITIALIZED</state>
			<runtimeenvironmentkey>APPS_BIO_WEKA_3_4_10-835614b62c98c4eb6cb03d74d3161b5d</runtimeenvironmentkey> <!-- at least CHECK -->
			<uses>/nfshome/knowarc/dredesign/src/services/dRE3/perl/spool/runtime/jre__57T1ke1UVz/runtime</uses> <!-- at least CHECK -->
			<uses>/nfshome/knowarc/dredesign/src/services/dRE3/perl/spool/runtime/weka_wHfyytarlE/runtime</uses> <!-- at least CHECK -->
		</job>
	</jobs>

	<runtimeenvironment type="local">      <!-- Needed for: LIST|SEARCH -->
		<package name="APPS/BIO/MUSTANG-3.0-1"/>
		<package name="APPS/BIO/EXONERATE-2.1.0-1"/>
	</runtimeenvironment>

	<runtimeenvironment type="dynamic">      <!-- Needed for: LIST -->
		<package name="APPS/BIO/WEKA-3.4.11">
			<state>INSTALLED_A</state>
			<lastused>1234567890</lastused>
			<jobid>1234</jobid>
		</package>
		<package name="APPS/BIO/WEKA-3.4.10">
			<state>INSTALLED_A</state>
			<lastused>1234567890</lastused>
			<jobid>1234</jobid>
			<jobid>4321</jobid>
		</package>
	</runtimeenvironment>

	<runtimeenvironment type="installable">      <!-- Needed for: LIST -->
		<package name="APPS/GRAPH/POVRAY-3.6">
			<description>The Persistence of Vision Raytracer</description>
			<lastupdate>1234567890</lastupdate>
		</package>
		<package name="APPS/BIO/WEKA-3.4.8A">
			<description>WEKA Machine Learning Software</description>
			<lastupdate>1234567890</lastupdate>
		</package>
	</runtimeenvironment>

</response>
\end{lstlisting}

\section{Janitor file system permissions}

At the time of writing, the Janitor needs to be executed as root. 
This may change over time, but as long as all other components
of ARC also demand superuser privileges, one can expect the
demand for the Janitor to be non-problematic.

\subsection{Security Consideration}

Security is a major concern for the grid systems. An automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of DRE that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention DRE names, a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lie with the maintainers of the Catalog,
who needs to be trusted.

All dynamic REs are installed in separate directories. The provisioning of
disk space is the duty of the site administrator.  In the current
implementation, the installation is completely transparent to the user:

\begin{itemize}
\item DREs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time an DRE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
side-effects by the installation of software.

The worst case scenario would be the installation of a RE
that overwrites system files. With the current implementation, which is
based solely on tar files, this is barely possible, unless such is
performed by the install scripts that accompany the tar files.
%However,
%hereto the installation would have to be performed by a user with system
%priviledges, for which there is no technical requirement.

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for DRE. In the
current implementation Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.

As a conclusion, there is general concern about the security of grid
computing.  Dynamic REs introduce new dangers since a manual control
at the grid site is substituted by a remote process that
is out the direct supervision of a local site administrator. The signing
of packages by known and directly or indirectly trusted developers is
a good indicator that no malevolent individuals have tampered with the
binary. The site administrators can limit the sources of packages and
specify packages that are eligible or excluded from installations.

% * Programming concept
% ** design decisions: Perl, RDF, Two classes entities with states: jobs and rtes
% ** State transitions of RTEs
% ** State transitions of Jobs
% ** RDF based knowledge base for runtime environments
% ** File based recovery of Jobs and RTEs
% ** Class structure
% ** Interaction with A-REX
%