### 2009-10-17 doc/tech_doc/janitor/tex_maintenance/maintenance.tex

\chapter{Maintenance}

This chapter explains how to maintain the catalog and the Janitor itself. 
In order to administrate the catalog, it is absolutly recommended (but not required) to use the ontology editor Prot\`eg\`e. The first
section will explain how this is done.
The second section gives a detailed explanation how to create new packages for the Janitor.
In the last section, a typical use case in maintaining Janitor will be demonstrated.

% * Maintainance
% ** Protege and RDF files
% ** Creating own tar balls

\section{Catalog}\label{sec:catalog}

The Catalog describes runtime environments and is either served through a web server or distributed
together with the Janitor.  It is specifed by a (Resource Description Framework) RDF file assigned to 
the Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by an RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory
\href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three
main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and
\texttt{Package}.  The \texttt{Metapackage} is a general platform
independent description of a \texttt{Package}. It can be understood as
a reference to a functionality that should be implemented at the remote
site. But the exact instruction on how to install it is not given.
The "instruction-level" comes with instances of its subclass
\texttt{Package}.

The \texttt{MetaPackage} is described by the subclasses of \texttt{Note}.
The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem}
describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid). The class \texttt{Tag} provides small keywords
which can be assigned to \texttt{MetaPackages} such that
they can be easier found.  \texttt{TarPackage} and \texttt{DebianPackage} or
currently the only subclasses of \texttt{Package}.

They are representing the necessary information (i.e. URL or Packagename)
for the installation. To provide an overview on how the classes are
interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of class \texttt{Metapackage}.}{}
         \label{tab:main_class_meta}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	homepage       & single                & String       \\
	instance       & multiple              & Instance of Package       \\
	lastupdated    & single                & String       \\
	name           & required single       & String       \\
	tag            & multiple              & Instance of Tag       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of class \texttt{BaseSystem}.}{}
         \label{tab:main_class_base}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	distribution   & required single       & String       \\
	name           & required single       & String       \\
	short\_description & required single   & String       \\
	url            & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of class \texttt{Tag}.}{}
         \label{tab:main_class_tag}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	name           & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of class \texttt{DebianPackage}.}{}
         \label{tab:main_class_debian}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	debconf        & multiple              & String       \\
	depends        & multiple              & Instance of MetaPackage  or Package      \\
	package        & required multiple     & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class \texttt{TarPackage}.}{}
         \label{tab:main_class_tar}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	basesystem     & required single       & Instance of BaseSystem       \\
	depends        & multiple              & Instance of MetaPackage or Package      \\
	environ        & multiple              & String       \\
	url            & required multiple     & String      \\
	\end{tabular} 
   \end{center}
\end{table}

\subsection{Debian packages - dysfunctional in current implementation}

The problem with Debian packages is that these are available only for the
local machine and not immediately also for the whole network. This feature
was meant for setups that use virtual machines for the execution of jobs.

Those entries can also be used to help the specification of further
dependencies of runtime environments. It would then be left to the
responsibility of the system administrator to manually (or assisted with
scripts) distribute a series of extra Debian packages throughout the
compute nodes and use the catalog entry merely to indicate their presence.

One can as such interpret the catalog to represent an interface between
software distributed via Linux distributions and independently from these
via grid communities. It should be noted that the Linux distributions
have now all started to accept communities to maintain packages, which
may bringt many scientific packages away from being traditional ARC 
runtime environments towards becoming regular packages of some Linux
distribution. For Debian, the Debian-Science and
Debian-Med\footnote{\href{http://debian-med.alioth.debian.org}{http://debian-med.alioth.debian.org}}
communities are known to be very open to grid and cloud computing.

\section{HTML interface of the catalog}

The dynamic Runtime Environments stored in the
catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also
links to both the formal Catalog in RDF syntax and an automated
transformation to HTML. The latter mimics the traditional
site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues
with an eventual transition to the new system.

That traditional page collects descriptions for Runtime Environments to
encourage human site administrators to install these.  This html page
listing the manually or automatically installable RTEs is prepared by
the script web/list.pl.  This script is meant to be run by a mod-perl
enabled Apache web server. The script itself is only loosely integrated
with the remaining functionality of the Janitor, used only for the
presentation to users. In the first lines of the script some variables
specific to the site are set. To configure the script these have to be
changed~\cite[p. 9]{BAYER_2007}.

\section{Introducing new packages}\label{sec:introducing_new_packages}

This section describes how to add new packages to the catalog. Currently
only tar based packages are processable by the Janitor.  Within the here
presented example the packages are assigned to be used together with
Debian Etch. This limitation is only literal. 
There should be no limitiation to newer Debian distributions.

\subsection{Debian Etch (tar based)}

At the time of writing, only the tape archive (tar) file format is accepted for dynamic runtime environment installation. 
It is well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files
for the representation of dynamic runtime environments. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

\begin{figure}
  \begin{center}
    \includegraphics[width=4cm]{images/tar_folder.pdf}
    \mycaption{Directory structure in the tar files for automated installation.}{}
    \label{fig:tar_folder}
  \end{center}
\end{figure}

The tar file contains two directories:
\begin{description}
\item [data/] contains all software that the grid-job may need
\item [control/] contains files formally specifying how to deal with the information in the data/ directory.
\end{description}

While installing such tar-based runtime environments, the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
But the working directory shall not be moved. All post-processing needs to be performed \textit{in situ}.

The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install}, just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty, implying that the working directory
\textdollar BAR shall be remove and no other action is required.

The remainder actions are regular actions performed upon execution of every grid-job. Upon submission,
the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. Every
ARC runtime environment must specify such a runtime script, new is only its specific location as control/runtime.
Since the directory \textdollar BAR is not known for the individual preparing the runtime environment, that
file will instead use the placeholder \%BASEDIR\%.  After installing the package, the Janitor changes all
occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.  To be offered to computing elements
for an installation, the such prepared runtime environment must be announced to a Catalog to which the
Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.

\subsection{Example: WEKA machine learning Java library}

%For an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\

The WEKA package for machine learning~\cite{FRANK_2004} and the Java Runtime Environment
are made available as dynamic Runtime Environments. Further packages for bioinformatics comprise dynamic
variants of tools for the analysis of transcription factor binding sites. These are already offered for manual
installation via the prior mentioned traditional page representing Runtime Environments for ARC. 
The corresponding tar file is named \task{somewhat??}. The \texttt{data} directory simply contains a ZIP file
which needs to be unzipped in the installation directory. For that reason, the \texttt{control/install} script is 
written as follows:
% dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
\begin{verbatim} 
#!/bin/sh
set -e  # Makes the script to terminate at the first line it fails.

WEKA_ZIP="weka-3-4-8a.zip"
unzip $WEKA_ZIP
rm -f $WEKA_ZIP
\end{verbatim}
The runtime script sets the environment variable of the Java Classpath:
\begin{verbatim}
#!/bin/sh

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}
The remove script, which will be executed right before WEKA is deinstalled, is empty. Janitor will delete the directory, so there
is nothing to be done.\\

\subsection{Example: R packages in CRAN and BioConductor.org}

To demonstrate the technical proximity to scientific communities that
provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
projects.

This effort has not found practical applications beyond this proof of concept. R packages are now
frequently found in regular Linux distributions. And for most packages, it is reasonably well possible
to install the packages at runtime.

\subsection{Example: ATLAS for High Energy Physics}

To address the concerns of the physicists using ARC, a dynamic runtime environment for the
ATLAS software suite was prepared. It extends prior work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}. The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}
What sets High Energy Physics software apart is it's sheer size. The package in question takes up more than
5 GB. This was a test illustrating the feasibility of using DREs in High Energy Physics. The application
of the DRE for ATLAS needs to wait for the planned web service extension of the Catalog. With such a
service, e.g. a software manager of a big experiment will be able to deploy software packages on production
sites simply by creating a tarball and adding an entry to the Catalog.

\subsection{Debian Etch}

These kind of packages are not yet supported.

\subsection{Debian Sid}

These kind of packages are not yet supported.

\section{Adminstrating the Catalog}