### 2009-10-25 doc/manuals/user_interface/ui.tex

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {ARC developers},
  pdftitle = {ARC Clients},
  pdfsubject = {User's manual},
  pdfkeywords = {Grid,KnowARC,ARC,NorduGrid,client},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}[1]{\colorbox{shadecolor}{\texttt #1}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\uicommand}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
\def\today{\number\day/\number\month/\number\year}

\begin{titlepage}

   \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
   \end{tabular}

   \hrulefill

   {\raggedleft NORDUGRID-MANUAL-1\par}

   {\raggedleft \today\par}

   \vspace*{2cm}

   %%%% ---- The title ----

   \begin{center}

      \textsc{\Large ARC Clients}
      \Large \par \textit{User's Manual}

      \vspace*{2cm}

      % {\centering \large Refers to ARC release series 0.4 and up \large \par}

   \end{center}

   \vspace*{2cm}

   % \abstract{
   %   This is the user's  manual to ARC User Interface.
   % }

\end{titlepage}
% \thispagestyle{empty} $ $
% \newpage
% $\ $
\tableofcontents
\newpage

\chapter{Introduction} \label{sec:intro}

The command line user interface of ARC consists of a set of commands
necessary for job submission and manipulation and data management.
This manual replaces the older version of \texttt{NORDUGRID-MANUAL-1} and is
valid for ARC versions 0.9 and above. Command line tools semantics is the same
as in earlier versions of ARC, roughly following that of basic Linux commands
and most common batch system commands. One obvious difference is change of
the legacy prefix from ``ng'' to the more appropriate ``arc''. This is not only
a cosmetic change: \textbf{behaviour of the commands also have changed}, as did
their functionalities and options.
\begin{framed}
  Users are strongly discouraged from modifying their old scripts by
  simply replacing ``ng'' with ``arc'' -- results may be unpredictable.
\end{framed}

\chapter{Commands} \label{sec:cli}

\section{Job submission and management} \label{sec:ui} 
\index{job management}
The following commands are used for job submission and management,
such as status check, results retrieval, cancellation, re-submission
and such. The jobs must be described using a job description language. ARC
supports the following languages: JSDL~\cite{jsdl}, xRSL~\cite{xrsl} and
JDL~\cite{jdl}.

\input{ui-job}

\section{Data manipulation} \label{sec:dm} 
\index{data management} ARC provides basic data management tools,
which are simple commands for file copy and removal, with eventual use
of data indexing services.

\input{ui-data}

%\section{Test suite} \label{sec:testsuite}
%\input{arctest}

%\section{Third-party commands} \label{sec:ui-other}
%\input{ui-other}

\chapter{URLs}\label{sec:urls}
\index{URLs}
\input{xrsl-urls}

\chapter{ARC Client Configuration} \label{sec:client.conf} 
\input{ui-conf}

\bibliography{grid}
\printindex

\end{document}