### 2009-10-25 doc/manuals/user_interface/ui-data.tex
arcsync, arcsub and chelonia fixed

\subsection{arcls}\label{sec:arcls}
The \texttt{arcls} \index{arcls}\index{commands:arcls} is a simple
utility that allows to list contents and view some attributes of
objects of a specified (by an URL) remote directory.

\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcls [options] $<$URL$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
    \texttt{-l} &  & detailed listing\\
    \texttt{-L} &  & detailed listing including URLs from which file can 
    be downloaded and\\ & & temporary cached locations\\
    \texttt{-m} && display all available metadata\\
    Arguments:&&\\
    \texttt{URL} && file or directory URL\\
\end{longtable}

This tool is very convenient not only because it allows to list files
at a Storage Element or records in an indexing service, but also
because it can give a quick overview of a job's working directory,
which is explicitly given by job ID.

Usage examples can be as follows:

\begin{verbatim}
    arcls rls://rc.host:38203/logical_file_name
    arcls -l gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535
    arcls -L srm://grid.uio.no:58000/srm/managerv1/johndoe/log2
\end{verbatim}

Examples of URLs accepted by this tool can be found in
Section~\ref{sec:urls}, though \texttt{arcls} won't be able to list a
directory at an HTTP server, as they normally do not return directory
listings.

\subsection{arccp}\label{sec:arccp}

The \texttt{arccp} \index{arccp}\index{commands:arccp} is a powerful
tool to copy files over the Grid. It is a part of the A-REX,
but can be used by the User Interface as well.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arccp [options] $<$source$>$ $<$destination$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
    \texttt{-y} & \textit{cache\_path} & path to local cache (use to put file into cache)\\
    \texttt{-p} && use passive transfer (does not work if secure is on, default if secure is not requested)\\
    \texttt{-n} && do not try to force passive transfer\\
    \texttt{-i} && show progress indicator\\
    \texttt{-u} && use secure transfer (insecure by default)\\
    \texttt{-r} & \textit{recursion\_level} & operate recursively (if possible) up to specified level (0 - no recursion)\\
    \texttt{-R} & \textit{number} & how many times to retry transfer of every file before failing\\
    \texttt{-t} & \textit{time} & timeout in seconds (default 20)\\
    \texttt{-f} && if the destination is an indexing service and not the same as the source and the destination is already registered, then the copy is normally not done. However, if this option is specified the source is assumed to be a replica of the destination created in an uncontrolled way and the copy is done like in case of replication\\
    \texttt{-T} && do not transfer file, just register it - destination must be non-existing meta-url\\
    Arguments:&&\\
    \texttt{source} && source URL\\
    \texttt{destination} && destination URL\\
\end{longtable}

This command transfers contents of a file between 2 end-points.
End-points are represented by URLs or meta-URLs. For supported
endpoints please refer to Section~\ref{sec:urls}.

\texttt{arccp} can perform multi-stream transfers if \texttt{threads}
URL option is specified and server supports it.

Source URL can end with \verb#"/"#. In that case, the whole fileset
(directory) will be copied. Also, if the destination ends with
\verb#"/"#, it is extended with part of source URL after last
\verb#"/"#, thus allowing users to skip the destination file or
directory name if it is meant to be identical to the source.

Usage examples of \texttt{arccp} are:

\begin{verbatim}
    arccp gsiftp://lscf.nbi.dk:2811/jobs/1323842831451666535/job.out \
              file:///home/myname/job2.out
    arccp gsiftp://aftpexp.bnl.gov;threads=10/rep/my.file \
              rc://;threads=4@grid.uio.no/lc=Collection,rc=Catalog/zebra4.f
    arccp http://www.nordugrid.org/data/somefile gsiftp://hathi.hep.lu.se/data/
\end{verbatim}

\subsection{arcrm}\label{sec:arcrm}

The \texttt{arcrm} \index{arcrm}\index{commands:arcrm}
command allows users to erase files at any location specified by a
valid URL.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcrm [options] $<$source$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
    Options:&&\\
    \texttt{-h} && short help\\
    \texttt{-v} && print version information\\
    \texttt{-d} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
    \texttt{-c} & &continue with meta-data even if it failed to delete real file\\
    Arguments:&&\\
    \texttt{source} && source URL\\
\end{longtable}

\begin{framed}
   A convenient use for \texttt{arcrm} is to erase the files in a data
   indexing catalog (RC, RLS or such), as it will not only remove the
   physical instance, but also will clean up the database record.
\end{framed}

Here is an \texttt{arcrm} example:

\begin{verbatim}
    arcrm rc://grid.uio.no/lc=Collection,rc=Catalog/badfile
\end{verbatim}

\subsection{arcacl}\label{sec:arcacl}

\index{arcacl}\index{commands:arcacl}This command retrieves or modifies
access control information associated with a stored object if service
supports GridSite GACL language~\cite{gacl} for access control.
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{arcacl [options] get$|$put $<$URL$>$}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
    \texttt{-d, -debug} & \textit{debuglevel} &debug level is one of  FATAL, ERROR, WARNING, INFO, DEBUG or VERBOSE\\
    \texttt{-v} && print version information\\
    \texttt{-h} && short help\\
   Arguments:&&\\
    \texttt{get} &\textit{URL}& get Grid ACL for the object\\
    \texttt{put} &\textit{URL}& set Grid ACL for the object\\
    \texttt{URL} && object URL; curently only gsiftp and sse URLs are supported\\
\end{longtable}

ACL document (an XML file) is printed to standard output when
\texttt{get} is requested, and is acquired from standard input when
\texttt{set} is specified\footnote{In ARC $\leq$ 0.5.28, \texttt{set}
  shoud be used instead of \texttt{put}}. Usage examples are:
\begin{verbatim}
    arcacl get gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename
    arcacl set gsiftp://se1.ndgf.csc.fi/ndgf/tutorial/dirname/filename $<$ myacl
\end{verbatim}

\subsection{arctransfer}\label{sec:arctransfer}
\index{arctransfer}\index{commands:arctransfer}The \texttt{arctransfer}
This command is not implemented.

\subsection{chelonia}\label{sec:chelonia}

\index{chelonia}\index{commands:chelonia}\texttt{chelonia} is a client tool for
accessing the Chelonia storage system. With it you can
create, remove and list file collections, upload, download and remove files,
and move and stat collections and files, using Logical Names (LN).
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{chelonia [options] $<$method$>$ [arguments]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-b} & \textit{URL} & URL of Bartender to connect\\
   \texttt{-x} && print SOAP XML messages\\
   \texttt{-v} && verbose mode\\
   \texttt{-z}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
   \texttt{-w} && allow to run without the ARC python client libraries (with limited functionality)\\
   Methods:&&\\
   \texttt{stat} &\textit{LN [LN ...]}& get detailed information about an entry or several\\
   \texttt{makeCollection, make, mkdir} &\textit{LN}& create a collection\\
   \texttt{unmakeCollection, unmake, rmdir} &\textit{LN}& remove an empty collection\\
   \texttt{list, ls} &\textit{LN}& list the content of a collection\\
   \texttt{move, mv} &\textit{source target}& move entries within the namespace (both LNs)\\
   \texttt{putFile, put} &\textit{source target}& upload a file from a \textit{source} to a \textit{target} (both specified as LNs))\\
   \texttt{getFile, get} &\textit{source [target]}& download a file from a \textit{source} to a \textit{target}\\
   \texttt{delFile, del, rm} &\textit{LN [LN ...]}& remove file(s))\\
   \texttt{modify, mod} &\textit{string}& modify metadata\\
   \texttt{policy, pol} &\textit{string}& modify access policy rules. The string has a form $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$.\\
   & &$<$changeType$>$ could be 'set', 'change' or 'clear'\\
   & & 'set': sets the action list to the given user overwriting the old one\\
   & & 'change': modify the current action list with adding and removing actions\\
   & & 'clear': clear the action list of the given user\\
   & &$<$identity$>$ could be a '$<$UserDN$>$' or a 'VOMS:$<$VO name$>$'\\
   & &$<$action list$>$ is a list actions prefixed with '+' or '-', e.g. '+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
   \texttt{unlink} &\textit{string}& remove a link to an entry from a collection without removing the entry itself\\
   \texttt{credentialsDelegation, cre} &\textit{string}& delegate credentials for using gateway\\
   \texttt{removeCredentials, rem} &\textit{string}& remove previously delegated credentials\\
   \texttt{makeMountPoint, makemount} &\textit{string}& create a mount point\\
\end{longtable}

Without arguments, each method prints its own help.

Examples:
\begin{verbatim}
    chelonia list /
    chelonia put orange /
    chelonia stat /orange
    chelonia get /orange /tmp
    chelonia mkdir /fruits
    chelonia mkdir /fruits/apple
    chelonia mv /orange /fruits
    chelonia ls /fruits
    chelonia rmdir /fruits/apple
    chelonia rmdir /fruits
    chelonia rm /fruits/orange
    chelonia policy / change ALL +read +addEntry
    chelonia modify /pennys-orange set states neededReplicas 2
\end{verbatim}