### 2009-10-28 doc/manuals/chelonia/chelonia-cli.tex
CLI manual: moved chelonia section under doc/chelonia, fixed Makefile

\subsection{chelonia}\label{sec:chelonia}

\index{chelonia}\index{commands:chelonia}\texttt{chelonia} is a client tool for
accessing the Chelonia storage system. With it you can
create, remove and list file collections, upload, download and remove files,
and move and stat collections and files, using Logical Names (LN).
\hspace*{0.5cm}
\begin{shaded}
   \uicommand{chelonia [options] $<$method$>$ [arguments]}
\end{shaded}
\versions{ARC 0.9}
\begin{longtable}{llp{8cm}}
   Options:&&\\
   \texttt{-b} & \textit{URL} & URL of Bartender to connect\\
   \texttt{-x} && print SOAP XML messages\\
   \texttt{-v} && verbose mode\\
   \texttt{-z}&\textit{filename}& configuration file (default {\$}HOME/.arc/client.xml)\\
   \texttt{-w} && allow to run without the ARC python client libraries (with limited functionality)\\
   Methods:&&\\
   \texttt{stat} &\textit{LN [LN ...]}& get detailed information about an entry or several\\
   \texttt{makeCollection, make, mkdir} &\textit{LN}& create a collection\\
   \texttt{unmakeCollection, unmake, rmdir} &\textit{LN}& remove an empty collection\\
   \texttt{list, ls} &\textit{LN}& list the content of a collection\\
   \texttt{move, mv} &\textit{source target}& move entries within the namespace (both LNs)\\
   \texttt{putFile, put} &\textit{source target}& upload a file from a \textit{source} to a \textit{target} (both specified as LNs))\\
   \texttt{getFile, get} &\textit{source [target]}& download a file from a \textit{source} to a \textit{target}\\
   \texttt{delFile, del, rm} &\textit{LN [LN ...]}& remove file(s))\\
   \texttt{modify, mod} &\textit{string}& modify metadata\\
   \texttt{policy, pol} &\textit{string}& modify access policy rules. The string has a form $<$LN$>$ $<$changeType$>$ $<$identity$>$ $<$action list$>$.\\
   & &$<$changeType$>$ could be 'set', 'change' or 'clear'\\
   & & 'set': sets the action list to the given user overwriting the old one\\
   & & 'change': modify the current action list with adding and removing actions\\
   & & 'clear': clear the action list of the given user\\
   & &$<$identity$>$ could be a '$<$UserDN$>$' or a 'VOMS:$<$VO name$>$'\\
   & &$<$action list$>$ is a list actions prefixed with '+' or '-', e.g. '+read +addEntry -delete'; possible actions are: \texttt{read}, \texttt{addEntry}, \texttt{removeEntry}, \texttt{delete}, \texttt{modifyPolicy}, \texttt{modifyStates}, \texttt{modifyMetadata}\\
   \texttt{unlink} &\textit{string}& remove a link to an entry from a collection without removing the entry itself\\
   \texttt{credentialsDelegation, cre} &\textit{string}& delegate credentials for using gateway\\
   \texttt{removeCredentials, rem} &\textit{string}& remove previously delegated credentials\\
   \texttt{makeMountPoint, makemount} &\textit{string}& create a mount point\\
\end{longtable}

Without arguments, each method prints its own help.

Examples:
\begin{verbatim}
    chelonia list /
    chelonia put orange /
    chelonia stat /orange
    chelonia get /orange /tmp
    chelonia mkdir /fruits
    chelonia mkdir /fruits/apple
    chelonia mv /orange /fruits
    chelonia ls /fruits
    chelonia rmdir /fruits/apple
    chelonia rmdir /fruits
    chelonia rm /fruits/orange
    chelonia policy / change ALL +read +addEntry
    chelonia modify /pennys-orange set states neededReplicas 2
\end{verbatim}