### 2009-11-25 doc/manuals/janitor/tex_maintenance/maintenance.tex

\chapter{Maintenance}

This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
It gives detailed instructions how to create new packages for the Janitor.
To administrate the catalog, it is strongly recommended (but not
ultimately required) to use the ontology editor Prot\`eg\`e, as explained
in the next section.
In the last section, a typical use case in maintaining the Janitor will be shown.

% * Maintainance
% ** Protege and RDF files
% ** Creating own tar balls

\section{Catalog}\label{sec:catalog}

The Catalog describes runtime environments and is either served
through a web server or is distributed together with the Janitor as a
regular file.  It is specifed by a (Resource Description Framework)
RDF file assigned to Janitor using the tag \texttt{catalog} within
the configuration file (see~\ref{sec:janitor_configuration}).
The format of the RDF file is defined by an RDF schema file
\texttt{knowarc.rdfs} which can be found along with an RDF
example file \texttt{knowarc.rdf} in the Janitor source directory
\href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.

To edit the catalog, the ontology editor Prot\`eg\`e should be
used. Figure~\ref{fig:protege_example} shows the editor while the
MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has
been selected.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}

On the left side of the editor the class browser is placed.  Three
main classes are prepared: \texttt{MetaPackage}, \texttt{Note}
and \texttt{Package}.  The \texttt{Metapackage} is a general
platform-independent description of a \texttt{Package}.  It has one
or more instances of the class \texttt{Package} and is described
by the subclasses of \texttt{Note}. The class \texttt{Note} has
two subclasses: \texttt{BaseSystem} and \texttt{Tag} to describe
the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the
Debian release a \texttt{Package} refers to (i.e. here etch or
sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned
to \texttt{MetaPackages} such that they can be found more easily.
\texttt{TarPackage} and \texttt{DebianPackage} or currently the only
subclasses of \texttt{Package}.  They are representing the necessary
information (i.e. URL or Packagename) for the installation. In order
to have an overview how the classes are interacting with each other
the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base},
\ref{tab:main_class_tag}, \ref{tab:main_class_debian} and
\ref{tab:main_class_tar} are pictured.

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Metapackage.}{}
         \label{tab:main_class_meta}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	homepage       & single                & String       \\
	instance       & multiple              & Instance of Package       \\
	lastupdated    & single                & String       \\
	name           & required single       & String       \\
	tag            & multiple              & Instance of Tag       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Basesystem.}{}
         \label{tab:main_class_base}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	distribution   & required single       & String       \\
	name           & required single       & String       \\
	short\_description & required single   & String       \\
	url            & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class Tag.}{}
         \label{tab:main_class_tag}
	\begin{tabular}{p{3cm}p{3cm}p{4cm}}
	\textbf{Name}  & \textbf{Cardinality}  & \textbf{Type}\\
	\hline
	description    & single                & String       \\
	name           & required single       & String       \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class DebianPackage.}{}
         \label{tab:main_class_debian}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name} & \textbf{Cardinality} & \textbf{Type}\\
	\hline
	basesystem    & required single      & Instance of BaseSystem \\
	debconf       & multiple             & String \\
	depends       & multiple             & Instance of MetaPackage or Package \\
	package       & required multiple    & String \\
	\end{tabular} 
   \end{center}
\end{table}

\begin{table}[!h]
   \begin{center}
        \mycaption{Specification of the class TarPackage.}{}
         \label{tab:main_class_tar}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
	\textbf{Name} & \textbf{Cardinality} & \textbf{Type}\\
	\hline
	basesystem    & required single      & Instance of BaseSystem \\
	depends       & multiple             & Instance of MetaPackage or Package \\
	environ       & multiple             & String \\
	url           & required multiple    & String \\
	\end{tabular} 
   \end{center}
\end{table}

\section{HTML interface of the catalog}

The dynamic Runtime Environments stored in the
Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also
links to both the formal Catalog in RDF syntax and its automated
transformation to HTML.  The latter mimics the traditional
site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with
an eventual transition to the new system.  That page collects descriptions
for Runtime Environments to encourage human site administrators to
install these.  This HTML page listing the manually or automatically
installable REs is prepared by the script web/list.pl.  This script is
meant to be run by a mod\_perl-enabled Apache. The script itself does
not contribute to the core functionality of the Janitor.  In the first
lines of the script some variables specific to the site are set. To
configure the script these have to be changed~\cite[p. 9]{BAYER_2007}.

\section{Introducing new packages}\label{sec:catalog}

This section describes how to add new packages to the Catalog. In the
current implementation, only tar based packages are processed by the
Janitor.  Within the example they are assigned to be used together with
Debian Etch. This limitation is only literal.  There is no restriction
for newer Debian distributions.

\subsection{Debian Etch (tar based)}

At the time of writing, only the tape archive (tar) file format is
accepted for dynamic Runtime Environment installation, a well accepted
file format throughout the UNIX community. The concept reflects the
traditional manual approach towards RTE in ARC, for which one directory
is made available to all compute nodes. This section explains the
inner structure of the tar files.  Subdirectories are visualised in
Figure~\ref{fig:tar_folder}.

\begin{figure}
  \begin{center}
    \includegraphics[width=4cm]{images/tar_folder.pdf}
    \mycaption{Directory structure in the tar files for automated installation.}{}
    \label{fig:tar_folder}
  \end{center}
\end{figure}

The tar file contains two directories, named {\tt control} and {\tt
data}. Software is stored in the latter subdirectory, while the files
formally specifying how to deal with such packages are stored in the
prior. Upon installation, the content of the data directory is extracted
to some directory \textdollar BAR.  After this unpacking of the tar file,
the Janitor executes the install script provided in the control directory.
It is executed within the working directory \textdollar BAR. The job of
this skript is to perform any necessary post-processing.  The Janitor
stores the file \texttt{control/remove}. It will be executed in the
same way as \texttt{control/install} just before the tar-package is
removed. In most cases \texttt{control/remove} will be empty.  Finally,
the file \texttt{control/runtime} is sourced multiple times by the
Grid Manager's job-submit script. After installing the package, the
Janitor changes all occurences of \%BASEDIR\% in the runtime script to
\textdollar BAR.  Once the tar file was prepared, it must its entry to
a RTE Catalogue~\cite[p. 10]{BAYER_2007}.

From such Catalogs, the Janitor finds all information to install
packages that possibly have never been installed on the site before. The
offers of RTEs in a Catalog are cross-checked against the local
infrastructure and a subset of the available packages will be accepted as
"installable". This list of installable RTEs is forwarded to the grid
information system.

\subsection{Automated transformation of install directory to dRTE}

The script 'prepareDRE.pl' was created to help with the transformation
of a readily installed software into a dynamic runtime environment.
I also prepares a complete catalog file that can be offered
individually or next to other catalogs. See the associated man page
prepareDRE(8) for details.

\subsection{Protoypes}

In order to have an impression how the tar files are created, several
prototypes are provided at \url{http://dre.knowarc.eu}.

The WEKA package for machine learning~\cite{FRANK_20074} and the Java
Runtime Environment are available as dynamic Runtime Environments. Further
packages for bioinformatics comprise dynamic variants of tools for
the analysis of transcription factor binding sites. These are already
offered for manual installation via the prior mentioned traditional page
representing Runtime Environments for ARC.  The corresponding tar file
is named \task{somewhat??}. The \texttt{data} directory simply contains
a ZIP file which needs to be unzipped in the installation directory. For
that reason, the \texttt{control/install} script is written as follows:

% dollar signs made trouble in listings *hrmlgrmpf* ... Costs too much time to figure out why *ARC!*
\begin{verbatim} 
#!/bin/sh
set -e  # Makes the script to terminate at the first line it fails.

WEKA_ZIP="weka-3-4-8a.zip"
unzip $WEKA_ZIP
rm -f $WEKA_ZIP
\end{verbatim}
The runtime script sets the environment variable of the Java Classpath:
\begin{verbatim}
#!/bin/sh

WEKA_JAR="weka-3-4-8a/weka.jar"
case "\$1" in
	0)	# Just before job submission
		# none
	;;
	1)	# Just before job execution
		# Initialize the java environment
		CLASSPATH="%BASEDIR%/$WEKA_JAR:$CLASSPATH"
		export CLASSPATH
	;;
	2)	# After job termination
		# none
	;;
	*)
		return 1
	;;
esac	
\end{verbatim}

The remove script, which will be executed right before WEKA is deinstalled, is empty.
Janitor will delete the directory, so there is nothing more to be done.

%\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
%provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
%dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
%thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
%projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

To address the concerns of the physicists using ARC, a dynamic runtime
environment for the ATLAS software suite was prepared. It extends prior
work on an automated installation that is available at
\href{http://guts.uio.no/atlas/12.0.6/}{http://guts.uio.no/atlas/12.0.6/}.
The preparation comprised the following steps:
\begin{itemize}
    \item The file system path specifications in the automated
       installation scripts were modified using the Janitor
       path variables.
    \item A tarball was prepared containing a directory structure as
       illustrated in Figure~\ref{fig:tar_folder}. The data directory
       was empty, since the automatic installation script downloads the
       software from a remote server.
    \item An entry was added to the Catalog file.
\end{itemize}

What sets High Energy Physics software apart is it's sheer size. The
package in question takes up more than 5 GB. This was a test illustrating
the feasibility of using dynamic REs in High Energy Physics. The
application of the dynamic RE for ATLAS needs to wait for the planned web
service extension of the Catalog. With such a service, e.g. a software
manager of a big experiment will be able to deploy software packages
on production sites simply by creating a tarball and adding an entry to
the Catalog.

%\subsection{Debian Etch}
%
%These kind of packages are not yet supported.
%
%\subsection{Debian Sid}
%
%These kind of packages are not yet supported.
%
%\subsection{Adminstrating the Catalog}

%\section{Typical use cases}\label{sec:catalog}

% ** setstate REMOVAL_PENDING
% no warranty interferences in the file syste,