### 2009-11-29 doc/tech_doc/janitor/tex_installation/installation.tex
Added symbolic link for compatibility until the tech-doc/janitor directory is removed.

../../../manuals/janitor/tex_installation/installation.tex