### 2009-11-29 doc/manuals/janitor/tex_introduction/introduction.tex
Merged manual/tech_doc janitor/introduction

\chapter{Introduction}

The {\em Janitor} is a service for the automated installation of runtime
environments for grid computing elements.  Its command line interface
allows for a direct interaction with site administrators. However, the
main stimulus for its development was the idea integrate such a service
with the regular handling of compute jobs. For ARC this is performed by
the \AREX module.

From the programmer's view, the Janitor is mostly a Perl script and the routines within \AREX to invoke it. The
site administrator will also associate with it also the Catalog files, that describe the availability of runtime
environments, and the repository of installable runtime envionments themselves, which are regular tar archives obeying
to particular structure and reside in a separate folder. In order to minimise the latency for the invocation of
the Perl script and the associated parsing of files, a Janitor web service was developed, which still is a Perl
script.

\section{Motivation}

A major motivation for grid projects is to stimulate new communities
to adopt the technology to start sharing their resources. From the
current grid user's viewpoint, the admission of users with a very different
education will suddenly impose difficulties in the communication between
site maintainers. One will not even understand the respective other side's
research aims. Hence, the proper installation of non-standard software
(Runtime Environments, RTEs) is not guaranteed. And extra time for manual
labour plus self-education is scarse.

A core problem remains to distribute a locally working solution, the
Know-How, quickly across all contributing sites, \ie without manual
interference. Every scientific discipline has its respective own
set of technologies for the distribution of work load. For instance,
research in bioinformatics requires access to so many different tools
and databases, that few sites, if any, install them all. Instead, the use
of web services became a commodity, with all their intrinsic problems as
there are bottlenecks and restrictions of repeated access.  The EU project
KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}},
amongst other challenges, with the here presented work
extends the NorduGrid's Advanced Research Connector (ARC) grid
middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic
Runtime Environments (dRTEs), seems the only approach to use the
computational grid to its full potential. Components of workflows
shall be spawned as jobs in a computational grid using dRTEs rather
than accessing a public web service at one particular machine that
is shared amongst all users.  The grid introduces an extra level of
parallelism that web services cannot provide. The demands for short
response times and the heterogeneous education of site-administrators
on a grid demand an automatism for the installation of software and
databases without manual interference~\cite{BAYER_2007}.

\section{Overview}

This document starts with a chapter on how to set-up the Janitor
locally. It is followed by a chapter that gives further instructions on
how to use the Janitor with A-REX and/or without A-REX. Afterwards, in the
third chapter, the maintenance of the program will be presented, which is
basically covering the method how to prepare new dRTEs. Deeper insights on
the design of the Janitor will be given by the forth chapter. The document
ends with an outlook to anticipated future developments and opportunities.

\section*{Abbreviations}

\begin{tabular}{l@{--}p{150mm}}
RTE&Runtime Environment\\
dRTE&dynamic Runtime Environment\\
RDF&Resource Description Framework (supporting the RTE Catalog)\\
\end{tabular}

% BASIC CONCEPT

% * Installation

% * Usage

% * Maintainance

% * Programming concept

% * Future work

% * APPENDIX