### 2009-11-29 doc/tech_doc/janitor/tex_introduction/introduction.tex
Added temporarily symbolic link from tech_doc to manual for janitor/introduction until transition is completed.

../../../manuals/janitor/tex_introduction/introduction.tex