### 2009-11-29 doc/tech_doc/janitor/tex_maintenance/maintenance.tex
Added symbolic link to help transition.

../../../manuals/janitor/tex_maintenance/maintenance.tex