### 2009-11-29 doc/manuals/janitor/tex_appendix/appendix.tex
deuglification

\chapter{Appendix}

\section{Useful tutorials and documentations}

\begin{itemize}
 \item \textbf{Another document describing Janitor.}\\
    D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs, Daniel Bayer and Steffen M\"oller and Frederik Orellana\cite{BAYER_2007}

\end{itemize}

% * Literature
% ** www.nordugrid.org/documents/xrsl.pdf
% ** Literature for JSDL is missing
% ** ``"D2.5-1 RDF Based Semantic Runtime Environment (RE) Description And Dynamic RE Management Framework Including Creating Proof Of Concept Bioinformatics REs" (see also prototype code and code documentation), May 2007''
%    http://www.knowarc.eu/documents/Knowarc_D2.5-1_07.pdf
% ** Code documentation http://www.knowarc.eu/download/D2.5-1_documentation.pdf