### 2011-01-11 doc/tech_doc/xrsl/xrsl.tex

\documentclass{book}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=3cm,lmargin=2cm,rmargin=3cm}
\usepackage{makeidx}
\makeindex
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color,framed}
\usepackage{longtable}                             %for breaking tables

%%%%---- Use EPS figures and graphics with LaTeX ----
%\usepackage{graphics}
%%%%---- USE PNG or JPG figures and graphicx with PDFLaTeX ----
\usepackage{graphicx}
\usepackage[linkbordercolor={0 0.8 0.8}]{hyperref}
\hypersetup{
  pdfauthor = {Oxana Smirnova},
  pdftitle = {Extended Resource Specification Language},
  pdfsubject = {Reference manual},
  pdfkeywords = {Grid,NorduGrid,ARC,Globus,RSL},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}

\bibliographystyle{IEEEtran}

\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\xrsl}[1]{\textbf{\sffamily #1}}
\newcommand{\versions}[1]{\vspace{-0.3cm}(#1)\par}
\newcommand{\GT}{Globus Toolkit$\mathrm{^{TM}}$}
\newcommand{\globus}{Globus\textsuperscript{\textregistered}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{cyan}{rgb}{0,1,1}

\begin{document}
  \def\today{\number\day/\number\month/\number\year}

  \begin{titlepage}

    \begin{tabular}{rl}
      \resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
      &\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
    \end{tabular}

    \hrulefill

    {\raggedleft NORDUGRID-MANUAL-4\par}

    {\raggedleft \today\par}

    \vspace*{2cm}

%%%%---- The title ----

    \begin{center}

    \textsc{\Large Extended Resource Specification Language}
     \Large \par \textit{Reference Manual}

    \vspace*{2cm}

%    {\centering \large Refers to ARC release series 0.8 and up \large \par}

    \end{center}

    \vspace*{2cm}

%    \abstract{
%      The paper represents an overview and a reference manual for the
%      NorduGrid Extended RSL, developed on the basis of the \globus\
%      RSL.
%    }

  \end{titlepage}
  %\thispagestyle{empty} $ $
  %\newpage
  %$\ $
  \tableofcontents
  \newpage

  \chapter{Introduction}
  \label{sec:intro}

  Computational tasks, such as data mining, data transformation, simulation of various conditions, execution of Monte Carlo algorihms and so on, are normally formalized as series of \textit{jobs} submitted to computing resources. Such computing resources can be mainframe supercomputers, high-performance computing clusters, or even pools of regular desktop PCs. In a conventional approach, applicaton experts obtain user acconts at computing centers and learn peculiarities of their \textit{batch systems} in order to formulate and submit computational jobs to the assigned resource.
  In the Grid infrastructure~\cite{grid}, experts get simultaneous access to a large number of very different computing resources, which are often widely distributed geographically. In order to abstract from the heterogeneous nature of Grid resources, a high-level job description language is necessary. The diversity of Grid resources implies special requirements for a
  proper description of a job, introducing many new options as opposed to a conventional high-performance computing center use case.

  Globus\ \index{Globus}
  Alliance~\cite{globus} developed a Grid middleware toolkit, early versions of which
  made use of the \textit{Resource Specification Language
  (RSL)}~\cite{rsl}\index{RSL} to parse job options and definitions to
  resource management systems. The NorduGrid project~\cite{nordugrid}
  \index{NorduGrid} decided to re-use RSL when they developed the Advanced Resource Connector (ARC),
  -- a
  solution~\cite{arc} for a Grid facility, suitable for
  complex tasks, like, for example, High Energy Physics
  data processing. To match the complexity of such tasks, this solution
  requires certain extensions to the RSL.

  To describe a task to be submitted to ARC-enabled resources, an
  extended version of the \globus\  RSL is used\footnote{One can also use a Grid-standard JSDL language; however, it is less versatile}. Extensions concern
  not only introduction of new attributes, but also differentiation
  between the two levels of job option specifications:

  \begin{itemize}
  \item[] \textbf{User-side RSL}, i.e., the set of attributes
    specified by a user in a job-specific file. This file is
    interpreted by a \textit{Client}~\cite{ui},
    \index{Client} and after the necessary modifications is
    passed to the ARC computing service: the \textit{Grid Manager (GM)}~\cite{gm}\index{Grid
    Manager}
  \item[] \textbf{GM-side RSL}, i.e., the set of attributes
    pre-processed by a client, and ready to be interpreted by the GM. Effectively, this is an \textit{internal} job representation of ARC.
  \end{itemize}

  A user only has to know the user-side part, and utilize it to
  describe the Grid tasks. The Grid Manager, however, uses slightly
  different notations, supplied by the client tools; therefore developers of such tools must take care of converting user-submitted RSL to the internal job description (GM-side RSL).

  In what follows, description of the NorduGrid-extended RSL,
  further denoted as \textbf{xRSL}\index{xRSL}, is given, using the
  following notations:

  \begin{tabular}{ll}
    \verb#<xxxx>#&parameter to be substituted with a corresponding string or a number\\
    \verb#[xxxx]# & optional parameter\\
    \verb#xxx|yyy|zzz# &list of possible values of a parameter\\
    \verb#-"-# &"same as above"\\
  \end{tabular}

Usage examples are given for the standard ARC command-line client (also known as \textit{User Interface})~\cite{ui}; other clients may introduce additional features, please refer to their respective documentation for details.

  \chapter{xRSL syntax and rules}
  \label{sec:remarks}

  For a complete description of \globus\  RSL, see
  reference~\cite{rsl}. xRSL uses the same syntax conventions,
  although changes the meaning and interpretation of some attributes.

  \section{Syntax}
  \label{sec:syntax}

  A Grid task is described by means of xRSL attributes, which can
  be either passed via a command-line, or, more conveniently, be
  collected in a so-called xRSL-file (suggested extension
  \textit{.xrsl}). Such a file contains a plain list of attribute-value pairs
  and boolean operators "\&" (for AND) and "\verb#|#" (for
  OR). Attribute names are case-insensitive.

  If the attributes are specified from the command line, the entire
  description must be enclosed either in single or in double quotes.
  Single quotes enclosure is completely analogous to xRSL-file usage,
  while double quotes imply standard shell expansion of the enclosed
  string. This has implications when strings contain local shell
  variables: they \textbf{will not be expanded} unless the task description is
  entered from the command line and is enclosed in double quotes.

  An attribute-value pair is a key element of specification. It
  consists of an expression that assigns one or more values to an
  attribute, and is enclosed in round brackets:

  \begin{shaded}
    \verb#(attribute="value")#\\
    \verb#(attribute="value1" "value2")# (for multi-valued attributes)
  \end{shaded}

  Certain attributes do not have assigned value; instead, they contain
  a list of values that should be put in proper correspondence with
  each other:

  \begin{shaded}
    \verb#(attribute=("value1" "value2")("value3" "value4"))#
  \end{shaded}

  In the example above, \verb#value1# and \verb#value3# are put in correspondence to
  \verb#value2# and \verb#value4# respectively, according to the
  context of the attribute.

  Values should be enclosed in quotes if they contain blank
  spaces or special characters. The \textbf{special characters} are:
  \begin{shaded}
    \verb. +   &   |   (   )   =   <   >   !   "   '   ^   #   $.%$
  \end{shaded}

  To quote a string containing special characters, you can use
  either single or double quotes. If your string, however, contains
  both such quotes, you can define any character as an own
  delimiter, by preceding it with the "carat" (\verb#^#) character:
  \verb#attribute=^*My "good" value^*# makes use of a carat-escaped
  asterisk as a delimiter.

  An xRSL job description starts with an ampersand
  ("\&") , to indicate implicit \textbf{conjunction} of all the
  attributes:
  \begin{shaded}
    \verb#&(attribute1=value1)(attribute2="value 2")...#
  \end{shaded}

  Whenever a \textbf{disjunct}-request of two or more attributes is
  needed, the following construction can be used:
  \begin{shaded}
    \verb#(|(attribute="value1")(attribute="value2")...)#
  \end{shaded}

  \begin{framed}
   Only few selected attributes (indicated further in the document) can be requested by the user multiple times, like in the disjunct request example above. Most attributes \textbf{must be unique}, i.e., appear only once in the job description document.
  \end{framed}

  In expressions, the following relational operators are allowed, in general:
  \begin{shaded}
    \verb#=   !=   >   <   >=   <=#
  \end{shaded}
  However, most attrbutes can \textbf{only be used with equality operator "="}. For few attributes (as indicated in the document), some other operators can be used as well.

  \textbf{Commented} lines should start with "(*" and be
  closed with "*)":
  \begin{shaded}
    \verb#(*attribute="value1"*)#
  \end{shaded}
  Comments can not be nested.

  \textbf{Multiple job} description in one file is realized via a standard
  \globus\  RSL multi-request operator "+", which should precede
  multiple job description:
  \begin{shaded}
    \verb#+(&(...))(&(...))(&(...))#
  \end{shaded}

  The xRSL attributes can be written in a single string, or split in
  lines arbitrary; blank spaces between and inside
  \verb#(attribute="value")# relations are ignored.

  \section{URLs}
  \label{sec:url}

  \input{xrsl-urls}

  \chapter{Attributes}
  \label{sec:attributes}

  Most of the job description attributes \index{attributes} introduced originally by \globus\ RSL~1.0 are
  supported, some with modifications as indicated in this
  document. Many new attributes are introduced by ARC, of which some are
  to be specified in the user's script, and others are internal for
  the GM (are added or modified by client tools).

  Attribute names are case-insensitive, although assigned values may
  well be case-sensitive, if they represent file names, environment
  variables etc..

  It is possible to use unsupported attributes in job description. Standard ARC client submission commands (\texttt{ngsub} and \texttt{ngresub}) must be used with a command line option ``-U'' in order to accept unknown attributes. Without this command line option, the client tool will concider job description invalid if it contains unsupported attributes.

  \section{User-side attributes}
  \label{sec:user}

  The following attributes \index{attributes:user-side} can be
  specified in a user's xRSL script. Some have to be modified by the client tool
  before being passed to the GM. If this is the case, the corresponding modified GM input is described in this document as well.

  \input{xrsl-user}

  \section{GM-side attributes}
  \label{sec:gm}

  The following attributes are a part of the internal ARC job representation, and must be provided by ARC client tools and passed to
  the GM. Users may even specify them manually (\textbf{not
  advised!}), while developers of new ARC client tools and utilities must make sure these attributes are added to the user job description before it is submitted to a GM.

  \input{xrsl-gm}

  \section{Unsupported Globus RSL attributes}
  \label{sec:globus}

  The following \globus\ attributes are not supported by the ARC
  middleware. Whenever they are specified, eiher an error or a warning should be issued by the
  client tool, and the corresponding attribute should be ignored\footnote{Standard ARC client will throw an error, unless option ''-U'' is specified}.

  \subsection{Unsupported RSL 1.0 attributes}

  \begin{itemize}
    \item \verb#(resourceManagerContact=<string>)#
    \item \verb#(directory=<string>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(project=<string>)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(label=<string>)#
    \item \verb#(subjobCommsType=blocking-join|independent)#
    \item \verb#(subjobStartType=strict-barrier|loose-barrier|no-barrier)#
  \end{itemize}

  \subsection{Unsupported GRAM RSL attributes}

  \begin{itemize}
    \item \verb#(directory=<string>)#
    \item \verb#(fileCleanUp=<array>)#
    \item \verb#(fileStageIn=<array>)#
    \item \verb#(fileStageInShared=<array>)#
    \item \verb#(fileStageOut=<array>)#
    \item \verb#(gassCache=<path>)#
    \item \verb#(gramMyJob=independent|collective)#
    \item \verb#(hostCount=<number>)#
    \item \verb#(jobType=<string>)#
    \item \verb#(libraryPath=<path>)#
    \item \verb#(maxCpuTime=<time>)#
    \item \verb#(maxWallTime=<time>)#
    \item \verb#(maxTime=<time>)#
    \item \verb#(maxMemory=<memory>)#
    \item \verb#(minMemory=<memory>)#
    \item \verb#(project=<string>)#
    \item \verb#(remoteIoUrl=<string>)#
    \item \verb#(scratchDir=<string>)#
  \end{itemize}

  \chapter{XRSL and JSDL}
  \label{sec:jsdl}

  \input{xrsl-jsdl}

  \appendix

  \chapter{Examples}
  \label{sec:examples}

  \section{User-side xRSL script}

  \begin{verbatim}
&
(* test run: if "yes", only submits RSL without actual job start *)
     (dryRun="no")
(* some local variables defined for further convenience *)
     (rsl_substitution=("TOPDIR" "/home/johndoe"))
     (rsl_substitution=("NGTEST" $(TOPDIR)/ngtest))
     (rsl_substitution=("BIGFILE" "/scratch/johndoe/100mb.tmp"))
(* some environment variables, to be used by the job *)
     (environment=("ATLAS" "/opt/atlas") ("CERN" "/cern"))
(* the main executable file to be staged in and submitted to the PBS *)
     (executable="checkall.sh")
(* the arguments for the executable above *)
     (arguments="pal")
(* files to be staged in before the execution *)
     (inputFiles = ("be_kaons" "")
     ("file1" gsiftp://grid.uio.no$(TOPDIR)/remfile.txt)
      ("bigfile.dat" $(BIGFILE) )    )
(* files to be given executable permissions after staging in *)
     (executables="be_kaons")
(* files to be staged out after the execution *)
     (outputFiles=
        ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp")
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile")
        ("be_kaons.hbook"  gsiftp://ce1.grid.org$(NGTEST)/kaons.hbook)   )
(* user-specified job name *)
     (jobName="NGtest")
(* standard input file *)
     (stdin="myinput.dat")
(* standard output file *)
     (stdout="myoutput.dat")
(* standard error file *)
     (stderr="myerror.dat")
(* GM logs directory name *)
     (gmlog="gmlog")
(* flag whether to merge stdout and stderr *)
     (join="no")
(* request e-mail notification on status change *)
     (notify="bqfe john.doe@gmail.com jane.doe@mail.org")
(* maximal CPU time required for the job, minutes for PBS*)
     (CpuTime="60")
(* maximal time for the session directory to exist on the remote node, days *)
     (lifeTime="7")
 (* memory required for the job, per rank, Mbytes *)
     (Memory="200")
(* wall time to start job processing *)
     (startTime="2002-04-28 17:15:00")
(* disk space required for the job, Mbytes *)
     (Disk="500")
(* required architecture of the execution node *)
     (architecture="i686")
(* required run-time environment *)
     (runTimeEnvironment="APPS/HEP/Atlas-1.1")
(* number of re-runs, in case of a system failure *)
     (rerun="2")
  \end{verbatim}

  \section{GM-side xRSL script}

Note that the client tool must expand all the \texttt{rsl\_substitution}
values and add double quotes to \textbf{all} strings. Client tool is also expected to insert \texttt{queue} attribute in case such is missing and to add checksum values for input files to be uploaded by the client. Standard ARC client also
streamlines blank spaces.

Comment lines are removed from the GM-side xRSL; below they are shown
to explain details.

  \begin{verbatim}
&
(* saves RSL in a temporary file if "yes" *)
     ("savestate" = "yes" )
(* job submission to be performed if action is "request" *)
     ("action" = "request" )
(* submission host name *)
     ("hostname" = "ce1.grid.org" )
(* client software version *)
     ("clientsoftware" = "nordugrid-arc-0.6.0.3" )
(* walltime value added by the client, in seconds *)
     ("walltime" = "3600" )
(* test run: if "yes", only submits RSL without actual job start *)
     ("dryRun" = "no" )
(* some local variables defined for further convenience *)
     ("rsl_substitution" = ("TOPDIR" "/home/johndoe" ) )
     ("rsl_substitution" = ("NGTEST" "/home/johndoe/ngtest" ) )
     ("rsl_substitution" = ("BIGFILE" "/scratch/johndoe/100mb.tmp" ) )
(* some environment variables, to be used by the job *)
     ("environment" = ("ATLAS" "/opt/atlas" ) ("CERN" "/cern" ) )
(* executable *)
     ("executable" = "checkall.sh" )
(* arguments *)
     ("arguments" = "pal" )
(* files to be staged in before the execution *)
     ("inputfiles" = ("checkall.sh" "279320" )
        ("myinput.dat" "39806" )
        ("be_kaons" "8807" )
        ("file1" "gsiftp://grid.uio.no/home/johndoe/remfile.txt" )
        ("bigfile.dat" "104857600" )
     )
(* files to be given executable permissions after staging in *)
     ("executables" = "checkall.sh" "be_kaons" )
(* files to be staged out after the execution *)
     ("outputfiles" = ("file1" "gsiftp://grid.tsl.uu.se/tmp/file1.tmp" )
        ("100mb.tmp" "rls://rls.nordugrid.org:39281/test/bigfile" )
        ("be_kaons.hbook" "gsiftp://ce1.grid.org/home/johndoe/ngtest/kaons.hbook" )
        ("myoutput.dat" "" )
        ("myerror.dat" "" )
    )
(* user-specified job name *)
     ("jobName" = "NGtest" )
(* standard input file *)
     ("stdin" = "myinput.dat" )
(* standard output file *)
     ("stdout"="myoutput.dat" )
(* standard error file *)
     ("stderr" = "myerror.dat" )
(* flag whether to merge stdout and stderr *)
     ("join" = "no" )
(* request e-mail notification on status change *)
     ("notify" = "bqfe john.doe@gmail.com jane.doe@mail.org" )
(* specific queue to submit the job *)
     ("queue" = "atlas" )
(* CPU time required for the job, converted into seconds *)
     ("cputime" = "3600" )
(* maximal time for the session directory to exist on the remote node, seconds *)
     ("lifetime" = "604800" )
(* memory required for the job, per rank, Mbytes *)
     ("memory" = "200" )
(* wall time to start job processing *)
     ("startTime" = "20020128171500" )
(* disk space required for the job, Mbytes *)
     ("disk" = "500" )
(* required architecture of the execution node *)
     ("architecture" = "i686" )
(* required run-time environment *)
     ("runtimeenvironment" = "APPS/HEP/Atlas-1.1" )
(* number of re-runs, in case of a system failure *)
     ("rerun" = "2" )
(* original client xRSL with expanded string substitutions; shortened here *)
     ("clientxrsl" = "&(""dryrun"" = ""no"" )(... )(""rerun"" = ""2"" )" )
  \end{verbatim}

  \section{JSDL script}

  \begin{verbatim}

<?xml version="1.0" encoding="UTF-8"?>
<jsdl:JobDefinition
  xmlns="http://schemas.ggf.org/jsdl/2005/11/jsdl"
  xmlns:posix="http://schemas.ggf.org/jsdl/2005/11/jsdl-posix"
  xmlns:arc="http://www.nordugrid.org/ws/schemas/jsdl-arc">
  <JobDescription>
    <!-- Generic job description (jobname in XRSL) -->
    <JobIdentification>
        <JobName>JSDL Hello World job</JobName>
        <Description>
          This is a simple Hello World job test
        </Description>
    </JobIdentification>
    <!-- Standard application (same as executable, arguments in XRSL) -->
    <Application>
        <ApplicationName>Echo</ApplicationName>
        <posix:POSIXApplication>
          <posix:Executable>/bin/echo</posix:Executable>
          <posix:Argument>Hello world</posix:Argument>
        </posix:POSIXApplication>
    </Application>
    <!-- Get the GM diagnostics (same as gmlog in XRSL) -->
    <arc:LocalLogging>
	      <arc:Directory>gmlog</arc:Directory>
    </arc:LocalLogging>
 </JobDescription>
</JobDefinition>

  \end{verbatim}

  \chapter{JSDL-ARC schema}
  \label{sec:jsdl-arc-schema}

  \input{jsdl-arc-schema}

  \bibliography{grid}
  \printindex

\end{document}